### 2020-09-17 paper/paper.tex
technical specification

\documentclass[runningheads]{llncs}

\usepackage{listings}
\usepackage[colorlinks]{hyperref}

\usepackage{xcolor}

\usepackage[ruled]{algorithm2e}
\usepackage{amsmath,amsfonts}

\usepackage[utf8]{inputenc}

\lstset{
    basicstyle=\fontsize{9}{9}\selectfont\ttfamily,
    keywordstyle=\color{blue!50}\bfseries
}

\usepackage{tikz}
\usetikzlibrary{decorations.pathreplacing, patterns}

\usepackage{subcaption}

\title{Verification of a TCP/IP stack}

\author{Guillaume Cluzel\inst{1,2} \and
        Yannick Moy\inst{1} \and
        Cl\'ement Zeller\inst{3}}

        \institute{Adacore \and
        \'Ecole Normale Sup\'erieure de Lyon \and
        Oryx Embedded}

\begin{document}

\authorrunning{G. Cluzel et al.} %, Y. Moy and C. Zeller}

\maketitle

\begin{abstract}
    The abstract should briefly summarize the contents of the paper in
    150--250 words.

    \keywords{First keyword  \and Second keyword \and Another keyword.}
\end{abstract}

\section{Introduction}

    In the development of the Internet, the TCP protocol that has a very special place.
    Initially described by the ``fathers of the Internet'' Vincent Cerf and Bob Kahn, the TCP protocol was standardized in 1981 by RFC 793~\cite{rfc793}.
    Since this date, TCP has been widely used on the Internet as the underlying protocol for application layers such as HTTP,
    FTP or TLS. TCP is based on the underlying IP protocol which is itself
    based on a link layer such as Ethernet. This stacking of protocols is referred to as the TCP/IP stack.

    Given the importance of the TCP protocol in network communications, some researchers have attempted to formalize the protocol to either prove key properties~\cite{smith1996formal}
    or even to fully prove the correctness of an implementation against a complete specification~\cite{ridge2008rigorous}.
    This latter implementation was performed in the HOL proof assistant, and while the authors note that it would be possible in theory to extract an implementation in Haskell from
    this work, this has not been adopted in practice.
    The formalization of TCP specification and verification of a TCP implementation face two difficulties:
    RFC 793 is written in English, leaving many parts underspecified; the TCP protocol is inherently concurrent which makes it complex to specify and verify.
    As a result, there is currently no formally verified implementation for TCP that is usable in industry.

    Other communication protocols have been formalized in recent years, leading to a formally verified implementation.
    The most prominent project in this area is the Everest project~\cite{bhargavan2017everest}, which led to a formally verified implementation in F* of TLS called miTLS~\cite{bhargavan2013implementing}.
    Note that miTLS still relies on an untrusted implementation of TCP, and thus could benefit from the results of our work.
    The approach taken in project Everest was to reimplement the formalized protocols completely in F*, a research programming language targeted at formal program verification.
    This approach is often difficult to adopt in industry, as codebases are evolved incrementally rather than replaced, both for economical and practical reasons.

    Our approach in this work has been to incrementally evolve an existing TCP/IP stack in C by replacing parts of the code with formally verified code in SPARK.
    The stack chosen, CycloneTCP\footnote{\url{https://www.oryx-embedded.com/products/CycloneTCP.html}}, is targeted at embedded processors with no Operating System.
    In Section~\ref{sec:TCP} we describe the TCP protocol, in particular the state machine that governs the behavior of a connection.
    In Section~\ref{sec:spec} we present specificities of the specification technique that we used, and in Section~\ref{sec:verif} we explain how we formalized and implemented in SPARK the TCP user functions and the underlying socket API.
    In Section~\ref{sec:results} we present the results of our work, in particular bugs that have been found and eliminated thanks to our approach, before we conclude.

\section{TCP protocol and CycloneTCP library}
\label{sec:TCP}

\subsection{TCP protocol}

    %\input{Automaton.tex}

In this part we should present:
\begin{itemize}
    \item The TCP protocol
    \begin{itemize}
        \item Different states
        \item Flags that can be received (mainly SYN, ACK, FIN \& RST)
        \item State machine
        \item User functions
        \item Precise that most of transitions are done when a message is received and then, outside of the user functions.
    \end{itemize}
\end{itemize}

Figures à inclures ici:
\begin{itemize}
    \item Une figure qui présente l'automate de transition. Expliquer que certaines transition sont effectuées par une action de l'utilisateur,
          et que d'autres sont effectuées quand un segment est reçu.
\end{itemize}

\subsection{CycloneTCP library organisation}

    \input{OryxStack.tex}

    The CycloneTCP library is TCP/IP stack that implements network protocols.
    Figure \ref{Fig:TcpStack} gives an overview of the different protocols supported. TCP and UDP protocols have a central place
    in the stack. An API is provided to use those network protocols: the Socket API. It is a common way to use transport protocols,
    and the application protocols in the stack, such as HTTP or FTP use it.

    The stack is particularly well adapted to embedded platforms and microcontrolers. As a result, the code is written
    to compile on those platforms and to be used with different real time OS, like for intance FreeRTOS.
    Hence, a model based on multiple tasks has been adopted to implement the TCP protocol as defined in the norm, with,
    in addition, complex cooperation mechanisms. A task is dedicated to the user code, another to timers, and the last is
    assigned to the processing of incoming segments. Then, user tasks can block without interfering the TCP/IP stack.

\section{Specification Techniques Used}
\label{sec:spec}

\subsection{The SPARK Programming Language}

Ada is a general-purpose procedural programming language. The design of the Ada
language puts great emphasis on the safety and correctness of the program. This
objective is realized by using a readable syntax that uses keywords instead of
symbols where reasonable. The type system is strong and strict and many
potential violations of type constraints can be detected statically by the
compiler. If not, a run-time check is inserted into the program, to guarantee
the detection of incorrect situations. Ada 2012 introduced contract based
programming to Ada. In particular, it is possible to attach pre- and
postconditions to subprograms\footnote{In Ada, a distinction is made between
  functions that return a value, and procedures, which do
  not. \emph{Subprogram} is the term that designates both.}.  These conditions
can be checked during the execution of the program, just like assertions.

SPARK~\cite{mccormick2015building} is the name of a platform that provides
formal verification for Ada. It uses the user-provided contracts and attempts
to prove that the runtime checks cannot fail and that postconditions are
established by the corresponding subprograms.  As formal verification for the
whole Ada language would be intractable, SPARK is also the name of the subset
of the Ada language that is supported by the SPARK tool called
GNATprove\footnote{\url{http://docs.adacore.com/spark2014-docs/html/ug/}}.
This subset contains almost all features of Ada, though sometimes in a
restricted form.  In particular, expressions should be free from side effects,
and aliasing is forbidden (no two variables should share the same memory
location or overlap in memory), including when using pointers thanks to the use
of an ownership policy~\cite{dross2020recursive}.  This restriction greatly
simplifies the memory model used in the SPARK tool: any program variables can
be reasoned about independently from other variables.

GNATprove performs two kinds of analysis modularly on individual subprograms:
information flow to detect reads of uninitialized data and violations of flow
contracts; deductive verification based on the Why3 platform to generate
verification conditions for SMT solvers via a weakest-precondition calculus to
detect possible runtime errors and violations of functional contracts.

\subsection{Interfacing SPARK and C Code}

    To reduce the size of the work, only few functions have been totaly translated
    in SPARK. Then, some C functions are reused in SPARK, and some SPARK functions
    are exported to be used in C code. The only constraint is that the
    declarations of both the SPARK and the C functions must match.
    The attribute \lstinline{Import} can be added to the declaration of a
    subprogram in SPARK to specify to the compiler that the function is defined
    in C. Symmetrically, the attribute \lstinline{Export} allows the subprogram
    to be reused in C. An \lstinline{External_Name} might be given for the
    programmers and the linker to locate the right symbol at the liking time.

    The types in SPARK can be more expressive than the ones used in C. However
    they must have the same representation in memory to be used as an argument
    of a function. The attribute \lstinline{Size} can be added to SPARK types
    to keep them coherent with the C types. It specifies to the compiler the
    size in bits requiered to store an object of this type.

    In addition, contracts can be added on imported subprograms. In this case,
    SPARK use their contracts to analyze other subprograms that call these imported
    subprograms, but their contracts are not proved. It is the responsability of the
    user to ensure that the contract he writes is correct.

\subsection{Dealing with Pointers}

    In CycloneTCP library, pointers are often manupulated at top-level and
    transmitted to C functions.  They are also used as fields in structures,
    for instance in the socket structure to store other complex data-structures
    and recursive pointer based data-structures.
    SPARK use a pointer ownership model to prevent bugs involving pointers such
    as memory leak or double free. When memory is allocated, the pointer pointed
    to this memory area is owned. The ownership of this memory can be moved.
    The memory must be deallocated before the end of the execution fo the program,
    and a pointer has to be owned at every point of the program.

% both at top-level and as fields in structures. explain how it relies on the
% ownership model. maybe a little bit on dynamic allocation/deallocation (which
% will introduce the section later on memory leaks)

\subsection{Specifying the Frame Condition}

    The socket structure is modified through the TCP user functions. In particular
    the state of the socket that we are interested in is modified and we want to
    track these modifications. But a lot of fields of the socket structure are not
    interesting to prove the correctness of the states transitions. Then, it is
    necessary to create a model of the socket to describe the revelant field of
    the socket to keep. Two models have been created, the first \lstinline{Model}
    that tracks the TCP state of the socket as well as basic fields of the socket,
    and the second \lstinline{Basic_Model} that only tracks the basic fields and
    loss all information about the TCP state.

    The post-conditions use these models in such way
    \begin{lstlisting}[language=Ada]
    Model(Sock) = (Model(Sock)'Old with delta
                        State => TCP_STATE_CLOSED)
    \end{lstlisting}
    where the \lstinline{Old} in \lstinline{Model(Sock)'Old} refers to the model
    of the socket before the call of the function, and \lstinline{with delta} is
    a delta aggregates that yield a copy of the old model with a new value for
    the field \lstinline{State} modified.

    The models used in contracts are based on ghost code. The ghost code is only
    used for the proofs with GNATprove, and is removed by the compiler. Thus, the size
    of the code does not increase by adding contracts.

    % describe how models of sockets are defined and used in contracts, based on
    % ghost code, delta aggregates and Old attribute

\section{Verification of the TCP protocol and the Socket API}
\label{sec:verif}

    Our work is focused on the key elements of the CycloneTCP library: the sockets and TCP protocol itself.
    These two parts have been replaced by SPARK code and the absence of run-time errors has been proved for
    these parts.
    However, to be integrated with the existing library, it has been necessary to reuse existing C code
    as an untrested black box, for which we cannot have strong guarantees. Figure \ref{Fig:TcpStack} gives an overview of
    what have been translated and what have been reused. The box for TCP is hatched because only the user level functions
    has been translated in SPARK. The parsing and processing of incoming segment is let to the C code.
    Thus, other techniques and tools have helped the verification process, in particular for the TCP part. They are presented in this section.
    The source code of the reworked TCP/IP stack can be found on Github\footnote{\url{https://github.com/AdaCore/Http_Cyclone}}.

\subsection{Hardening the user's socket API}

\begin{figure}[t]
\begin{subfigure}{\textwidth}
\begin{lstlisting}[language=C, frame=bottomline]
Socket *socketOpen(uint_t type, uint_t protocol);
error_t socketConnect(Socket *socket, IpAddr *remoteIpAddr,
                      uint16_t remotePort);
\end{lstlisting}
\caption{C interface.}
\label{Fig:socketInterface:C}
\end{subfigure}
\begin{subfigure}{\textwidth}
\begin{lstlisting}[language=Ada, frame=bottomline]
type Socket_Struct is private;
type Socket is access Socket_Struct;
subtype Not_Null_Socket is not null Socket;
type Socket_Type is (SOCKET_TYPE_STREAM, SOCKET_TYPE_DGRAM);
type Socket_Proto is (SOCKET_IP_PROTO_TCP, SOCKET_IP_PROTO_UDP);

procedure Socket_Open
    (Sock       :    out Socket;
     S_Type     : in     Socket_Type;
     S_Protocol : in     Socket_Protocol)
with
    Post => (if Sock /= null then
        Sock.S_Type = S_Type and then
        not Is_Initialized_Ip(Sock.S_Remote_Ip_Addr) and then
        not Is_Initialized_Ip(Sock.S_localIpAddr));

procedure Socket_Connect
    (Sock           : in out Not_Null_Socket;
     Remote_Ip_Addr : in     IpAddr;
     Remote_Port    : in     Port;
     Error          :    out Error_T)
with
    Pre => Is_Initialized_Ip (Remote_Ip_Addr) and then
           Remote_Port > 0,
    Post => (if Error = NO_ERROR then
            Is_Initialized_Ip(Sock.S_localIpAddr) and then
            Sock.S_localIpAddr = Remote_Ip_Addr);
\end{lstlisting}
\caption{SPARK interface.}
\label{Fig:socketInterface:SPARK}
\end{subfigure}
\caption{Comparison between the C and the SPARK interface for \texttt{Socket\_Open} et \texttt{Socket\_Connect}.}
\label{Fig:socketInterface}
\end{figure}

    The socket API is a set of functions used to communicate with the protocol stack. These are therefore the functions used
    by programmers in their programs to use network protocols. However, one of the most significant problems pointed to by the
    primary author of the CycloneTCP library is the incorrect usage of the library's API. More specifically, the library's users
    tend to call the API functions with wrong arguments and they tend to forget to check the value of the return code of the
    function call before continuing the processing. This can lead to incorrect program behaviors when an error occurs in a function
    and it can be difficult to debug.

    In order to design a better socket API, we have taken advantage of the strong-typing characteristic of Ada,
    and the features of the SPARK language, in particular the functional specifications.

\subsubsection{Using types to carry information}

    Strong-typing is a first mechanism to ensure correctness of programs and exclude a range of programing errors. Meaningful names for
    variables and types give a hint about the subprogram behavior, but also the valid input values accepted.
    Figure \ref{Fig:socketInterface} is a comparison between the interface of the C functions of the original library
    \lstinline{socketOpen} and \lstinline{socketConnect} and their SPARK version \lstinline{Socket_Open} and
    \lstinline{Socket_Connect}. The C function \lstinline{socketOpen} takes two arguments, the type of the socket and the protocol,
    both as unsigned integers and it returns a new socket in case of success.
    Its SPARK equivalent takes its two arguments, type and protocol, as enumeration values, which ensures correct calls
    of this function. Indeed, all the unsigned values that do not correspond to any protocol or any type of socket are eliminated.

    In SPARK, types can be constrained by a valid range of values. Another kind of constraint on types can appear in
    access types. The example on figure \ref{Fig:socketInterface:SPARK} shows a such mechanism for the type \lstinline{Not_Null_Socket},
    that is a socket with a constraint that states that the pointer has to be not null. The function \lstinline{socketConnect},
    used to connect to a remote server, requires a socket not null. While the C function contains defensive code to ensure that
    the socket passed as first argument is not null, \lstinline{Socket_Connect} directly uses information on the type of the argument
    \lstinline{Sock} to know that the socket is not null. Moreover, this information can be checked statically with GNATprove.
    Then removing the defensive code has two interests: first, it reduces the size of the program by removing lines of code,
    and secondly it reduces the number of instructions to execute, and then the global execution time.

\subsubsection{Order of the functions call}

    The TCP connection follows an order of execution as seen in the section \ref{sec:TCP}: the connection is first opened,
    then the data are sent and received, and finally the connection can be closed. Since the sockets are an interface to use
    network protocols, they follow the same order of execution. Thus, an order can be enforced thanks to the pre- and postconditions.
    If a function $f_1$ must be executed before a function $f_2$, the strategy is to add a postcondition on the socket on $f_1$
    that is also a precondition for $f_2$. Then, for the precondition of $f_2$ to be true, the function $f_1$ must have been called first.
    Such postcondition is presented on figure \ref{Fig:socketInterface:SPARK} for the function \lstinline{Socket_Connect}. After its
    execution, the field \lstinline{S_Remote_Ip_Addr} of the socket \lstinline{Sock}. Then, since all the functions to send or receive
    data and to close the connection require a connection established, these functions have a precondition that contains a condition
    that states that the condition is established, which can be written as
    \begin{lstlisting}
Pre => Is_Ip_Initialized(Sock.S_Remote_Ip_Addr);
    \end{lstlisting}
    and because the function \lstinline{Socket_Connect} is the only one that can initialize this value, the order is respected
    if and only if the code is valid.

    The same mechanism is used with the functions \lstinline{Socket_Open} that create a valid socket when the function
    succeeds and \lstinline{Socket_Connect} that require a valid not null socket.

    GNATprove can prove that the order of the calls is respected, and then it can invalidate an incorrect program without requiring any
    tests. It helps programmers to write correct programs.

\subsubsection{Checking error codes}

    Most of the functions of the socket API can fail. To know if the function has failed or not, and error code is returned.
    It must be checked by the caller before continuing processing at the risk of leading to an incorrect state.
    In the previous paragraph we explained that the functions that send or receive data check in their precondition if the connection
    has been initialized or not, by checking the field \lstinline{S_Remote_Ip_Addr}, and we explained that this value is initialized
    by \lstinline{Socket_Connect}. The presence in the postcondition of \lstinline{Socket_Connect} or if statement
    \lstinline{(if Error = NO_ERROR then ...)} makes that \lstinline{S_Remote_Ip_Addr} is initialized only when no error occurs during
    the execution of the function. Thanks to the mechanism exposed in the previous section the next function can only be called
    if it is known that no error has occurred previously.

    Let \lstinline{Socket_Send} be the function to send data. The following example of code is incorrect because if \lstinline{Error} is
    different to \lstinline{NO_ERROR}, nothing ensures that \lstinline{S_Remote_Ip_Addr} is initialized.
    \begin{lstlisting}[language=Ada]
Socket_Connect(Sock, Remote_Ip_Addr, Remote_Port, Error);
-- GNATprove: medium: precondition might fail.
Socket_Send(Sock, Data, Written, Flags, Error);
    \end{lstlisting}
    In comparison this second example of code produces no error because when \lstinline{Socket_Send} is called, GNATprove knows that
    \lstinline{ERROR} is equal to \lstinline{NO_ERROR} and then \lstinline{S_Remote_Ip_Addr} is initialized.
    \begin{lstlisting}[language=Ada]
Socket_Connect(Sock, Remote_Ip_Addr, Remote_Port, Error);
if Error = NO_ERROR then
    Socket_Send(Sock, Data, Written, Flags, Error);
end if;
    \end{lstlisting}

    The same mechanism can be observed on figure \ref{Fig:socketInterface:SPARK}. The user has to test if the socket returned by
    \lstinline{Socket_Open} is null or not before calling \lstinline{Socket_Connect}, otherwise GNATprove will emit a message.

    Thanks to this use of GNATprove, one can ensure the correct use of the socket API, by ensuring a correct order for the
    calls of the functions and ensuring that error codes are checked.

\subsection{Conformance to the TCP protocol}

    The user functions of the TCP protocol have been rewritten in SPARK. Besides proving the absence of run-time error in these functions,
    it has been possible to check that some properties of the TCP protocol are not violated in the code.

\subsubsection{Extracting a specification from the TCP norm}

    Extracting a specification is mandatory to provide later functional contracts to the user functions.
    That can be achieved by reading the norm. However, the norm is underspecified in many cases and only one possible implementation
    is depicted. Thus, the specifications we have extracted are conformant with the norm but they differ a little to
    what can be found in the RFC.

    Among the specifications we have extracted:
    \begin{itemize}
        \item The transitions between the states must respect the order given by the state machine described
              in figure \ref{Fig:statemachine}.
              In particular, a transition from a state to another is valid only if there exists a transition in the state machine
              between these two states.
        \item User functions contains functional specifications in the norm, that describe what have to be done when the function
              is called depending on the state of the socket~\cite[p.~52]{rfc793}.
              In particular, a socket in some state will return an error for certain calls of user functions. Figure \ref{Fig:constraint}
              lists all the constraints on the state of the socket before and after the call.
    \end{itemize}

\subsubsection{Rewriting TCP user functions}

    The user functions have been translated in SPARK to prove the absence of run-time errors in them. The low level IP functions
    that are called are assumed to be correct and then are used as an untrusted black box. They are supposed to not interact with TCP
    special fields. However, there exists another kind of subprogram that interacts and can modify the state of the socket
    during the execution of the program: the procedure \lstinline{Tcp_Wait_For_Events}. This procedure is called when the code needs
    to wait for an event to become true. The event waited can become true, for example, when a segment is received. But when a
    segment is received, the state of state can change. Then it is necessary to know what are the exact possible states when the
    procedure \lstinline{Tcp_Wait_For_Events} returns if we want to be sure that there is no violation of the norm in the further transitions.

    SPARK does not have a very advanced native way to deal with concurrency. In particular, it is not possible for the tool to infer precisely the
    interferences that comes from other tasks. Thus, using SPARK is not the right way to do, and a manual solution has been considered. It consists
    of analysing what is done in other tasks. Then a contract can be added to \lstinline{Tcp_Wait_For_Events} to describe the possible
    interractions of the other tasks during the wait.

\subsubsection{The use of KLEE to extract contracts}

    When a TCP segment is received, that segment is processed through a C function named \lstinline{tcpProcessSegment}.
    Depending on the flags contains in the segment, and the state of the socket at the moment of the reception of the segment, different
    transitions can be done. Reading the code might be enough to infer contracts connecting the state before the reception of the segment
    and after. But from experience it is not enough and a more formal approach is necessary. Translating this function in SPARK to analyze
    its behavior would have been a correct way to proceed, but because of a lack of time, another method has been chosen: symbolic execution.

    Symbolic execution is a means to execute a program using symbolic variables instead of concrete values. It accumulates contraints
    on those values which are used to generate proof obligations. Symbolic execution has been successfully used with other verification
    methods. Indeed, in~\cite{vanoverberghe2008using} the authors present how to use symbolic execution to improve deductive verification,
    and in~\cite{kassios2012comparing} the authors explain that symbolic execution can be good at infering contracts.
    For our work, KLEE has been used. It is built on top of LLVM as it takes LLVM bytecode in input
    and works well with C code thanks to a very simple interface.

    In order to extract contracts from the C function \lstinline{tcpProcessSegment}, we have started by building a small main function that
    simulates the reception of a segment. This function is then executed by KLEE. More precisely, first, with the KLEE intrinsic functions, we
    create a symbolic segment (representing any segment that can be received). Secondly, the socket is changed for a particular state $S_1$. Thirdly, the
    function is executed by KLEE. After its call, an assertion about the states $S'_1, \dots, S'_k$ of the socket after the execution
    of the function is added to check what are the resulting states. The states $S'_1, \dots, S'_k$ are extracted from the norm and adapted to the CycloneTCP
    implementation that can differ on some points.
    After this work, we then have a postcondition for \lstinline{tcpProcessSegment}, that can be reported in the SPARK code for further proofs, of the form:
    \begin{lstlisting}[language=Ada,mathescape=true]
    if Sock.State'Old = $S_1$ then Sock.State in $S'_1$ | $S'_2$ | ... | $S'_k$
    \end{lstlisting}

\subsubsection{Solution to the concurrency challenge}

    The previous part shows how we can compute all the transitions that can be done when a single message is received.
    It is now necessary to understand how the waiting procedure \lstinline{Tcp_Wait_For_Events} works.
    When it is called, it checks if the event is already true, thanks to a call to the function \lstinline{Tcp_Update_Events}.
    If it is not the case, the mutex that enforces the mutual exclusion of access to the socket is unlocked.
    Messages can be received until the event waited becomes true.
    Once a reception makes true the event, the function \lstinline{Tcp_Wait_For_Events} locks again the mutex and returns.
    A possible algorithm to simulate the reception of segment and compute the least set of states that can result of the wait can be done
    by the algorithm \ref{algo:waitForEvents}.

    \begin{algorithm}
        \SetKwFunction{ProcessSegm}{\lstinline[language=Ada]{Tcp\_Wait\_For\_Events}}
        \SetKwProg{Fn}{function}{}{end}
        \Fn{\ProcessSegm{Socket, Event, Event\_Mask}}{
            $S_{last} := \text{\textit{Socket}}$\;
            $S := S_{last}$\;
            $E :=$ \lstinline[language=Ada]{Tcp_Update_Events}($S_{last}$)\;
            \If{$(E\ \&\ \text{Event\_Mask}) \neq 0$}{
                \KwRet{$S$}\;
            }
            \For{$i=1$ \KwTo $3$}{
                $S_{last} :=$ \lstinline[language=Ada]{Tcp_Process_One_Segment}($S_{last}$) \;
                $S := S \cup S_{last}$\;
                $E :=$ \lstinline[language=Ada]{Tcp_Update_Events}($S_{last}$)\;
                \If{$(E\ \&\ \text{Event\_Mask}) \neq 0$}{
                    \KwRet{$S$}\;
                }
            }
            \KwRet{$\emptyset$}\;
        }
        \caption{Function to compute the possible states after a wait for a particular event.}
        \label{algo:waitForEvents}
    \end{algorithm}

\section{Results and bugs found}
\label{sec:results}

    \begin{itemize}
        \item Bug de concurrence -> SPARK
        \item Memory leak -> SPARK
        \item Problèmes trouvés.
        \item Résultats sur la nouvelle implémentation (performance, taille du code).
        \item Activation des préconditions pour voir si elles sont toutes valides.
    \end{itemize}

\section{Conclusion and future work}

    The use of SPARK for the verification of the TCP protocol has given good results. It has helped to find bugs
    in the existing implementation. But more than finding bugs, the interest of this work lie in the face that it is
    a big step toward a secure implementation. It has been proved by the tool GNATprove that our implementation is free
    of run-time errors and that all the transitions in the TCP state machine are done with respect to the norm.

    However weeknesses exist our implementation, in particular because all the underlying layers are still written in C.
    The principal function of those layers is to format packet before they are sent, or parse incoming packets, check
    their integrity and transmit the resulting payload to the corresponding upper level layer.
    This processing part can be a source of errors and bugs as explain in~\cite{Reiher2019RecordFluxFM}.
    Using the RecordFlux DSL to parse the packets of the different protocols is the next step to make the stack safer.
    Using RecordFlux to parse TCP message would be a big step to finish the translation of the TCP protocol in SPARK.
    The postconditions extracted by KLEE could be proved by GNATprove for a safer result.

    Our verification of the TCP protocol only focuses on the validity of the transitions in the TCP state machine,
    but some other properties are formulated in the TCP norm, and could be verified in the implementation to make
    it even more robust.

\bibliographystyle{splncs04}
\bibliography{biblio}

\end{document}

### 2020-10-22 paper/Automaton.tex
Description du protocol TCP

\begin{figure}[t]
{\centering\scalebox{.7}{
\begin{tikzpicture}[>=latex]
    \def\tfrac#1#2{\ensuremath{\displaystyle\frac{\text{\footnotesize#1}}{\text{\footnotesize#2}}}}
    %
    % Styles for states, and state edges
    %
    \tikzstyle{state} = [draw, very thick, fill=white, rectangle, minimum height=3em, minimum width=7em, node distance=8em, font={\sffamily\bfseries}]
    \tikzstyle{stateEdgePortion} = [black,thick];
    \tikzstyle{stateEdge} = [stateEdgePortion,->];
    \tikzstyle{edgeLabel} = [pos=0.5, text centered, font={\sffamily\small}];

    %
    % Position States
    %
    \node[state, name=closedStart] {CLOSED};
    \node[state, name=listen, below of=closedStart] {LISTEN};
    \node[state, name=synSent, below of=listen, right of=listen, xshift=8em] {SYN\_SENT};
    \node[state, name=synRcvd, below of=listen, left of=listen, xshift=-8em] {SYN\_RCVD};
    \node[state, name=established, below of=listen, node distance=14em] {ESTABLISHED};
    \node[state, name=finWait1, below of=established, left of=established, node distance=7em, xshift=-9em] {FIN\_WAIT\_1};
    \node[state, name=finWait2, below of=finWait1] {FIN\_WAIT\_2};
    \node[state, name=closeWait, below of=established, right of=established, node distance=7em, xshift=9em] {CLOSE\_WAIT};
    \node[state, name=closing, below of=established, node distance=14em] {CLOSING};
    \node[state, name=lastAck, below of=closeWait] {LAST\_ACK};
    \node[state, name=timeWait, below of=closing] {TIME\_WAIT};

    %
    % Connect States via edges
    %
    \draw ($(closedStart.south) + (-.5em,0)$)
    edge[stateEdge] node[edgeLabel, xshift=-3em]{\tfrac{\emph{Passive open}}{--}}
    ($(listen.north) + (-.5em,0)$);
    \draw ($(listen.north) + (.5em,0)$)
    edge[stateEdge] node[edgeLabel, xshift=2em]{\tfrac{\emph{Close}}{--}}
    ($(closedStart.south) + (.5em,0)$);

    \draw ($(listen.south) + (-1em,0)$)
    edge[stateEdge, bend left=22.5] node[edgeLabel, xshift=-2em, yshift=2em]{\tfrac{rcv SYN}{snd SYN, ACK}}
    ($(synRcvd.east) + (0,1em)$);

    % \draw ($(synRcvd.north) + (.5em,0)$)
    % edge[stateEdge, bend left=45] node[edgeLabel,xshift=-4em]{\emph{Timeout}/RST}
    % ($(closedStart.west) + (0,-.5em)$);

    \draw ($(synSent.north) + (-.5em,0)$)
    edge[stateEdge, bend right=45] node[edgeLabel,xshift=-1em, yshift=-2em]{\tfrac{\emph{Close}}{--}}
    ($(closedStart.east) + (0,-.5em)$);
    \draw ($(closedStart.east) + (0,.5em)$)
    edge[stateEdge, bend left=45] node[edgeLabel,xshift=4em]{\tfrac{\emph{Active open}}{snd SYN}}
    ($(synSent.north) + (.5em,0)$);

    \draw (synSent.west)
    edge[stateEdge] node[edgeLabel, yshift=1.3em]{\tfrac{rcv SYN}{snd SYN, ACK}}
    (synRcvd.east);
    \draw (synRcvd)
    edge[stateEdge] node[edgeLabel, xshift=-2.5em]{\tfrac{\emph{Close}}{snd FIN}}
    (finWait1);

    \draw ($(synRcvd.east) + (0,-1em)$)
    edge[stateEdge, bend left=22.5] node[edgeLabel, xshift=-1.5em, yshift=-2em]{\tfrac{rcv ACK}{--}}
    ($(established.north) + (-1em,0)$);
    \draw ($(synSent.west) + (0,-1em)$)
    edge[stateEdge, bend right=22.5] node[edgeLabel, xshift=3em, yshift=-1em]{\tfrac{rcv SYN, ACK}{snd ACK}}
    ($(established.north) + (1em,0)$);

    \draw ($(established.south) + (-1em,0)$)
    edge[stateEdge, bend left=22.5] node[edgeLabel, xshift=-1em, yshift=1.5em]{\tfrac{\emph{Close}}{snd FIN}}
    ($(finWait1.east) + (0,.5em)$);
    \draw ($(established.south) + (1em,0)$)
    edge[stateEdge, bend right=22.5] node[edgeLabel, xshift=1em, yshift=1.5em]{\tfrac{rcv FIN}{snd ACK}}
    ($(closeWait.west) + (0,1em)$);

    \draw (finWait1.south)
    edge[stateEdge] node[edgeLabel, xshift=-2em]{\tfrac{rcv ACK}{--}}
    (finWait2.north);
    \draw ($(finWait1.east) + (0,-.5em)$)
    edge[stateEdge, bend left=22.5] node[edgeLabel, yshift=0em, xshift=4.5em]{\tfrac{rcv FIN}{snd ACK}}
    (closing.north);
    \draw (finWait1.south east)
    edge[stateEdge] node[edgeLabel, xshift=0em, yshift=2em, text width=3em]{\tfrac{rcv FIN, ACK}{snd ACK}}
    (timeWait.north west);

    \draw (finWait2.south)
    edge[stateEdge, bend right=22.5] node[edgeLabel, xshift=-2em, yshift=-1em]{\tfrac{rcv FIN}{snd ACK}}
    (timeWait.west);

    \draw (closing)
    edge[stateEdge] node[edgeLabel, xshift=2.5em]{\tfrac{rcv ACK}{--}}
    (timeWait);

    \draw (closeWait)
    edge[stateEdge] node[edgeLabel,xshift=-2.5em]{\tfrac{\emph{Close}}{snd FIN}}
    (lastAck);

    %
    % Connect lastAck to closed is slightly more complicated
    % no direct line-of-sight, so we need to take the scenic route
    %
    \coordinate (lastAck2ClosedA) at ($(lastAck.east) + (2em,0)$);
    \coordinate (lastAck2ClosedB) at ($(closedStart.north -| lastAck.east) + (2em,1em)$);
    \coordinate (lastAck2ClosedC) at ($(closedStart.north) + (0.5em,1em)$);
    \draw (lastAck.east) edge[stateEdgePortion] (lastAck2ClosedA);
    \draw (lastAck2ClosedA) edge[stateEdgePortion] node[edgeLabel,xshift=-2.5em, yshift=-4em]{\tfrac{rcv ACK}{--}} (lastAck2ClosedB);
    \draw (lastAck2ClosedB) edge[stateEdgePortion] (lastAck2ClosedC);
    \draw (lastAck2ClosedC) edge[stateEdge] ($(closedStart.north) + (0.5em,0)$);

    %
    % likewise for timeWait to closed
    %
    \coordinate (timeWait2ClosedA) at ($(timeWait.south) + (0,-1em)$);
    \coordinate (timeWait2ClosedB) at ($(timeWait.south -| finWait2.west) + (-2em,-1em)$);
    \coordinate (timeWait2ClosedC) at ($(closedStart.north -| finWait2.west) + (-2em,1em)$);
    \coordinate (timeWait2ClosedD) at ($(closedStart.north) + (-0.5em,1em)$);
    \draw (timeWait.south) edge[stateEdgePortion] (timeWait2ClosedA);
    \draw (timeWait2ClosedA) edge[stateEdgePortion] (timeWait2ClosedB);
    \draw (timeWait2ClosedB) edge[stateEdgePortion] (timeWait2ClosedC);
    \draw (timeWait2ClosedC) edge[stateEdgePortion]
    node[edgeLabel, text width=12.25em, yshift=1.5em]{\emph{Timeout after two maximum segment lifetimes (2*MSL)}}
    (timeWait2ClosedD);
    \draw (timeWait2ClosedD) edge[stateEdge] ($(closedStart.north) + (-0.5em,0)$);

    % draw dotted lines around passive and active closes
    % \begin{pgfonlayer}{background}
    %     \draw [join=round,black,dotted] ($(closeWait.north west) + (-1em, -1em)$) rectangle ($(lastAck.south east) + (1em, 1em)$);
    %     \draw [join=round,black,dotted] ($(finWait1.north west) + (-1em, -1em)$) rectangle ($(timeWait.south east) + (1em, 1em)$);
    % \end{pgfonlayer}

\end{tikzpicture}}\par}
\caption{TCP state machine.}
\label{Fig:statemachine}
\end{figure}

### 2020-10-22 paper/paper.tex
Description du protocol TCP

@@
 
-\usepackage{xcolor}
+\ifLuaTeX
+    \usepackage{fontspec}
+    \IfFontExistsTF{Fira Code}
+    {\setmonofont[Scale=MatchLowercase]{Fira Code}}
+    {\IfFontExistsTF{DejaVu Sans Mono}{
+        \setmonofont{DejaVu Sans Mono}[Scale=MatchLowercase]
+    }{}}
+\fi
+
+\usepackage[x11names]{xcolor}
 
@@
     basicstyle=\fontsize{9}{9}\selectfont\ttfamily,
-    keywordstyle=\color{blue!50}\bfseries
+    keywordstyle={[1]\color{Blue4}\bfseries}, % keywords style
+    keywordstyle={[2]\color{Pink4}}, % type style
+    keywordstyle={[3]\color{OrangeRed3}}, % function style
+    literate =
+        {=>}{{=>}}2
+        {/=}{{/=}}2,
+    commentstyle=\itshape\color{gray}
 }
+
+\lstdefinestyle{C}{
+    language = C,
+    classoffset = 1,
+    morekeywords = {Socket, IpAddr, uint_t, uint16_t, error_t},
+    classoffset = 2,
+    morekeywords = {socketOpen, socketConnect},
+    classoffset = 0
+}
+
+\lstdefinestyle{Spark}{
+    language = Ada,
+    classoffset = 1,
+    morekeywords = {Socket_Struct, Socket, Not_Null_Socket, Socket_Type, Socket_Prot, IpAddr, Error_T, Port},
+    classoffset = 2,
+    morekeywords = {Socket_Open, Socket_Connect, Socket_Send},
+    classoffset = 7,
+    morekeywords = {Pre, Post},
+    keywordstyle = \bfseries,
+    classoffset = 0,
+}
+
+\let\state\textsf
+\let\flag\textsf
 
@@
 \usetikzlibrary{decorations.pathreplacing, patterns}
+\usetikzlibrary{calc, backgrounds}
 
@@
 
-    %\input{Automaton.tex}
+The TCP protocol is one of the most used protocol on the Internet.
+It is connection-oriented in the sense that a connection requires to be
+established between the two processes that exchange data. The protocol is
+governed by a state machine described in figure~\ref{Fig:statemachine}.
+The whole description of the protocol can be found in RFC~793~\cite{rfc793}.
+Its implementation is rather free and lots of choices are left to the programmer.
+However, the implementation must ensure some properties to be able to communicate
+with other TCPs.
+We present here the main elements that are useful for our verification.
 
-In this part we should present:
-\begin{itemize}
-    \item The TCP protocol
-    \begin{itemize}
-        \item Different states
-        \item Flags that can be received (mainly SYN, ACK, FIN \& RST)
-        \item State machine
-        \item User functions
-        \item Precise that most of transitions are done when a message is received and then, outside of the user functions.
-    \end{itemize}
-\end{itemize}
+\input{Automaton.tex}
 
-Figures à inclures ici:
-\begin{itemize}
-    \item Une figure qui présente l'automate de transition. Expliquer que certaines transition sont effectuées par une action de l'utilisateur,
-          et que d'autres sont effectuées quand un segment est reçu.
-\end{itemize}
+\subsubsection{States}
 
-\subsection{CycloneTCP library organisation}
+In this automaton, the states represent the state of the connection. \state{CLOSED}
+stands for a closed connection. Both states \state{SYN-SENT} and \state{SYN-RCVD} denotes
+the states that a TCP process can take during the initiation  of a new connection. The state
+\state{ESTABLISHED} represent the state when the connection is established. Then, two groups
+of states can be distinguished: the group formed by \state{FIN\_WAIT\_1}, \state{FIN\_WAIT\_2},
+\state{CLOSING} and \state{TIME\_WAIT} and that represents the states that a TCP can take
+when it is closing its connection and it is at the origin of the closing; and the group formed
+by the states \state{CLOSE\_WAIT} and \state{LAST\_ACK} when the connection is closing
+and the closing was initiated by the remote TCP.
+Finally, the state \state{LISTEN} stands for a TCP that waits for an incoming connection.
+
+\subsubsection{Flags}
+
+The transitions between the states are dictated by three things: user actions, the reception
+of flags and timers. When a transition is triggered, a response is given by sending of flags.
+The user actions, written in italic in the automaton, can be an active or
+passive opening, or a close. The corresponding functions are detailed in the next paragraph.
+The main flags that can be sent appear in figure~\ref{Fig:statemachine}. They are sent
+alongside TCP segments, in the header.
+\flag{SYN} is sent to synchronize two TCPs when they want to initiate a connection.
+The \flag{ACK} flag confirms the reception of a previous message. \flag{FIN} indicates that
+the TCP server wants to close the connection. Finally, a flag named \flag{RST} can be sent or
+received at any instant and it forces the connection to be reset.
+If so, Both TCPs jump to the state \state{CLOSED}.
+The transition are represented in figure~\ref{Fig:statemachine} in the format
+$\frac{\ \ x\ \ }{\ \ y\ \ }$ where $x$ is the event that triggers the transition and $y$ is
+the response given to fulfil the transition.
+For instance, if the TCP is in the state \state{SYN\_SENT} and it receives a segment that
+contains the flag \flag{SYN} and the flag \flag{ACK} that acknowledges the previous \flag{SYN}
+sent, then the TCP responds by sending a flag \flag{ACK} to acknowledge the segment received
+and it proceeds to state \state{ESTABLISHED}.
+
+\subsubsection{User functions}
+
+In order to use the TCP protocol in programs, the norm defines user functions that can be invoked
+by the user in its code. This is the well-known \emph{socket API}, available on almost all operating
+systems, and known on POSIX systems as the \emph{Berkeley sockets}.
+This API specifies the already mentioned \texttt{open} and
+\texttt{close} functions, as well as functions \texttt{send} and \texttt{receive} to send
+or receive data. The behavior of each function is defined depending on the state of the
+TCP when the function is called.
+It appears in the state machine that most of the transitions are due to the reception of
+messages.
+
+\subsubsection{Tasks}
+
+A model with multiple tasks is adopted in the norm.
+Each task has a dedicated job. On task is dedicated to the sending and the reception of messages.
+Another is dedicated to the user functions and the last to timers. Timers can resend data or close
+a connection if the remote server does not respond.
+The first task performs most of the transitions in the state machine.
+
+\subsection{CycloneTCP library organization}
 
@@
     The CycloneTCP library is TCP/IP stack that implements network protocols.
-    Figure \ref{Fig:TcpStack} gives an overview of the different protocols supported. TCP and UDP protocols have a central place
-    in the stack. An API is provided to use those network protocols: the Socket API. It is a common way to use transport protocols,
-    and the application protocols in the stack, such as HTTP or FTP use it.
+    Figure \ref{Fig:TcpStack} gives an overview of the different protocols supported.
+    TCP and UDP protocols have a central place in the stack as well as the socket API.
+    This API is a common way to use transport protocols, but they can also be used
+    hidden the high level application protocols in the stack, such as HTTP or FTP.
 
     The stack is particularly well adapted to embedded platforms and microcontrolers. As a result, the code is written
-    to compile on those platforms and to be used with different real time OS, like for intance FreeRTOS.
+    to compile on those platforms and to be used with different real time OS, like, for instance, FreeRTOS.
     Hence, a model based on multiple tasks has been adopted to implement the TCP protocol as defined in the norm, with,
@@
 
-    To reduce the size of the work, only few functions have been totaly translated
+    To reduce the size of the work, only few functions have been totally translated
     in SPARK. Then, some C functions are reused in SPARK, and some SPARK functions
@@
     in C. Symmetrically, the attribute \lstinline{Export} allows the subprogram
-    to be reused in C. An \lstinline{External_Name} might be given for the
+    to be reused in C. An \lstinline{External_Name} must be given for the
     programmers and the linker to locate the right symbol at the liking time.
@@
     The types in SPARK can be more expressive than the ones used in C. However
-    they must have the same representation in memory to be used as an argument
-    of a function. The attribute \lstinline{Size} can be added to SPARK types
-    to keep them coherent with the C types. It specifies to the compiler the
-    size in bits requiered to store an object of this type.
+    they must have the same representation in memory in particular when an
+    object is shared between SPARK and C code.
+    The attribute \lstinline{Size} can be added to SPARK types to keep them
+    coherent with the corresponding C types. It specifies to the compiler the
+    size in bits required to store an object of this type.
 
@@
 
-    In CycloneTCP library, pointers are often manupulated at top-level and
+    In CycloneTCP library, pointers are often manipulated at top-level and
     transmitted to C functions.  They are also used as fields in structures,
@@
     interesting to prove the correctness of the states transitions. Then, it is
-    necessary to create a model of the socket to describe the revelant field of
+    necessary to create a model of the socket to describe the relevant field of
     the socket to keep. Two models have been created, the first \lstinline{Model}
@@
     The post-conditions use these models in such way
-    \begin{lstlisting}[language=Ada]
-    Model(Sock) = (Model(Sock)'Old with delta
-                        State => TCP_STATE_CLOSED)
+    \begin{lstlisting}[style=Spark]
+    Model(Sock) = (Model(Sock)'Old with delta State => TCP_STATE_CLOSED)
     \end{lstlisting}
@@
     However, to be integrated with the existing library, it has been necessary to reuse existing C code
-    as an untrested black box, for which we cannot have strong guarantees. Figure \ref{Fig:TcpStack} gives an overview of
+    as an untrusted black box, for which we cannot have strong guarantees. Figure \ref{Fig:TcpStack} gives an overview of
     what have been translated and what have been reused. The box for TCP is hatched because only the user level functions
@@
 \begin{subfigure}{\textwidth}
-\begin{lstlisting}[language=C, frame=bottomline]
+\begin{lstlisting}[style=C, frame=bottomline]
 Socket *socketOpen(uint_t type, uint_t protocol);
-error_t socketConnect(Socket *socket, IpAddr *remoteIpAddr,
-                      uint16_t remotePort);
+error_t socketConnect(Socket *socket, IpAddr *remoteIpAddr, uint16_t remotePort);
 \end{lstlisting}
@@
 \begin{subfigure}{\textwidth}
-\begin{lstlisting}[language=Ada, frame=bottomline]
+\begin{lstlisting}[style=Spark, frame=bottomline]
 type Socket_Struct is private;
@@
 type Socket_Type is (SOCKET_TYPE_STREAM, SOCKET_TYPE_DGRAM);
-type Socket_Proto is (SOCKET_IP_PROTO_TCP, SOCKET_IP_PROTO_UDP);
+type Socket_Prot is (SOCKET_IP_PROTO_TCP,SOCKET_IP_PROTO_UDP);
 
@@
      S_Type     : in     Socket_Type;
-     S_Protocol : in     Socket_Protocol)
+     S_Protocol : in     Socket_Prot)
 with
@@
     that is also a precondition for $f_2$. Then, for the precondition of $f_2$ to be true, the function $f_1$ must have been called first.
-    Such postcondition is presented on figure \ref{Fig:socketInterface:SPARK} for the function \lstinline{Socket_Connect}. After its
-    execution, the field \lstinline{S_Remote_Ip_Addr} of the socket \lstinline{Sock}. Then, since all the functions to send or receive
+    Such postcondition is presented on figure \ref{Fig:socketInterface:SPARK} for the function \lstinline[style=Spark]{Socket_Connect}. After its
+    execution, the field \lstinline[style=Spark]{S_Remote_Ip_Addr} of the socket \lstinline[style=Spark]{Sock}. Then, since all the functions to send or receive
     data and to close the connection require a connection established, these functions have a precondition that contains a condition
     that states that the condition is established, which can be written as
-    \begin{lstlisting}
+    \begin{lstlisting}[style=Spark]
     Pre => Is_Ip_Initialized(Sock.S_Remote_Ip_Addr);
@@
 
-    The same mechanism is used with the functions \lstinline{Socket_Open} that create a valid socket when the function
-    succeeds and \lstinline{Socket_Connect} that require a valid not null socket.
+    The same mechanism is used with the functions \lstinline[style=Spark]{Socket_Open} that creates a valid socket when the function
+    succeeds and \lstinline[style=Spark]{Socket_Connect} that requires a valid not null socket.
 
@@
     In the previous paragraph we explained that the functions that send or receive data check in their precondition if the connection
-    has been initialized or not, by checking the field \lstinline{S_Remote_Ip_Addr}, and we explained that this value is initialized
-    by \lstinline{Socket_Connect}. The presence in the postcondition of \lstinline{Socket_Connect} or if statement
-    \lstinline{(if Error = NO_ERROR then ...)} makes that \lstinline{S_Remote_Ip_Addr} is initialized only when no error occurs during
+    has been initialized or not, by checking the field \lstinline[style=Spark]{S_Remote_Ip_Addr}, and we explained that this value is initialized
+    by \lstinline[style=Spark]{Socket_Connect}. The presence of the if statement \lstinline[style=Spark]{(if Error = NO_ERROR then ...)} in the postcondition
+    of \lstinline[style=Spark]{Socket_Connect}  makes that \lstinline[style=Spark]{S_Remote_Ip_Addr} is initialized only when no error is reported during
     the execution of the function. Thanks to the mechanism exposed in the previous section the next function can only be called
@@
 
-    Let \lstinline{Socket_Send} be the function to send data. The following example of code is incorrect because if \lstinline{Error} is
+    Let \lstinline[style=Spark]{Socket_Send} be the function to send data. The following example of code is incorrect because if \lstinline{Error} is
     different to \lstinline{NO_ERROR}, nothing ensures that \lstinline{S_Remote_Ip_Addr} is initialized.
-    \begin{lstlisting}[language=Ada]
+    \begin{lstlisting}[style=Spark]
     Socket_Connect(Sock, Remote_Ip_Addr, Remote_Port, Error);
@@
     \end{lstlisting}
-    In comparison this second example of code produces no error because when \lstinline{Socket_Send} is called, GNATprove knows that
-    \lstinline{ERROR} is equal to \lstinline{NO_ERROR} and then \lstinline{S_Remote_Ip_Addr} is initialized.
-    \begin{lstlisting}[language=Ada]
+    In comparison this second example of code does not produce any check message because when \lstinline[style=Spark]{Socket_Send} is called, GNATprove knows that
+    \lstinline[style=Spark]{ERROR} is equal to \lstinline[style=Spark]{NO_ERROR} and then \lstinline[style=Spark]{S_Remote_Ip_Addr} is initialized.
+    \begin{lstlisting}[style=Spark]
     Socket_Connect(Sock, Remote_Ip_Addr, Remote_Port, Error);
@@
     The same mechanism can be observed on figure \ref{Fig:socketInterface:SPARK}. The user has to test if the socket returned by
-    \lstinline{Socket_Open} is null or not before calling \lstinline{Socket_Connect}, otherwise GNATprove will emit a message.
+    \lstinline[style=Spark]{Socket_Open} is null or not before calling \lstinline[style=Spark]{Socket_Connect}, otherwise GNATprove will emit a message.
 
@@
     implementation that can differ on some points.
-    After this work, we then have a postcondition for \lstinline{tcpProcessSegment}, that can be reported in the SPARK code for further proofs, of the form:
+    After this work, we get a postcondition for \lstinline{tcpProcessSegment}, that can be reported in the SPARK code for further proofs, of the form:
     \begin{lstlisting}[language=Ada,mathescape=true]
@@
     When it is called, it checks if the event is already true, thanks to a call to the function \lstinline{Tcp_Update_Events}.
-    If it is not the case, the mutex that enforces the mutual exclusion of access to the socket is unlocked.
-    Messages can be received until the event waited becomes true.
+    This function returns all the events that the socket satisfies when it is called.
+    If the event is not satisfied, the mutex that enforces the mutual exclusion of access to the socket is unlocked.
+    Several messages can be received until the event waited becomes true.
     Once a reception makes true the event, the function \lstinline{Tcp_Wait_For_Events} locks again the mutex and returns.
     A possible algorithm to simulate the reception of segment and compute the least set of states that can result of the wait can be done
-    by the algorithm \ref{algo:waitForEvents}.
+    by the algorithm~\ref{algo:waitForEvents}.
+
+    The parameter \textit{Event\_Mask} represents the event that we are waiting for. A loop of three steps simulates the
+    reception of three messages with a call to the function \lstinline[style=Spark]{Tcp_Process_One_Segment}. The new events
+    satisfied are computed, and if the searched event is reached, the socket is returned.
+    Then, it is possible for SPARK to compute all the relations that link the state of the input socket with the states
+    of the returned socket, depending on the event waited.
+    The loop is limited to three steps because all the paths in the automaton that do not require a user action have
+    a length of at most three.
 
@@
         \SetKwProg{Fn}{function}{}{end}
-        \Fn{\ProcessSegm{Socket, Event, Event\_Mask}}{
+        \Fn{\ProcessSegm{Socket, Event\_Mask}}{
             $S_{last} := \text{\textit{Socket}}$\;
@@
         }
-        \caption{Function to compute the possible states after a wait for a particular event.}
+        \caption{Function to compute the possible states after a wait for an event.}
         \label{algo:waitForEvents}
@@
 \label{sec:results}
+
+    The work done on the TCP implementation has shown the powerful
+    of SPARK in finding bugs. One of the bugs found, a memory leak,
+    relies on the use of the pointers.
 
@@
 
-    However weeknesses exist our implementation, in particular because all the underlying layers are still written in C.
+    However weeknesses exist in our implementation, in particular because all the underlying layers are still written in C.
     The principal function of those layers is to format packet before they are sent, or parse incoming packets, check

### 2021-03-29 paper/paper.tex
Denrière partie

@@
 \usepackage[colorlinks]{hyperref}
-
-\ifLuaTeX
-    \usepackage{fontspec}
-    \IfFontExistsTF{Fira Code}
-    {\setmonofont[Scale=MatchLowercase]{Fira Code}}
-    {\IfFontExistsTF{DejaVu Sans Mono}{
-        \setmonofont{DejaVu Sans Mono}[Scale=MatchLowercase]
-    }{}}
-\fi
 
@@
 }
+
+\let\Spark\lstinline
 
@@
 Socket *socketOpen(uint_t type, uint_t protocol);
-error_t socketConnect(Socket *socket, IpAddr *remoteIpAddr, uint16_t remotePort);
+error_t socketConnect(Socket *socket, IpAddr *remoteIpAddr,
+                      uint16_t remotePort);
 \end{lstlisting}
@@
 with
-    Post => (if Sock /= null then
+  Post =>
+    (if Sock /= null then
       Sock.S_Type = S_Type and then
@@
     that is also a precondition for $f_2$. Then, for the precondition of $f_2$ to be true, the function $f_1$ must have been called first.
-    Such postcondition is presented on figure \ref{Fig:socketInterface:SPARK} for the function \lstinline[style=Spark]{Socket_Connect}. After its
-    execution, the field \lstinline[style=Spark]{S_Remote_Ip_Addr} of the socket \lstinline[style=Spark]{Sock}. Then, since all the functions to send or receive
+    Such postcondition is presented on figure \ref{Fig:socketInterface:SPARK} for the function \Spark{Socket_Connect}. After its
+    execution, the field \Spark{S_Remote_Ip_Addr} of the socket \Spark{Sock}. Then, since all the functions to send or receive
     data and to close the connection require a connection established, these functions have a precondition that contains a condition
@@
     \end{lstlisting}
-    and because the function \lstinline{Socket_Connect} is the only one that can initialize this value, the order is respected
+    and because the function \Spark{Socket_Connect} is the only one that can initialize this value, the order is respected
     if and only if the code is valid.
 
-    The same mechanism is used with the functions \lstinline[style=Spark]{Socket_Open} that creates a valid socket when the function
-    succeeds and \lstinline[style=Spark]{Socket_Connect} that requires a valid not null socket.
+    The same mechanism is used with the functions \Spark{Socket_Open} that creates a valid socket when the function
+    succeeds and \Spark{Socket_Connect} that requires a valid not null socket.
 
@@
     In the previous paragraph we explained that the functions that send or receive data check in their precondition if the connection
-    has been initialized or not, by checking the field \lstinline[style=Spark]{S_Remote_Ip_Addr}, and we explained that this value is initialized
-    by \lstinline[style=Spark]{Socket_Connect}. The presence of the if statement \lstinline[style=Spark]{(if Error = NO_ERROR then ...)} in the postcondition
-    of \lstinline[style=Spark]{Socket_Connect}  makes that \lstinline[style=Spark]{S_Remote_Ip_Addr} is initialized only when no error is reported during
+    has been initialized or not, by checking the field \Spark{S_Remote_Ip_Addr}, and we explained that this value is initialized
+    by \Spark{Socket_Connect}. The presence of the if statement \Spark{(if Error = NO_ERROR then ...)} in the postcondition
+    of \Spark{Socket_Connect}  makes that \Spark{S_Remote_Ip_Addr} is initialized only when no error is reported during
     the execution of the function. Thanks to the mechanism exposed in the previous section the next function can only be called
@@
 
-    Let \lstinline[style=Spark]{Socket_Send} be the function to send data. The following example of code is incorrect because if \lstinline{Error} is
+    Here, \Spark{Socket_Send} refers to the function to send data. The following example of code is incorrect because if \lstinline{Error} is
     different to \lstinline{NO_ERROR}, nothing ensures that \lstinline{S_Remote_Ip_Addr} is initialized.
@@
     \end{lstlisting}
-    In comparison this second example of code does not produce any check message because when \lstinline[style=Spark]{Socket_Send} is called, GNATprove knows that
-    \lstinline[style=Spark]{ERROR} is equal to \lstinline[style=Spark]{NO_ERROR} and then \lstinline[style=Spark]{S_Remote_Ip_Addr} is initialized.
+    In comparison this second example of code does not produce any check message because when \Spark{Socket_Send} is called, GNATprove knows that
+    \Spark{ERROR} is equal to \Spark{NO_ERROR} and then \Spark{S_Remote_Ip_Addr} is initialized.
     \begin{lstlisting}[style=Spark]
@@
     The same mechanism can be observed on figure \ref{Fig:socketInterface:SPARK}. The user has to test if the socket returned by
-    \lstinline[style=Spark]{Socket_Open} is null or not before calling \lstinline[style=Spark]{Socket_Connect}, otherwise GNATprove will emit a message.
+    \Spark{Socket_Open} is null or not before calling \Spark{Socket_Connect}, otherwise GNATprove will emit a message.
 
@@
     The parameter \textit{Event\_Mask} represents the event that we are waiting for. A loop of three steps simulates the
-    reception of three messages with a call to the function \lstinline[style=Spark]{Tcp_Process_One_Segment}. The new events
+    reception of three messages with a call to the function \Spark{Tcp_Process_One_Segment}. The new events
     satisfied are computed, and if the searched event is reached, the socket is returned.
@@
 
-    The work done on the TCP implementation has shown the powerful
+    The work done on the TCP implementation has shown the strength
     of SPARK in finding bugs. One of the bugs found, a memory leak,
     relies on the use of the pointers.
+
+\subsubsection{Concurrency bug}
+
+    The most significant problem found in the TCP implementation relies on the concurrency.
+    Concurrency is one of the biggest source of bug when writing code. It is hard to understand
+    all the possible interactions between different threads which leads to many bugs that can be
+    difficult to reproduce and to debug. The use of SPARK helps us to find one.
+
+    %% TODO : Show the relevant code snippet here
+
+\subsubsection{Memory leak}
+
+    A memory leak has been found thanks by SPARK and the recent work of Dross and Kanig~\cite{dross2020recursive}.
+    They added a pointer support to SPARK inspired by Rust. At any program point, every memory cell must be owned by a pointer.
+    If a memory cell has not be freed and is not owned anymore by any pointer, there is a memory leaks.
+    This mechanism of ownership has helped in finding one memory leak in the original C code that has been fixed in the
+    SPARK version of the code.
+
+%% Ici on peut parler d'une ou deux typos trouvées dans le code. Mais plutôt liées à une relecture de code et pas vraiment dues à SPARK
+
+\subsubsection{Metrics of the new implementation}
+
+\subsubsection{Tests to validate the new implementation}
 

### 2021-04-13 paper/paper.tex
Suite du papier

@@
     classoffset = 2,
-    morekeywords = {Socket_Open, Socket_Connect, Socket_Send},
+    morekeywords = {Socket_Open, Socket_Connect, Socket_Send, Tcp_Send, Tcp_Wait_For_Events, Tcp_Send_Segment, Tcp_Change_State},
     classoffset = 7,
@@
     The most significant problem found in the TCP implementation relies on the concurrency.
-    Concurrency is one of the biggest source of bug when writing code. It is hard to understand
+    Concurrency is one of the biggest source of bugs when writing code. It is hard to understand
     all the possible interactions between different threads which leads to many bugs that can be
     difficult to reproduce and to debug. The use of SPARK helps us to find one.
+\begin{figure}[t]
+\begin{lstlisting}[style=Spark, frame=bottomline, basicstyle=\ttfamily\footnotesize,
+                    numbers=left, numberstyle=\tiny, escapechar=\%]
+case Sock.State is
+  when TCP_STATE_SYN_RECEIVED
+      | TCP_STATE_ESTABLISHED =>
+  -- Flush the send buffer
+  Tcp_Send (Sock, Buf, Ignore_Written,      % \label{bugProg:tcpSend} %
+            SOCKET_FLAG_NO_DELAY, Error);
+  if Error /= NO_ERROR then return; end if;
+  -- Make sure all the data has been sent out
+  Tcp_Wait_For_Events   % \label{bugProg:tcpWaitForEvents} %
+    (Sock       => Sock,
+    Event_Mask => SOCKET_EVENT_TX_DONE,
+    Timeout    => Sock.S_Timeout,
+    Event      => Event);
+  if Event /= SOCKET_EVENT_TX_DONE then
+    Error := ERROR_TIMEOUT; return;
+  end if;
+  -- Send a FIN segment
+  Tcp_Send_Segment      % \label{bugProg:tcpWaitForEventsPost} %
+    (Sock         => Sock,
+    Flags        => TCP_FLAG_FIN or TCP_FLAG_ACK,
+    Seq_Num      => Sock.sndNxt,
+    Ack_Num      => Sock.rcvNxt,
+    Length       => 0,
+    Add_To_Queue => True,
+    Error        => Error);
+  if Error /= NO_ERROR then return; end if;
+  -- Switch to the FIN-WAIT-1 state
+  Tcp_Change_State (Sock, TCP_STATE_FIN_WAIT_1); % \label{bugProg:tcpChangeState} %
+\end{lstlisting}
+\caption{Reproduction of the code where the concurrency bug has been found.}
+\label{fig:concurrencyBug}
+\end{figure}
 
-    %% TODO : Show the relevant code snippet here
+    A snippet in figure~\ref{fig:concurrencyBug} shows a bug of this kind found in the initial code.
+    At line~\ref{bugProg:tcpSend}, the buffer is flushed, if no error happened during this stage,
+    the state of the socket is either \state{ESTABLISHED} or \state{CLOSE\_WAIT} because the mutex
+    that protected the socket has been released and a transition has possibly be done. The call to
+    \Spark{Tcp_Wait_For_Events} at line~\ref{bugProg:tcpWaitForEvents}, produces, if succeeds, the
+    state of the socket to be in the set \state{ESTABLISHED}, \state{CLOSE\_WAIT} or \state{CLOSED}.
+    After, those two calls, \Spark{Tcp_Send_Segment} is called to initiate a closing of the connection
+    by sending a \flag{FIN} flag at line~\ref{bugProg:tcpWaitForEventsPost}.
+
+    The programmer who wrote this code forgot that the state of the socket has changed during the last
+    called and when the \flag{FIN} flag is sent, he though that the state was still \state{SYN\_RECEIVED}
+    or \state{ESTABLISHED}. The transition that looks the most appropriate for him is to jump in the state
+    \state{FIN\_WAIT\_1}, as done at line~\ref{bugProg:tcpChangeState}. Unfortunately, this change of
+    state makes the transitions \state{CLOSE\_WAIT} $\rightarrow$ \state{FIN\_WAIT\_1} and
+    \state{CLOSED} $\rightarrow$ \state{FIN\_WAIT\_1} possible. Those two transitions do not
+    appear in the state machine in figure~\ref{Fig:statemachine} because they are forbidden by the
+    norm.
+
+    As the precondition given to the function \Spark{Tcp_Change_State} lists all the authorized transition,
+    SPARK warmed up that an incorrect transition was possible in the current code. The code has been reported to
+    the maintainer of the TCP/IP stack and fixed.
 
@@
 %% Ici on peut parler d'une ou deux typos trouvées dans le code. Mais plutôt liées à une relecture de code et pas vraiment dues à SPARK
+\subsubsection{Tests to validate the new implementation}
+
+    In addition to the guarantees given by SPARK, we have carried out some tests by enabling the assertions at the runtime.
+    Doing so strengthen our confidence in the assertions written in SPARK, and in particular it provides more sureness
+    that the assertions written are not broken by C code. The assertion are checked during the execution time
+    and if one of them is violated, a runtime error is raised and it is possible to investigate.
+
+    In order to do that, we have written some HTTP drivers with different scenarios. During the different runs,
+    no assertion has been violated.
 
@@
 
-\subsubsection{Tests to validate the new implementation}
-
-    \begin{itemize}
-        \item Bug de concurrence -> SPARK
-        \item Memory leak -> SPARK
-        \item Problèmes trouvés.
-        \item Résultats sur la nouvelle implémentation (performance, taille du code).
-        \item Activation des préconditions pour voir si elles sont toutes valides.
-    \end{itemize}
+    \textcolor{red}{TODO : Quoi mettre ?}
 

### 2021-05-12 SecDev21/Automaton.tex
First draft of the paper

--- 2020-10-22 paper/Automaton.tex
@@
 \begin{figure}[t]
-{\centering\scalebox{.7}{
+    {\centering\scalebox{.55}{
     \begin{tikzpicture}[>=latex]
-    \def\tfrac#1#2{\ensuremath{\displaystyle\frac{\text{\footnotesize#1}}{\text{\footnotesize#2}}}}
+        \def\tfrac#1#2{\ensuremath{\displaystyle\frac{\text{\small#1}}{\text{\small#2}}}}
         %
@@
         \node[state, name=listen, below of=closedStart] {LISTEN};
-    \node[state, name=synSent, below of=listen, right of=listen, xshift=8em] {SYN\_SENT};
-    \node[state, name=synRcvd, below of=listen, left of=listen, xshift=-8em] {SYN\_RCVD};
+        \node[state, name=synSent, below of=listen, right of=listen, xshift=8em] {SYN-SENT};
+        \node[state, name=synRcvd, below of=listen, left of=listen, xshift=-8em] {SYN-RCVD};
         \node[state, name=established, below of=listen, node distance=14em] {ESTABLISHED};
-    \node[state, name=finWait1, below of=established, left of=established, node distance=7em, xshift=-9em] {FIN\_WAIT\_1};
-    \node[state, name=finWait2, below of=finWait1] {FIN\_WAIT\_2};
-    \node[state, name=closeWait, below of=established, right of=established, node distance=7em, xshift=9em] {CLOSE\_WAIT};
+        \node[state, name=finWait1, below of=established, left of=established, node distance=7em, xshift=-9em] {FIN-WAIT-1};
+        \node[state, name=finWait2, below of=finWait1] {FIN-WAIT-2};
+        \node[state, name=closeWait, below of=established, right of=established, node distance=7em, xshift=9em] {CLOSE-WAIT};
         \node[state, name=closing, below of=established, node distance=14em] {CLOSING};
-    \node[state, name=lastAck, below of=closeWait] {LAST\_ACK};
-    \node[state, name=timeWait, below of=closing] {TIME\_WAIT};
+        \node[state, name=lastAck, below of=closeWait] {LAST-ACK};
+        \node[state, name=timeWait, below of=closing] {TIME-WAIT};
 

### 2021-05-12 SecDev21/OryxStack.tex
First draft of the paper

\newlength\hatchdistance
\hatchdistance=10pt
\pgfdeclarepatternformonly[\hatchdistance,2pt]{north east hatch}% name
    {\pgfqpoint{-1pt}{-1pt}}% below left
    {\pgfqpoint{\hatchdistance}{\hatchdistance}}% above right
    {\pgfpoint{\hatchdistance-1pt}{\hatchdistance-1pt}}%
    {
        \pgfsetcolor{orange!40}
        \pgfsetlinewidth{2pt}
        \pgfpathmoveto{\pgfqpoint{0pt}{0pt}}
        \pgfpathlineto{\pgfqpoint{\hatchdistance}{\hatchdistance}}
        \pgfusepath{stroke}
    }
\begin{figure}[t]
    \centering\begin{tikzpicture}[x=.12\linewidth,y=.45cm,
        application/.style={font=\sffamily\tiny,
                            draw,
                            minimum width=.115\linewidth,
                            minimum height=.4cm},
        session/.style={font=\sffamily\tiny,
                        draw,
                        minimum width=.715\linewidth,
                        minimum height=.4cm},
        transport/.style={font=\sffamily\tiny,
                          draw,
                          minimum width=.295\linewidth,
                          minimum height=.4cm},
        network/.style={font=\sffamily\tiny,
                        draw,
                        minimum width=.355\linewidth,
                        minimum height=.4cm},
        datalink/.style={font=\sffamily\tiny,
                         draw,
                         minimum width=.139\linewidth,
                         minimum height=.4cm}
    ]
        \node[application] at (0,10) {HTTP};
        \node[application] at (1,10) {HTTP/2};
        \node[application] at (2,10) {MQTT};
        \node[application] at (3,10) {MQTT-SN};
        \node[application] at (4,10) {CoAP};
        \node[application] (CoinA) at (5,10) {FTP};
        \node[application] at (0,9)  {SMTP};
        \node[application] at (1,9)  {SNTP};
        \node[application] at (2,9)  {DNS};
        \node[application] at (3,9)  {NetBIOS};
        \node[application] at (4,9)  {SNMPv3};
        \node[application] at (5,9)  {TFTP};
        \node[application, minimum width=.235\linewidth] at (.5,8) {WebSocket};
        \node[application] at (2,8)  {mDNS};
        \node[application] at (3,8)  {DNS-SD};
        \node[application] at (4,8)  {DHCP};
        \node[application] (CoinB) at (5,8)  {DHCPv6};
        \node[session, fill=green!70!yellow!20] (Socket) at (2.5, 7) {Socket};
        \node[transport, preaction={fill, green!70!yellow!20}, pattern=north east hatch, pattern color=orange!40] at (.75,6) {TCP};
        \node[transport, fill=orange!40] at (3.25,6) {UDP};
        \node[application, fill=orange!40] (RAW) at (5,6) {RAW};
        \node[network, fill=orange!40] at (1, 5) {IPv4};
        \node[network, fill=orange!40] (IPv6) at (4, 5) {IPv6};
        \node[application, fill=orange!40] at (0,4) {ARP};
        \node[application, fill=orange!40] at (1,4) {Auto-IP};
        \node[application, fill=orange!40] at (2,4) {NDP};
        \node[application, fill=orange!40] at (3,4) {SLAAC};
        \node[application, fill=orange!40] (ICMP) at (4,4) {ICMP};
        \node[application, fill=orange!40] (MLDv1) at (5,4) {MLDv1};
        \node[datalink, fill=orange!40] at (0.10, 3) {Ethernet};
        \node[datalink, fill=orange!40] at (1.3, 3) {Wi-Fi};
        \node[datalink, fill=orange!40] at (2.5, 3) {PPP};
        \node[datalink, fill=orange!40] at (3.7, 3) {USB/RNDIS};
        \node[datalink, fill=orange!40] (G3) at (4.9, 3) {G3-PLC};

        \draw[decorate, decoration={brace}, line width=.5pt]
            ([xshift=2pt]CoinA.north east) -- ([xshift=2pt]CoinB.south east)
            node [midway, anchor=west, font=\sffamily\tiny] {Application layer};
        \draw[decorate, decoration={brace}, line width=.5pt]
            ([xshift=2pt]Socket.north east) -- ([xshift=2pt]Socket.south east)
            node [midway, anchor=west, font=\sffamily\tiny] {Session layer};
        \draw[decorate, decoration={brace}, line width=.5pt]
            ([xshift=2pt]RAW.north east) -- ([xshift=2pt]RAW.south east)
            node [midway, anchor=west, font=\sffamily\tiny] {Transport layer};
        \draw[decorate, decoration={brace}, line width=.5pt]
            ([xshift=2pt]IPv6.north east) -- ([xshift=2pt]MLDv1.south east)
            node [midway, anchor=west, font=\sffamily\tiny] {Network layer};
        \draw[decorate, decoration={brace}, line width=.5pt]
            ([xshift=2pt]G3.north east) -- ([xshift=2pt]G3.south east)
            node [midway, anchor=west, font=\sffamily\tiny] {Data Link layer};
    \end{tikzpicture}

    \caption{Overview of the TCP/IP CycloneTCP stack in the OSI model. Green parts represent
             what have been translated in SPARK and in orange the underlying untrusted part written in C.}
    \label{Fig:TcpStack}
\end{figure}

### 2021-05-12 SecDev21/paper.tex
First draft of the paper

\documentclass[conference]{IEEEtran}

% Class modification
\IEEEilabelindent=0pt

\usepackage[utf8]{inputenc}
\usepackage[x11names]{xcolor}

\usepackage[np]{numprint}

\definecolor{linkcolor}{RGB}{36,113,163}
\definecolor{citecolor}{RGB}{17,122,101}
\definecolor{urlcolor}{RGB}{148,39,38}

\usepackage[
    colorlinks,
    linkcolor=linkcolor,
    citecolor=citecolor,
    urlcolor=urlcolor
]{hyperref}

\usepackage{algorithm}
\usepackage{algpseudocode}

\usepackage{amsmath, amsfonts}

\usepackage{lipsum}

\usepackage{tikz}
\usetikzlibrary{decorations.pathreplacing, patterns}
\usetikzlibrary{calc, backgrounds}

\usepackage{xspace}

\usepackage{listings}

\lstset{
    basicstyle=\ttfamily,
    keywordstyle={[1]\color{Blue4}\bfseries}, % keywords style
    % keywordstyle={[2]\color{Pink4}}, % type style
    % keywordstyle={[3]\color{OrangeRed3}}, % function style
    % literate =
    %     {=>}{{=>}}2
    %     {/=}{{/=}}2,
    commentstyle=\itshape\color{gray}
}

\def\spark#1{\lstinline[language=Ada]{#1}}

\def\todo#1{\textcolor{red}{[#1]}}

\def\state#1{\textsf{\MakeUppercase{#1}}\xspace}
\def\sclosed{\state{closed}}
\def\ssynsent{\state{syn-sent}}
\def\ssynrcv{\state{syn-rcvd}}
\def\slisten{\state{listen}}
\def\sestab{\state{established}}
\def\sfwone{\state{fin-wait-1}}
\def\sfwtwo{\state{fin-wait-2}}
\def\sclosing{\state{closing}}
\def\sclosew{\state{close-wait}}
\def\slastack{\state{last-ack}}
\def\stimewait{\state{time-wait}}

\def\flag#1{\textsf{#1}\xspace}
\def\syn{\flag{SYN}}
\def\ack{\flag{ACK}}
\def\rst{\flag{RST}}
\def\fin{\flag{FIN}}

\let\code\texttt

\begin{document}

\title{Verification of a TCP/IP stack}

\author{%
\IEEEauthorblockN{Guillaume Cluzel}
\IEEEauthorblockA{\textit{Adacore \& ENS de Lyon}}
\and
\IEEEauthorblockN{Kyriakos Georgious}
\IEEEauthorblockA{\textit{Adacore \& University of Bristol}}
\and
\IEEEauthorblockN{Yannick Moy}
\IEEEauthorblockA{\textit{Adacore}}
\and
\IEEEauthorblockN{Clément Zeller}
\IEEEauthorblockA{\textit{Oryx Embedded}}
}

\maketitle

\begin{abstract}
\todo{Write the abstract.}
\end{abstract}

\begin{IEEEkeywords}
Network protocols,
deductive verification.
\end{IEEEkeywords}

\section{Introduction}

The usage of IoT devices has been put more within everyone's reach. Their
vulnerability is a major cause for concern, and as their number
increases, the number of attacks they suffer increases, putting at risk
the security of the owner's data collected by the object. \todo{Try to find a
relevant ref. I have found examples about the possibility to leak personnal data
on Google home/Alexa}.

The core of the IoT devices is built on top of the TCP protocol~\cite{rfc793} as
it is the most widely protocol used on the Internet, as the underlying protocol
for application layers such as HTTP, FTP or TLS. TCP is based on the underlying
IP protocol which is itself based on a link layer such as Ethernet. This
stacking of protocols is referred to as the TCP/IP stack.

Given the importance of the TCP protocol in network communications, some
researchers have attempted to formalize the protocol to either prove key
properties~\cite{smith1996formal} or even to fully prove the correctness of an
implementation against a complete specification~\cite{ridge2008rigorous}.
This latter implementation was performed in the HOL proof assistant, and while
the authors note that it would be possible in theory to extract an
implementation in Haskell from this work, this has not been adopted in practice.
The formalization of TCP specification and verification of a TCP implementation
face two difficulties:
RFC 793 is written in English, leaving many parts underspecified; the TCP
protocol is inherently concurrent which makes it complex to specify and verify.
As a result, there is currently no formally verified implementation for TCP that
is usable in industry.

Other communication protocols have been formalized in recent years, leading to a
formally verified implementation. The most prominent project in this area is the
Everest project~\cite{bhargavan2017everest}, which led to a formally verified
implementation in F* of TLS called miTLS~\cite{bhargavan2013implementing}. The
approach of the Everest project is to implement TLS in F*, a programming
language targeted at formal program verification.
Note that miTLS still relies on an untrusted implementation of TCP, and thus
could benefit from the results of our work.
This approach is often difficult to adopt in industry, as codebases are evolved
incrementally rather than replaced, both for economical and practical reasons.

Our approach in this work has been to incrementally evolve an existing TCP/IP
stack in C by replacing parts of the code with formally verified code in SPARK.
The stack chosen, CycloneTCP\footnote{\url{https://www.oryx-embedded.com/products/CycloneTCP.html}},
a mature TCP/IP stack targeted at embedded processors with no Operating System.
We present the TCP protocol in section~\ref{sec:TCP} and then briefly the
choosen stack, the Oryx Embedded one, in section~\ref{sec:stack}
before presenting the specification techniques used in section~\ref{sec:spec}
and then in sections~\ref{sec:verif} and~\ref{sec:API} we explain how we
formally implemented in SPARK the TCP user functions and how we hardened the
user's API. Finally we end in section~\ref{sec:results} by showing the results
of our work.

\section{TCP protocol}
\label{sec:TCP}

Initially described by the ``fathers of the Internet'' Vincent Cerf and Bob
Kahn, the TCP protocol was standardized in 1981 by RFC 793~\cite{rfc793}.
Its sturdiness makes it still one of the most widely used protocols today with
UDP.

In the OSI networking communication model, the TCP protocol is central and
located in the \emph{Transport layer}. It itself uses the \emph{Network layer}
to work.
The security of the \emph{Application} protocols depends on the security of
the \emph{Session layer} and consequently on the security of the TCP and UDP
protocols. Figure~\ref{Fig:TcpStack} gives an overview of how protocols are
stacked.

The purpose of this section is to give enough context to enable the reader to
follow the work conducted in this article.
The TCP protocol is connection-oriented in the sense that a connection requires
to be established between the two processes that exchange data.
The protocol is governed by a state machine described in
figure~\ref{Fig:statemachine}.
The implementation is rather free and lots of choices are left to the
programmer. However, the implementation must ensure some properties to be able
to communicate with other TCPs.

\input{Automaton.tex}

\subsection{States}

Generally, every TCP communication session will go through the following three
phases (if no error occurs): 1)~opening the connection 2)~sending and receiving
the data and 3)~closing the connection.

The behavior of the three-phase communication can be described by the state
machine.
Every group of states corresponds to a one phase. More precisely \sclosed
stands for a closed connection. Both states \ssynsent and \ssynrcv denotes the
state that a TCP process can take during the initiation of a new connection. The
state \sestab represents the state when the connection is established,
\textit{i.e.} when two TCPs can communicate by sending and receiving data.
Then, two groups of states can be distinguished: the group formed by \sfwone,
\sfwtwo, \sclosing and \stimewait and that represents the state that a TCP can
reach when it is closing its connection and it is at the origin of the closing;
and the group formed by the state \sclosew and \slastack when the connection is
closing and the closing was initiated by the remote TCP.
Finally, the state \slisten stands for a TCP that is waiting for an incoming
connection.

\subsection{Flags}

The transition between the states are dictated by three things: user
actions, the reception of flags and timers. When a transition is triggered, a
response is given by sending of flags. The user actions, written in italic in
the automaton, can be an active or passive opening, or a close.
The corresponding user functions are detailed in the next paragraph.
The main flags that can be sent appear in figure~\ref{Fig:statemachine}.
They are sent alongside TCP segments, in the header.

The \syn flag is sent to synchronize two TCPs when they want to initiate a
connection. With a \ack flag, the last message received by the sender is
acknowledged. \fin is sent when a TCP wants to close the connection.
Moreover, a flag named \rst can be sent or received at any time and it forces the
connection to be reset. If so, both TCPs jump to the state \sclosed.

The state machine in figure~\ref{Fig:statemachine} does not reflect
error-conditions or any action which are not connected with the state changes.
Rather, it gives an overview of all the possible states a TCP connection could
reach over its lifetime. The transition are represented in the format
$\frac{\ \ x \ \ }{\ \ y \ \ }$ where $x$ is the event that triggers the
transition and $y$ is the response given to fulfil the transition.

\subsection{User functions}

In order to use the TCP protocol in programs, the nom defines user functions that
can be invoked by the software programmer in his code. This is the well-known
\emph{socket API}, available on almost all operating systems and known on POSIX
systems as the Berkeley sockets. This API specifies the already mentioned
\code{open} and \code{close} functions, as well as functions \code{send} and
\code{receive} to send or receive data once the connection is opened. The
behavior of each function is defined depending on the state of the TCP when the
function is called.
Even if user functions can initiate a change of state, it appears in the state
machine that most of transition are initiated by the reception of flags.

\subsection{Tasks}

A model with multiple tasks is adopted in the norm. Each task has a dedicated
job. One task is dedicated to the sending and the reception of messages. Another
is dedicated to the user functions and the last to the timers. Timers can resend
data or close a connection if the remote server does not acknowledged a message.
The first task mentioned performs most of the transitions.

\section{CycloneTCP library}
\label{sec:stack}

\input{OryxStack.tex}

CycloneTCP is a professional-grade embedded TCP/IP library developed by the
Oryx Embedded company. It provides several network protocols whose implementations
are meant to conform with the Request for Comments (RFC) Internet standards.
Figure~\ref{Fig:TcpStack} shows the protocols implemented in the library.
These include the IPv4 or Ipv6 of the network layer, TCP and UDP of the transport
layer and, among other, DHCP and HTTP of the application layer.
In particular, the TCP implementation is conform with the RFC 793
norm~\cite{rfc793}.

The library is written in ANSI C, and it supports a large number of 32-bits
embedded processors and a large number of Real-time Operating System (RTOS).
Hence, a model based on multiple tasks has been adopted to implement the TCP
protocol as described in the norm, with, in addition, complex cooperation
mechanisms. A task is dedicated to the user code, another to the timers and the
last is assigned to the processing of the incoming segments.
It can also run on bare metal environment.

The quality of the C implementation hailed in the AMNESIA:33 report~\cite{AMNESIA33}
as one of the most resilient TCP/IP stack. This gives a really good assurance
in the code and it strengthens the confidence that we can have in a formally
verified code running inside the library environment.
\todo{Add anything else relevant about the CycloneTCP library}.

\section{Specification techniques used}
\label{sec:spec}

\subsection{The SPARK programming language}

Ada is a general-purpose procedural programming language. The design of the Ada
language puts great emphasis on the safety and correctness of the program. This
objective is realized by using a readable syntax that uses keywords instead of
symbols where reasonable. The type system is strong and strict and many
potential violations of type constraints can be detected statically by the
compiler. If not, a run-time check is inserted into the program, to guarantee
the detection of incorrect situations. Ada 2012 introduced contract based
programming to Ada. In particular, it is possible to attach pre- and
postconditions to subprograms\footnote{In Ada, a distinction is made between
  functions that return a value, and procedures, which do
  not. \emph{Subprogram} is the term that designates both.}.  These conditions
can be checked during the execution of the program, just like assertions.

SPARK~\cite{mccormick2015building} is the name of a platform that provides
formal verification for Ada. It uses the user-provided contracts and attempts
to prove that the runtime checks cannot fail and that postconditions are
established by the corresponding subprograms.  As formal verification for the
whole Ada language would be intractable, SPARK is also the name of the subset
of the Ada language that is supported by the SPARK tool called
GNATprove\footnote{\url{http://docs.adacore.com/spark2014-docs/html/ug/}}.
This subset contains almost all features of Ada, though sometimes in a
restricted form.  In particular, expressions should be free from side effects,
and aliasing is forbidden (no two variables should share the same memory
location or overlap in memory), including when using pointers thanks to the use
of an ownership policy~\cite{dross2020recursive}.  This restriction greatly
simplifies the memory model used in the SPARK tool: any program variables can
be reasoned about independently from other variables.

GNATprove performs two kinds of analysis modularly on individual subprograms:
information flow to detect reads of uninitialized data and violations of flow
contracts; deductive verification based on the Why3 platform to generate
verification conditions for SMT solvers via a weakest-precondition calculus to
detect possible runtime errors and violations of functional contracts.

\subsection{Interfacing SPARK and C code}

To reduce the size of the work, only few functions of interest have been
translated in SPARK. Most of the C code has been kept.
Hopefully, SPARK integrates a mechanism to interface C code. No verification are
performed on the C code, but pre- and postconditions can be attached to
subprograms and they can be called in SPARK code.
The functional specification are not proved, thus it is the role of the
programmer to ensure the correctness of the contracts.

The types in SPARK can be more expressive than the types in C which adds more
safety to the SPARK code.
But the C and the SPARK code can share objects and it is essential that the
memory representations of an object are the same in the C and in the SPARK side.
For instance, if a C function is supposed to be called with a 32-bits integer,
it is the responsibility of the programmer to ensure that the function is
actually called with a 32-bits integer in the SPARK code.

Interfacing C and SPARK code is a great opportunity to plug our work in an
existing library, but it is also be a source of errors if it is not done
carefully. In section~\ref{sec:verif} it is explain how we have dealt with these
problems to ensure the quality of our work.

\subsection{Dealing with pointers}
\label{sec:pointers}

In CycloneTCP library, pointers are often manipulated at top-level and transmitted
to C functions. They are also used as fields in structures, for instance in
the socket structure to store other complex data-structures and recursive pointer
based data-structures. SPARK use a pointer ownership model to prevent bugs
involving pointers such as memory leak or double free. When a memory area is
allocated, a pointer points to this memory area and thus it is owned.
The ownership of the memory can be moved but the memory must be deallocated
before the end of the execution of the program.

\subsection{Specifying the frame condition}

The socket structure is shared between different functions, and contains a lot
of fields. Most of the fields are used in other parts of library and are not
relevant for our verification. We only need to concentrate on a subset of fields
to make the specifications easier to write, and to understand.
This can be achieve in SPARK by using \emph{models} and \emph{ghost code}.
The ghost code is code uses by GNATprove to simplify the proof that is removed
by the compiler to keep the size of executable file reasonable.

In the contract, we use the construction \spark{Model(Sock)} to refer the
relevant fields of the object \spark{Sock} where
\spark{Model} is a ghost function that extract the fields of interest from
\spark{Sock}, like in the construction frequently used in postconditions
\begin{lstlisting}[language=Ada, basicstyle=\small\ttfamily]
Model(Sock) = (Model(Sock)'Old with delta
                    State => TCP_STATE_CLOSED)
\end{lstlisting}
that states that relevant fields of \spark{Sock} selected by \spark{Model} are
not changed by the function call, except the fields \spark{State} that is equal
to \spark{TCP_STATE_CLOSED} after the call.

\section{Conformance to the TCP protocol}
\label{sec:verif}

The reader can refer to the code at \url{https://github.com/AdaCore/Http_Cyclone}.

\subsection{Extracting a specification from the TCP norm}

Extracting a specification
is mandatory to provide functional contracts to the user functions. That can
be achieved by reading the norm. However, the norm is underspecified in many
cases and only one possible implementation is depicted. Thus, the specifications
we have extracted are conformant with the norm but they slightly differ to what
can be found in the RFC.
Among the specification we have extracted:
\begin{itemize}
\item The transitions between the states must respect the order given by the
state machine described in figure~\ref{Fig:statemachine}.
In particular, a transition from a state to
another is valid only if there exists a transition in the state machine between
these two states.
\item The user functions contain functional specifications in the norm, that
have to be done  when a function is called depending on the state of the
socket~\cite[p. 52]{rfc793}. In particular it is specified in what state the
function can be called without returning an error.
\end{itemize}

\subsection{Rewriting the TCP user functions}

The user functions of the TCP protocol have been rewritten in SPARK. Besides
proving the absence of run-time error in these functions, it has been possible
to check that the functional specifications of the TCP protocol are not violated
by the code.

% dire que les fonctions appelées dans le code on était modélisées pour prendre
% en compte les changements d'état TCP qu'elles occasionnent.

% Problème, certains changement d'état se font entre l'appel de deux fonctions,
% ça doit donc être pris en compte. Et certains état se font directement lors de
% l'appel d'une fonction particulière et ça doit aussi être pris en compte.

When a transition has to be made in the code, it is made through the helper
function \spark{TCP_Change_State}. The precondition of this function describes
all the allowed transitions. Therefore, we try to do an forbidden transition,
GNATprove will display a warning. This is the most direct way to make a
transition in the state machine.
As the TCP user functions are all protected with a mutex to avoid interferences
from other threads that could change the TCP state during their execution, in
particular when a new segment is received, it is quite easy to know the
possible set of state that can have the socket during the all execution of the
functions.

There is only one function that makes the verification harder because it
releases the mutex: the function \spark{TCP_Wait_For_Events}.
It takes as an arguments an event to wait and returns once the event
is triggered by the reception of a segment.
The role of this
function is to release the mutex and allow the reception task to process the
incoming segments. This function
can be used, for example, in the function \spark{open} when the connection
is initialized to wait until the connection has been established or not to
correctly set the error that indicates to the user the state of the connection.
If we consider that everything can happen during a call to \spark{TCP_Wait_For_Events},
the verification becomes too imprecise and it becomes impossible to prove the
functional specifications extracted from the norm. Instead we have decided to
find a better contract to this function.

\subsection{Dealing with synchronous changes of states}

The function \spark{TCP_Wait_For_Events} first checks if the events is not already
true. If not, the mutex is released until a segment that triggered the event is
received or until the timeout is reach. Each time a new segment is received, it
is processed and if it contains a segment that causes a change of state, the TCP
state is updated. Then the events are updated, and if the waited event becomes
true due to the change of state, it is triggered.

\begin{algorithm}[t]
    \caption{Function to compute the possible state after the completion of a
particular event that is requested by a user-task related function.}
\label{algo:waitForEvents}
\begin{algorithmic}[1]
\footnotesize
\Function{TCP\_Wait\_For\_Events\_Proof}{Socket, Event\_Mask}
    \State $S_{last} := \text{Socket}$
    \State $E :=$ \spark{TCP_Update_Events}($S_{last}$)
    \If{$(E\ \&\ \text{Event\_Mask}) \neq 0$}
        \State \textbf{return}  $S_{last}$
    \EndIf
    \For{$i=1$ \textbf{to} $3$}
        \State $S_{last} :=$ \spark{TCP_Process_One_Segment}($S_{last}$)
        \State $E :=$ \spark{TCP_Update_Events}($S_{last}$)
        \If{$(E\ \&\ \text{Event\_Mask}) \neq 0$}
            \State \textbf{return} $S_{last}$
        \EndIf
    \EndFor
    \State \textbf{return} $\emptyset$
\EndFunction
\end{algorithmic}
\end{algorithm}

An event is considered to be true or not depending on the state of the socket.
This computation is done by the function \spark{TCP_Update_Events}. We have put
a contract on it, which is proved to be correct by SPARK. Then we can write a
function that emulates what is done when \spark{TCP_Wait_For_Events} is called.
We have written a procedure \spark{TCP_Wait_For_Events_Proof}, reproduced in
algorithm~\ref{algo:waitForEvents} to compute all the possible TCP states that a
socket can have after a wait. This algorithm simulates the reception of
one message, update the events and return if the event waited is reached.
Otherwise it repeats the procedures. The for loop only requires three steps
because all the step in the state machine can be reached in three segment
receptions without user interaction.

On \spark{TCP_Wait_For_Events_Proof}, a contract has been put to describe all
the possible states that a socket can have after the wait. It is only possible
to prove this contract with SPARK if the function named
\spark{TCP_Process_One_Segment} in algorithm~\ref{algo:waitForEvents}
has a postcondition.

\subsection{Symbolic execution to extract contracts}

Rewriting the part that process the incoming segments was out of the scope of
this work. Instead, we have opted for the use of symbolic execution on the
original CycloneTCP C code.
Symbolic execution is a mean to execute programs with symbolic values rather
than concrete values. Symbolic execution has already been successfully
used with other verification methods.
Indeed, in~\cite{vanoverberghe2008using} the authors present how to use symbolic
execution to improve deductive verification, and in~\cite{kassios2012comparing}
the authors explain that symbolic execution can be good at infering contracts.
For our work, KLEE has been used. It is built on top of LLVM as it takes LLVM
bytecode in input and works well with C code thanks to a very simple interface.

\begin{figure}
\begin{lstlisting}[language=C, basicstyle=\footnotesize\ttfamily, numbers=left,
    numberstyle=\tiny, frame=bottomline, escapeinside={(*@}{@*)}]
// creation of a fake incoming segment
TCPheader *segment = malloc(sizeof(TCPheader));
klee_make_symbolic(segment, sizeof(segment), "seg");
klee_assume(segment->flag <= 31);
// Creation of the socket
Socket *sock = malloc(sizeof(Socket)), oldSock;
klee_make_symbolic(sock, sizeof(sock), "sock");
memcpy(&oldSock, sock, sizeof(Socket));

// The function that process the segments is called
tcpProcessSegment(sock, segment); (*@\label{code:kleedriver:tcpProcessSegment}@*)

// Contract to check
klee_assert( (*@\label{code:kleedriver:assert}@*)
    (oldSock.state == TCP_STATE_ESTABLISHED) ?
        sock.state == TCP_STATE_ESTABLISHED ||
        sock.state == TCP_STATE_CLOSE_WAIT ||
        sock.state == TCP_STATE_CLOSED :
    (oldSock.state == ...) ? ... )
)
\end{lstlisting}
\caption{Driver for the verification of \lstinline[language=C]{tcpProcessSegment}
with KLEE}
\label{code:kleedriver}
\end{figure}

The first step in order to extract contracts was to read the state machine to
know all the possible transitions that can be done by the reception of a flag,
and adapt it to
the subtleties of CycloneTCP. The next step is to write a driver to run KLEE and
know if the contracts that we have infered are correct or not.
A simplified example is shown in figure~\ref{code:kleedriver}. The first lines
create a symbolic incoming segment to process as well as a symbolic socket.
At line~\ref{code:kleedriver:tcpProcessSegment} the function that processes the
segment is called. It is followed at line~\ref{code:kleedriver:assert} by a KLEE
builtin that check if an assert is valid for all the symbolic paths
explored before.

Once the contract is validated by KLEE, it can be copied out on the SPARK
function \spark{TCP_Process_One_Segment}. With this step done, the concurrency
challenge has been solved and the function
\spark{TCP_Wait_For_Events} have a correct and an accurate contract, which
makes possible the verification of the functional correctness of the TCP user
functions thanks to SPARK.

\section{Hardening the user's API}
\label{sec:API}

It appears that lot of programmers do not correctly use the socket API that is
provided by the library. Providing verified user functions for TCP is not enough
to guarantee the safety of software if the users do not correctly use them.
An important part of the work was to ensure that code that uses the socket API's
functions are correctly called. By ``correctly'' we mean: 1)~the functions
are called with respect to a defined order and 2)~the return code is checked
before continuing the processing.

\subsection{Enforcing the correct order of API functions}

The TCP protocol implies a specific order such that the user can call the API's
functions without breaking the functional specification of the protocol. This
order also is being conveyed by the protocol's state machine given in
figure~\ref{Fig:statemachine}.
Ada's pre- and postconditions are a powerful tool to express such inter-function
dependencies, while SPARK technology can be used to guarantee that the
assertions always hold. Thus, post- and preconditions were introduced to model
a partial order on the calls to the API's functions.

\begin{figure}
\begin{lstlisting}[language=Ada,basicstyle=\footnotesize\ttfamily]
procedure Socket_Connect
   (Sock           : in out Not_Null_Socket;
    Remote_Ip_Addr : in     IpAddr;
    Remote_Port    : in     Port;
    Error          :    out Error_T)
   with
   Pre => Is_Initialized_Ip (Remote_Ip_Addr),
   Post =>
    if Sock.S_Type = SOCKET_TYPE_STREAM then
      (if Error = NO_ERROR then
        Sock.S_Remote_Ip_Addr = Remote_Ip_Addr)
    else
      Sock.S_Remote_Ip_Addr = Remote_Ip_Addr

procedure Socket_Send
   (Sock    : in out Not_Null_Socket;
    Data    : in     Send_Buffer;
    Written :    out Natural;
    Flags   :        Socket_Flags;
    Error   :    out Error_T)
   with
   Pre => Is_Initialized_Ip (Sock.S_Remote_Ip_Addr)
\end{lstlisting}
\caption{An example of how function calls can be ordered by pre- and
postconditions}
\label{fig:functionorder}
\end{figure}

Simplified contracts extracted from our library are reproduced in
figure~\ref{fig:functionorder}. If no error occurs, the function
\spark{Socket_Connect} sets \spark{Sock.S_Remote_Ip_Addr} to
\spark{Remote_Ip_Addr} which is supposed to be
initialized when the function is called. \spark{Socket_Send} requires to be
called with a socket not null such that its field \spark{Remote_Ip_Addr} is
initialized. The only way to ensure that the precondition hold is to
call \spark{Socket_Connect} before \spark{Socket_Send}.
The tool GNATprove can statically find if a call to \spark{Socket_Send}
is not preceded by a call to \spark{Socket_Connect}.

\subsection{Checking the correctness of return codes}

A common mistake is to forget to check the return codes after a call to a socket
API's function. For instance, the following code would be reject by SPARK:
\begin{lstlisting}[language=Ada,basicstyle=\small\ttfamily]
Socket_Connect(Sock, Ip_Addr, Port, Error);
Socket_Send(Sock, Data, Written, Flags, Error);
\end{lstlisting}
Indeed, if \spark{Socket_Connect} fails and \spark{Error} is not equal to
\spark{NO_ERROR}, \spark{Sock.S_Remote_Ip_Addr} might be non initialized, and
then the function \spark{Socket_Send} cannot be called because of its
precondition.

\section{Results}
\label{sec:results}

\subsection{Bugs found}

\subsubsection{Memory leak}

Thanks to the pointer ownership policy adopted by SPARK and detailed in
section~\ref{sec:pointers}, we have found that a memory leak can happen when
the connection have been closed and the buffer are cleaning.

\subsubsection{Violation of the TCP protocol}

Another very interesting and very subtle bug have been found thanks to SPARK is
reproduced in the snippet bellow.

\begin{lstlisting}[language=Ada, basicstyle=\footnotesize\ttfamily,
                    numbers=left, numberstyle=\tiny, escapechar=\%]
case Sock.State is
  when TCP_STATE_SYN_RECEIVED
     | TCP_STATE_ESTABLISHED =>
  -- Flush the send buffer
  TCP_Send (Sock, Buf, Ignore_Written,
            SOCKET_FLAG_NO_DELAY, Error);
  if Error /= NO_ERROR then
      return;
  end if;

  -- Make sure all the data has been sent out
  TCP_Wait_For_Events % \label{bugProg:tcpWaitForEvents} %
     (Sock       => Sock,
      Event_Mask => SOCKET_EVENT_TX_DONE,
      Timeout    => Sock.S_Timeout,
      Event      => Event);

  -- Timeout error?
  if Event /= SOCKET_EVENT_TX_DONE then
     Error := ERROR_TIMEOUT;
     return;
  end if;

  -- Send a FIN segment
  TCP_Send_Segment   % \label{bugProg:tcpWaitForEventsPost} %
     (Sock         => Sock,
      Flags        => TCP_FLAG_FIN or TCP_FLAG_ACK,
      Seq_Num      => Sock.sndNxt,
      Ack_Num      => Sock.rcvNxt,
      Length       => 0,
      Add_To_Queue => True,
      Error        => Error);

  -- Failed to send FIN segment?
  if Error /= NO_ERROR then
     return;
  end if;

  -- Switch to the FIN-WAIT-1 state
  TCP_Change_State (Sock, TCP_STATE_FIN_WAIT_1); % \label{bugProg:tcpChangeState} %
\end{lstlisting}

In this scenario, after the call to \spark{Tcp_Send}, when the error cases have
been filter, the state of the socket can either be \sestab or \sclosew. The call
at line~\ref{bugProg:tcpWaitForEvents} and now the state of the socket can be
still be \sestab or \sclosew, or if a \rst flag has been received when the mutex
was released, the state can be \sclosed too.
At line~\ref{bugProg:tcpWaitForEventsPost}, a \fin segment is sent, which does
not modify the state of the socket. Finally, at line~\ref{bugProg:tcpChangeState}
we try to jump to the state \sfwone and thus the transitions $\text{\sclosew}
\rightarrow \text{\sfwone}$ and $\text{\sclosed} \rightarrow \text{\sfwone}$ are
possible.

But these transition are not allowed by the norm, and thus not allowed by the
precondition of \spark{TCP_Change_State}. The problem is warm by SPARK and a fix
has been proposed.

\subsection{Verified parts}

For achieving this verification, 50 subprograms have been translated from C to
SPARK. This accounts for \np{2266} lines of code of which \np{1165} are logical
code to write and prove the specifications.

\section{Conclusion and future work}

The use of SPARK for the verification of the TCP protocol has given good
results. It has helped to find bugs in the existing implementation. But more
than finding bugs, the interest of this work lie in the face that it is a big
step toward a secure implementation. It has been proved by the tool GNATprove
that our implementation is free of run-time errors and that all the transitions
in the TCP state machine are done with respect to the norm.

However weeknesses exist in our implementation, in particular because all the
underlying layers are still written in C. The principal function of those layers
is to format packet before they are sent, or parse incoming packets, check their
integrity and transmit the resulting payload to the corresponding upper level
layer. This processing part can be a source of errors and bugs as explain
in~\cite{Reiher2019RecordFluxFM}.
Using the RecordFlux DSL to parse the packets of the different protocols is the
next step to make the stack safer. Using RecordFlux to parse TCP message
would be a big step to finish the translation of the TCP protocol in SPARK.
The postconditions extracted by KLEE could be proved by GNATprove for a
safer result.

Our verification of the TCP protocol only focuses on the validity of the
transitions in the TCP state machine, but some other properties are formulated
in the TCP norm, and could be verified in the implementation to make it even
more robust.

\bibliographystyle{IEEEtran}
\bibliography{biblio}

\end{document}

### 2021-05-12 SecDev21/paper.tex
Minor fixes

@@
 \IEEEauthorblockN{Guillaume Cluzel}
-\IEEEauthorblockA{\textit{Adacore \& ENS de Lyon}}
+\IEEEauthorblockA{\textit{AdaCore \& ENS de Lyon}}
 \and
 \IEEEauthorblockN{Kyriakos Georgious}
-\IEEEauthorblockA{\textit{Adacore \& University of Bristol}}
+\IEEEauthorblockA{\textit{AdaCore \& University of Bristol}}
 \and
 \IEEEauthorblockN{Yannick Moy}
-\IEEEauthorblockA{\textit{Adacore}}
+\IEEEauthorblockA{\textit{AdaCore}}
 \and
@@
 We present the TCP protocol in section~\ref{sec:TCP} and then briefly the
-choosen stack, the Oryx Embedded one, in section~\ref{sec:stack}
+chosen stack, the Oryx Embedded one, in section~\ref{sec:stack}
 before presenting the specification techniques used in section~\ref{sec:spec}
@@
 Every group of states corresponds to a one phase. More precisely \sclosed
-stands for a closed connection. Both states \ssynsent and \ssynrcv denotes the
+stands for a closed connection. Both states \ssynsent and \ssynrcv denote the
 state that a TCP process can take during the initiation of a new connection. The
@@
 to make the specifications easier to write, and to understand.
-This can be achieve in SPARK by using \emph{models} and \emph{ghost code}.
-The ghost code is code uses by GNATprove to simplify the proof that is removed
-by the compiler to keep the size of executable file reasonable.
+This can be achieved in SPARK by using \emph{models} and \emph{ghost code}
+which are used by GNATprove for analysis but not compiled in the final executable.
 
@@
 function \spark{TCP_Change_State}. The precondition of this function describes
-all the allowed transitions. Therefore, we try to do an forbidden transition,
-GNATprove will display a warning. This is the most direct way to make a
+all the allowed transitions. Therefore, when we try to do a forbidden transition,
+GNATprove will display a message. This is the most direct way to make a
 transition in the state machine.
@@
 releases the mutex: the function \spark{TCP_Wait_For_Events}.
-It takes as an arguments an event to wait and returns once the event
+It takes as argument an event to wait and returns once the event
 is triggered by the reception of a segment.
@@
 
-The function \spark{TCP_Wait_For_Events} first checks if the events is not already
+The function \spark{TCP_Wait_For_Events} first checks if the event is not already
 true. If not, the mutex is released until a segment that triggered the event is
@@
 dependencies, while SPARK technology can be used to guarantee that the
-assertions always hold. Thus, post- and preconditions were introduced to model
+assertions always hold. Thus, pre- and postconditions were introduced to model
 a partial order on the calls to the API's functions.
@@
 But these transition are not allowed by the norm, and thus not allowed by the
-precondition of \spark{TCP_Change_State}. The problem is warm by SPARK and a fix
+precondition of \spark{TCP_Change_State}. The problem is detected by SPARK and a fix
 has been proposed.

### 2021-05-13 SecDev21/paper.tex
Rewrite of the TCP protocol description

@@
 \documentclass[conference]{IEEEtran}
-
-% Class modification
-\IEEEilabelindent=0pt
 
@@
 
-\section{TCP protocol}
+\section{Overview of the TCP protocol specification}
 \label{sec:TCP}
 
+This section gives an overview of the TCP functional specification without going
+into unnecessary depth. A complete description of the protocol’s specification
+can be found at the RFC 793 document~\cite{rfc793}.
+
+Networking communication models define the processes of transferring information
+from one network component to another. Figure~\ref{Fig:TcpStack}
+shows an overview of the OSI model with its different layers.
+Layering is an essential design
+requirement for both models to achieve modularity, flexibility, and abstraction.
+Applications that only need to use the functionality of the lower levels can
+drop all the unnecessary upper levels. Furthermore, layering enables easier
+troubleshooting as the network's functionality is well-defined and split between
+the different layers. Each layer supports a wide range of protocols that are
+compatible with the layer's specifications. The lowest layers, bottom layers in
+Figure~\ref{Fig:TcpStack},
+are more hardware-oriented, while the top layers are more software specific.
+While the \emph{application} level
+has a wide range of protocols that can be used depending on the application
+specification, the \emph{transport} layer is heavily depended on the two
+dominant protocols: Transmission Control Protocol (TCP) and
+User Datagram Protocol (UDP).
+
+TCP is a connection-oriented protocol; a
+connection between the sender and the receiver must be established before
+transmitting any data. Furthermore, TCP is a reliable protocol since it
+guarantees the delivery of all the messages and that they are delivered in the
+order they were sent. It also provides error-checking mechanisms to discard and
+recover any corrupted data.
+
 Initially described by the ``fathers of the Internet'' Vincent Cerf and Bob
-Kahn, the TCP protocol was standardized in 1981 by RFC 793~\cite{rfc793}.
-Its sturdiness makes it still one of the most widely used protocols today with
-UDP.
-
-In the OSI networking communication model, the TCP protocol is central and
-located in the \emph{Transport layer}. It itself uses the \emph{Network layer}
-to work.
-The security of the \emph{Application} protocols depends on the security of
-the \emph{Session layer} and consequently on the security of the TCP and UDP
-protocols. Figure~\ref{Fig:TcpStack} gives an overview of how protocols are
-stacked.
-
-The purpose of this section is to give enough context to enable the reader to
-follow the work conducted in this article.
-The TCP protocol is connection-oriented in the sense that a connection requires
-to be established between the two processes that exchange data.
-The protocol is governed by a state machine described in
-figure~\ref{Fig:statemachine}.
-The implementation is rather free and lots of choices are left to the
-programmer. However, the implementation must ensure some properties to be able
-to communicate with other TCPs.
+Kahn, the TCP protocol was standardized in 1981 in~\cite{rfc793}.
+Its sturdiness still makes it one of the most widely used protocols 40 years
+later. Indeed, almost any networking application is using either the TCP or the
+UDP protocols to traffic data between the Network and the Application layers.
 
@@
 
-\subsection{States}
+\subsection{TCP state machine}
 
@@
 The behavior of the three-phase communication can be described by the state
-machine.
-Every group of states corresponds to a one phase. More precisely \sclosed
-stands for a closed connection. Both states \ssynsent and \ssynrcv denote the
-state that a TCP process can take during the initiation of a new connection. The
-state \sestab represents the state when the connection is established,
-\textit{i.e.} when two TCPs can communicate by sending and receiving data.
-Then, two groups of states can be distinguished: the group formed by \sfwone,
-\sfwtwo, \sclosing and \stimewait and that represents the state that a TCP can
-reach when it is closing its connection and it is at the origin of the closing;
-and the group formed by the state \sclosew and \slastack when the connection is
-closing and the closing was initiated by the remote TCP.
-Finally, the state \slisten stands for a TCP that is waiting for an incoming
-connection.
+machine in figure~\ref{Fig:statemachine}.
 
-\subsection{Flags}
+An edge represents a state transition. A transition is either triggered by an
+explicit action, stated in italics format on the label's edge, for example,
+\textit{Close}, or is triggered by the arrival of a specific flag/s.
+An edge-label in the format of $\frac{\ \ x \ \ }{\ \ y \ \ }$
+represents the associated flags transmitted with each transition, where $x$
+is either send or receive and $y$ refers to the actual
+flag/s transmitted in response to the action that caused the state-transition.
+An explicit action is either a user-triggered action (through the user's API)
+or an automatically performed action triggered by a timeout event. The flags are
+embedded in the header section of a transmitted segment and can be one of the
+following:
+\begin{itemize}
+\item \ack: the last message received by the sender is acknowledged.
+\item \syn: this flag is sent to establish a connection.
+\item \fin: this flag is sent to close the connection.
+\item \rst: this flag is sent to reset the connection, often when an error
+occurred in one or the other side.
+\end{itemize}
 
-The transition between the states are dictated by three things: user
-actions, the reception of flags and timers. When a transition is triggered, a
-response is given by sending of flags. The user actions, written in italic in
-the automaton, can be an active or passive opening, or a close.
-The corresponding user functions are detailed in the next paragraph.
-The main flags that can be sent appear in figure~\ref{Fig:statemachine}.
-They are sent alongside TCP segments, in the header.
+The state machine in figure~\ref{Fig:statemachine} does not represent the
+complete protocol specification; for example, it does not reflect
+error-conditions or any actions which are not connected with the state changes.
+Rather, it gives an overview of all the possible states a TCP connection could
+reach over its lifetime.
 
-The \syn flag is sent to synchronize two TCPs when they want to initiate a
-connection. With a \ack flag, the last message received by the sender is
-acknowledged. \fin is sent when a TCP wants to close the connection.
-Moreover, a flag named \rst can be sent or received at any time and it forces the
-connection to be reset. If so, both TCPs jump to the state \sclosed.
+Each state has a meaning. \slisten represents waiting for a connection request
+from any remote TCP and port. \ssynsent and \ssynrcv represent the
+initialization of the connection, that ends when \sestab is reached.
+\sfwone, \sfwtwo, \sclosing and \stimewait are a group of states that represents
+waiting for a connection termination request from remote TCP. \sclosew and
+\slastack represent waiting for a connection termination request from local
+user. Finally, \sclosed represents no connection state at all.
 
-The state machine in figure~\ref{Fig:statemachine} does not reflect
-error-conditions or any action which are not connected with the state changes.
-Rather, it gives an overview of all the possible states a TCP connection could
-reach over its lifetime. The transition are represented in the format
-$\frac{\ \ x \ \ }{\ \ y \ \ }$ where $x$ is the event that triggers the
-transition and $y$ is the response given to fulfil the transition.
+\subsection{TCP Multitasking model}
 
-\subsection{User functions}
-
-In order to use the TCP protocol in programs, the nom defines user functions that
-can be invoked by the software programmer in his code. This is the well-known
-\emph{socket API}, available on almost all operating systems and known on POSIX
-systems as the Berkeley sockets. This API specifies the already mentioned
-\code{open} and \code{close} functions, as well as functions \code{send} and
-\code{receive} to send or receive data once the connection is opened. The
-behavior of each function is defined depending on the state of the TCP when the
-function is called.
-Even if user functions can initiate a change of state, it appears in the state
-machine that most of transition are initiated by the reception of flags.
-
-\subsection{Tasks}
-
-A model with multiple tasks is adopted in the norm. Each task has a dedicated
-job. One task is dedicated to the sending and the reception of messages. Another
-is dedicated to the user functions and the last to the timers. Timers can resend
-data or close a connection if the remote server does not acknowledged a message.
-The first task mentioned performs most of the transitions.
+Different tasks can interact and update the socket data structure to handle the
+various events that can occur within a TCP session. The TCP norm, under
+Section “3.9. Event Processing” page 52, describes a possible implementation
+of how to handle these events based on three tasks: one for the
+\emph{user calls}, one for the \emph{arriving segments} and one for the
+\emph{timers}.
+The role of each of the three tasks can be summarized as follow:
+User calls refer to functions, namely \texttt{open}, \texttt{close},
+\texttt{abort}, \texttt{send} and \texttt{receive}, that can be called
+by the user to control the connection, send, or receive data.
+These functions can trigger transitions between the connection's states since
+they are intended to control the connection.
+In the arriving segments task, the received segments are being processed,
+and the corresponding messages are sent back. Transitions between
+states can be triggered on the reception of a segment.
+Finally, timers control the timeouts, for example, the retransmission
+timeout to retransmit a message, or the time-wait timeout to close the
+connection after a specified amount of time elapsed. Thus, corresponding
+transitions to the timeout events can also be triggered by this task.
 
@@
 It can also run on bare metal environment.
+
+In the CycloneTCP implementation, the socket data structure is used to retain
+the status of a TCP connection. Furthermore, a socket can be
+manipulated by the user through the TCP's library Application Programming
+Interface (API) to perform various TCP operations, such as the transmission of
+data.
 

### 2021-05-18 SecDev21/paper.tex
reshaping the paper

@@
 \usepackage{listings}
+\usepackage{cleveref}
 
@@
 
-The usage of IoT devices has been put more within everyone's reach. Their
-vulnerability is a major cause for concern, and as their number
-increases, the number of attacks they suffer increases, putting at risk
-the security of the owner's data collected by the object. \todo{Try to find a
-relevant ref. I have found examples about the possibility to leak personnal data
-on Google home/Alexa}.
+By 2030 it is estimated that one trillion Internet of Things (IoT) devices will be deployed and be part of every aspect of our daily life~\cite{IoT_route:ARM_white_paper}.
+At the same time, it is made evident that security for such systems is a major challenge~\cite{Sec_IoT_Challenges:paper,Cyber_Systems_security:paper}.
+Any vulnerability in software libraries that are widely used by embedded systems can potentially open the door for compromising the security of billions of deployed IoT devices.
+In 2019, critical security vulnerabilities were discovered in a TCP implementation that has been used potentially by billions of IoT devices for more than 13 years \cite{zero_days_vuln:ARMIS_white_paper}, with some of them even allowing remote code execution.
 
-The core of the IoT devices is built on top of the TCP protocol~\cite{rfc793} as
-it is the most widely protocol used on the Internet, as the underlying protocol
-for application layers such as HTTP, FTP or TLS. TCP is based on the underlying
-IP protocol which is itself based on a link layer such as Ethernet. This
-stacking of protocols is referred to as the TCP/IP stack.
+Networking communication models define the processes of transferring information from one network component to another.
+Currently, the two most popular networking communication models are the TCP/IP (Transmission Control Protocol/Internet Protocol) \cite{TCP_IP:ietf_tutorial} and the OSI (Open System Interconnection) \cite{ISO:35.100} models.
+Layering is an essential design requirement for both models to achieve modularity, flexibility, and abstraction.
+Applications that only need to use the lower levels' functionality can drop all the unnecessary upper levels.
+\Cref{Fig:TcpStack} shows the various layers of the OSI model.
+Each layer supports a wide range of protocols that are compatible with the layer's specifications.
+The lowest layers, bottom layers in \Cref{Fig:TcpStack}, are more hardware-oriented, while the top layers are more software-specific.
+While the \emph{application} level has a wide range of protocols that can be used depending on the application specification, the \emph{transport} layer is heavily dependent on the two dominant protocols: Transmission Control Protocol (TCP) and User Datagram Protocol (UDP).
+Today, almost every networking application, including IoT devices, is using either the TCP or the UDP protocols to traffic data between the Network and the Application layers.
+%Thus, the reliability, assurance, and security of these protocols are of paramount importance for critical applications.
 
-Given the importance of the TCP protocol in network communications, some
-researchers have attempted to formalize the protocol to either prove key
-properties~\cite{smith1996formal} or even to fully prove the correctness of an
-implementation against a complete specification~\cite{ridge2008rigorous}.
-This latter implementation was performed in the HOL proof assistant, and while
-the authors note that it would be possible in theory to extract an
-implementation in Haskell from this work, this has not been adopted in practice.
-The formalization of TCP specification and verification of a TCP implementation
-face two difficulties:
-RFC 793 is written in English, leaving many parts underspecified; the TCP
-protocol is inherently concurrent which makes it complex to specify and verify.
-As a result, there is currently no formally verified implementation for TCP that
-is usable in industry.
+\input{OryxStack.tex}
 
-Other communication protocols have been formalized in recent years, leading to a
-formally verified implementation. The most prominent project in this area is the
-Everest project~\cite{bhargavan2017everest}, which led to a formally verified
-implementation in F* of TLS called miTLS~\cite{bhargavan2013implementing}. The
-approach of the Everest project is to implement TLS in F*, a programming
-language targeted at formal program verification.
-Note that miTLS still relies on an untrusted implementation of TCP, and thus
-could benefit from the results of our work.
-This approach is often difficult to adopt in industry, as codebases are evolved
-incrementally rather than replaced, both for economical and practical reasons.
+Existing research deals with formally verifying several protocols TCP/IP stack levels.
+For example, the work in \emph{miTLS} \cite{bhargavan2013implementing} formally verifies an SSL/TLS protocol implementation, and the work in \cite{Reiher_2020} uses a technology called RecordFlux to safely parse data segments.
+Given the reliance of the other software-level protocols on the TCP protocol, formalizing the its specification and verifying its implementation is critical.
+Achieving these two goals in practice is challening.
+This is mainly because TCP's specification, \cite{rfc793}, is written in English rather than any formal language, leaving many parts underspecified.
+Moreover, the TCP protocol is inherently concurrent, making it complex to specify and verify.
+Previous work focused on formalizing the protocol to either prove key properties \cite{smith1996formal} or fully verify the TCP protocol's specifications \cite{ridge2008rigorous} using the HOL proof assistant tool.
+While for the later work, the authors claim that it is possible from their formalization to extract an actual implementation in Haskell, this has not been achieved yet.
+At this point of writing, we are not aware of any formally verified TCP protocol implementation.
 
-Our approach in this work has been to incrementally evolve an existing TCP/IP
-stack in C by replacing parts of the code with formally verified code in SPARK.
-The stack chosen, CycloneTCP\footnote{\url{https://www.oryx-embedded.com/products/CycloneTCP.html}},
-a mature TCP/IP stack targeted at embedded processors with no Operating System.
+In this paper, we take a practical approach, where the security of an existing TCP/IP stack implementation written in ANSI C programming language is incrementally enhanced by replacing parts of its code with formally verified code in SPARK \cite{mccormick_chapin_2015}.
+SPARK 2014 is a programming language designed as a subset of the Ada language to produce highly reliable software through formal verification.
+For this work, the CycloneTCP, a professional-grade embedded TCP/IP library developed by the Oryx Embedded company\footnote{\url{https://www.oryx-embedded.com/products/CycloneTCP.html}}.
+\Cref{Fig:TcpStack} shows the protocols implemented in the library.
+In this work, we focused on enhancing the TCP and socket components of the stack.
+The TCP implementation is meant to conform with the RFC 793 protocol specifications \cite{rfc793}.
+The quality of the C implementation hailed in the AMNESIA:33 report \cite{AMNESIA33} as one of the most resilient TCP/IP stack.
+Using a starting point, an existing well-tested and widely used TCP/IP implementation, the IoT industry can have an immediate benefit by adopting the proposed by this work security-enhanced version of this implementation.
+
 We present the TCP protocol in section~\ref{sec:TCP} and then briefly the
@@
 \label{sec:TCP}
-
-This section gives an overview of the TCP functional specification without going
-into unnecessary depth. A complete description of the protocol’s specification
-can be found at the RFC 793 document~\cite{rfc793}.
-
-Networking communication models define the processes of transferring information
-from one network component to another. Figure~\ref{Fig:TcpStack}
-shows an overview of the OSI model with its different layers.
-Layering is an essential design
-requirement for both models to achieve modularity, flexibility, and abstraction.
-Applications that only need to use the functionality of the lower levels can
-drop all the unnecessary upper levels. Furthermore, layering enables easier
-troubleshooting as the network's functionality is well-defined and split between
-the different layers. Each layer supports a wide range of protocols that are
-compatible with the layer's specifications. The lowest layers, bottom layers in
-Figure~\ref{Fig:TcpStack},
-are more hardware-oriented, while the top layers are more software specific.
-While the \emph{application} level
-has a wide range of protocols that can be used depending on the application
-specification, the \emph{transport} layer is heavily depended on the two
-dominant protocols: Transmission Control Protocol (TCP) and
-User Datagram Protocol (UDP).
 
@@
 \label{sec:stack}
-
-\input{OryxStack.tex}
 

### 2021-05-18 SecDev21/paper.tex
more fixes

@@
 SPARK 2014 is a programming language designed as a subset of the Ada language to produce highly reliable software through formal verification.
-For this work, the CycloneTCP, a professional-grade embedded TCP/IP library developed by the Oryx Embedded company\footnote{\url{https://www.oryx-embedded.com/products/CycloneTCP.html}}.
-\Cref{Fig:TcpStack} shows the protocols implemented in the library.
-In this work, we focused on enhancing the TCP and socket components of the stack.
-The TCP implementation is meant to conform with the RFC 793 protocol specifications \cite{rfc793}.
-The quality of the C implementation hailed in the AMNESIA:33 report \cite{AMNESIA33} as one of the most resilient TCP/IP stack.
-Using a starting point, an existing well-tested and widely used TCP/IP implementation, the IoT industry can have an immediate benefit by adopting the proposed by this work security-enhanced version of this implementation.
+For this work, the CycloneTCP, a professional-grade embedded TCP/IP library developed by the Oryx Embedded company \cite{CycloneTCP}.
+\Cref{Fig:TcpStack} shows the protocols supported by the library.
+This work focuses on the TCP and socket components of the stack.
+The library's TCP implementation is meant to conform with the RFC 793 protocol specifications \cite{rfc793}.
+The quality of the CycloneTCP is acknowledged by the AMNESIA:33 report \cite{AMNESIA33}, which classifies it as one of the most resilient TCP/IP stack.
+By using as a starting point such a well-tested and widely used TCP/IP implementation, the IoT industry can have an immediate benefit by adopting this work's security-enhanced version of the CycloneTCP library.
 
@@
 
-Initially described by the ``fathers of the Internet'' Vincent Cerf and Bob
-Kahn, the TCP protocol was standardized in 1981 in~\cite{rfc793}.
-Its sturdiness still makes it one of the most widely used protocols 40 years
-later. Indeed, almost any networking application is using either the TCP or the
-UDP protocols to traffic data between the Network and the Application layers.
-
 \input{Automaton.tex}
 
-\subsection{TCP state machine}
+%\subsection{TCP state machine}
 
-Generally, every TCP communication session will go through the following three
-phases (if no error occurs): 1)~opening the connection 2)~sending and receiving
-the data and 3)~closing the connection.
-
-The behavior of the three-phase communication can be described by the state
-machine in figure~\ref{Fig:statemachine}.
+Generally, every TCP communication session will go through the following three phases (if no error occurs): 1)~opening the connection 2)~sending and receiving
+the data and 3)~closing the connection. The behavior of the three-phase communication can be described by the state machine in \cref{Fig:statemachine}.
 

### 2021-05-18 SecDev21/paper.tex
shortening the information sections

@@
 
-TCP is a connection-oriented protocol; a
-connection between the sender and the receiver must be established before
-transmitting any data. Furthermore, TCP is a reliable protocol since it
-guarantees the delivery of all the messages and that they are delivered in the
-order they were sent. It also provides error-checking mechanisms to discard and
-recover any corrupted data.
+\input{Automaton.tex}
 
-\input{Automaton.tex}
+TCP is a connection-oriented protocol; a connection between the sender and the receiver must be established before transmitting any data.
+Furthermore, TCP is a reliable protocol since it guarantees the delivery of all the messages and that they are delivered in the order, they were sent.
+It also provides error-checking mechanisms to discard and recover any corrupted data.
 
@@
 
-Generally, every TCP communication session will go through the following three phases (if no error occurs): 1)~opening the connection 2)~sending and receiving
-the data and 3)~closing the connection. The behavior of the three-phase communication can be described by the state machine in \cref{Fig:statemachine}.
-
-An edge represents a state transition. A transition is either triggered by an
-explicit action, stated in italics format on the label's edge, for example,
-\textit{Close}, or is triggered by the arrival of a specific flag/s.
-An edge-label in the format of $\frac{\ \ x \ \ }{\ \ y \ \ }$
-represents the associated flags transmitted with each transition, where $x$
-is either send or receive and $y$ refers to the actual
-flag/s transmitted in response to the action that caused the state-transition.
-An explicit action is either a user-triggered action (through the user's API)
-or an automatically performed action triggered by a timeout event. The flags are
-embedded in the header section of a transmitted segment and can be one of the
-following:
+Generally, every TCP communication session will go through the following three phases (if no error occurs): 1)~opening the connection, 2)~sending and receiving the data, and 3)~closing the connection.
+The behaviour of the three-phase communication can be described by the state machine given in \cref{Fig:statemachine}.
+An edge represents a state transition.
+A transition is either triggered by an explicit action, stated in italics format on the label's edge, for example, \textit{Close}, or is triggered by the arrival of a specific flag/s.
+An edge-label in the format of $\frac{\ \ x \ \ }{\ \ y \ \ }$ represents the associated flags transmitted with each transition, where $x$ is either send or receive and $y$ refers to the actual flag/s transmitted in response to the action that caused the state transition.
+An explicit action is either a user-triggered action (through the user's API) or an automatically performed action triggered by a timeout event.
+The flags are embedded in the header section of a transmitted segment and can be one of the following:
 \begin{itemize}
@@
 \item \rst: this flag is sent to reset the connection, often when an error
-occurred in one or the other side.
+occurred on one or the other side.
 \end{itemize}
 
-The state machine in figure~\ref{Fig:statemachine} does not represent the
-complete protocol specification; for example, it does not reflect
-error-conditions or any actions which are not connected with the state changes.
-Rather, it gives an overview of all the possible states a TCP connection could
-reach over its lifetime.
+The state machine in \cref{Fig:statemachine} does not represent the complete protocol specification; for example, it does not reflect error conditions or any actions which are not connected with the state changes.
+Instead, it gives an overview of all the possible states a TCP connection could reach over its lifetime.
+Each state has a meaning.
+\slisten represents waiting for a connection request from any remote TCP and port.
+\ssynsent and \ssynrcv represent the initialization of the connection that ends when \sestab is reached.
+\sfwone, \sfwtwo, \sclosing, and \stimewait are a group of states that represents waiting for a connection termination request from remote TCP.
+\sclosew and \slastack represent waiting for a connection termination request from the local user.
+Finally, \sclosed represents no connection state at all.
 
-Each state has a meaning. \slisten represents waiting for a connection request
-from any remote TCP and port. \ssynsent and \ssynrcv represent the
-initialization of the connection, that ends when \sestab is reached.
-\sfwone, \sfwtwo, \sclosing and \stimewait are a group of states that represents
-waiting for a connection termination request from remote TCP. \sclosew and
-\slastack represent waiting for a connection termination request from local
-user. Finally, \sclosed represents no connection state at all.
+%\subsection{TCP Multitasking model}
 
-\subsection{TCP Multitasking model}
+TCP is based on a multitasking model.
+Different tasks can interact and update the socket data structure, used to retain the status of a TCP connection, to handle the various events that can occur within a TCP session.
+The TCP norm, \cite{rfc793} , under Section "3.9.
+Event Processing" page 52, describes a possible implementation of how to handle these events based on three tasks: one for the \emph{user calls}, one for the \emph{arriving segments} and one for the \emph{timers}.
+The role of each of the three tasks can be summarized as follow: User calls refer to functions, namely \texttt{open}, \texttt{close}, \texttt{abort}, \texttt{send} and \texttt{receive}, that can be called by the user to control the connection, send, or receive data.
+These functions can trigger transitions between the connection's states since they are intended to control the connection.
+In the arriving segments task, the received segments are being processed,
+and the corresponding messages are sent back.
+Transitions between states can be triggered on the reception of a segment.
+Finally, timers control the timeouts, for example, the retransmission timeout to retransmit a message or the time-wait timeout to close the connection after a specified amount of time elapsed.
+Thus, corresponding transitions to the timeout events can also be triggered by this task.
+A complete description of the protocol's specification can be found at the \emph{RFC 793} document \cite{rfc793}.
 
-Different tasks can interact and update the socket data structure to handle the
-various events that can occur within a TCP session. The TCP norm, under
-Section “3.9. Event Processing” page 52, describes a possible implementation
-of how to handle these events based on three tasks: one for the
-\emph{user calls}, one for the \emph{arriving segments} and one for the
-\emph{timers}.
-The role of each of the three tasks can be summarized as follow:
-User calls refer to functions, namely \texttt{open}, \texttt{close},
-\texttt{abort}, \texttt{send} and \texttt{receive}, that can be called
-by the user to control the connection, send, or receive data.
-These functions can trigger transitions between the connection's states since
-they are intended to control the connection.
-In the arriving segments task, the received segments are being processed,
-and the corresponding messages are sent back. Transitions between
-states can be triggered on the reception of a segment.
-Finally, timers control the timeouts, for example, the retransmission
-timeout to retransmit a message, or the time-wait timeout to close the
-connection after a specified amount of time elapsed. Thus, corresponding
-transitions to the timeout events can also be triggered by this task.
-
-\section{CycloneTCP library}
-\label{sec:stack}
-
-CycloneTCP is a professional-grade embedded TCP/IP library developed by the
-Oryx Embedded company. It provides several network protocols whose implementations
-are meant to conform with the Request for Comments (RFC) Internet standards.
-Figure~\ref{Fig:TcpStack} shows the protocols implemented in the library.
-These include the IPv4 or Ipv6 of the network layer, TCP and UDP of the transport
-layer and, among other, DHCP and HTTP of the application layer.
-In particular, the TCP implementation is conform with the RFC 793
-norm~\cite{rfc793}.
-
-The library is written in ANSI C, and it supports a large number of 32-bits
-embedded processors and a large number of Real-time Operating System (RTOS).
-Hence, a model based on multiple tasks has been adopted to implement the TCP
-protocol as described in the norm, with, in addition, complex cooperation
-mechanisms. A task is dedicated to the user code, another to the timers and the
-last is assigned to the processing of the incoming segments.
-It can also run on bare metal environment.
-
-In the CycloneTCP implementation, the socket data structure is used to retain
-the status of a TCP connection. Furthermore, a socket can be
-manipulated by the user through the TCP's library Application Programming
-Interface (API) to perform various TCP operations, such as the transmission of
-data.
-
-The quality of the C implementation hailed in the AMNESIA:33 report~\cite{AMNESIA33}
-as one of the most resilient TCP/IP stack. This gives a really good assurance
-in the code and it strengthens the confidence that we can have in a formally
-verified code running inside the library environment.
-\todo{Add anything else relevant about the CycloneTCP library}.
+The CycloneTCP implementation adopts the three-task-based model and the socket data structure.
+Thus, one task is dedicated to the user's code, one to the timers, and one to the processing of the incoming segments.
+The user can manipulate a socket through the TCP's library Application Programming Interface (API) to perform various TCP operations, such as data transmission.
 

### 2021-05-18 SecDev21/paper.tex
minor fixes

@@
 
-\def\todo#1{\textcolor{red}{[#1]}}
+\usepackage{todonotes}
+%\def\todo#1{\textcolor{red}{[#1]}}
 
@@
 By using as a starting point such a well-tested and widely used TCP/IP implementation, the IoT industry can have an immediate benefit by adopting this work's security-enhanced version of the CycloneTCP library.
-
+\todo[inline]{pending contributions list and fixing the next paragraph.}
 We present the TCP protocol in section~\ref{sec:TCP} and then briefly the
@@
 
-The CycloneTCP implementation adopts the three-task-based model and the socket data structure.
-Thus, one task is dedicated to the user's code, one to the timers, and one to the processing of the incoming segments.
-The user can manipulate a socket through the TCP's library Application Programming Interface (API) to perform various TCP operations, such as data transmission.
+The CycloneTCP implementation adopts the three-task-based model and the socket data structure defined by the \cite{rfc793}.
+Thus, it has tasks dedicated to the user's code, to the timers, and to the processing of the incoming segments.
+Moreover, the user can manipulate a socket through the TCP's library Application Programming Interface (API) to perform various TCP operations, such as data transmission.
 
@@
 
-SPARK~\cite{mccormick2015building} is the name of a platform that provides
+SPARK~\cite{mccormick_chapin_2015} is the name of a platform that provides
 formal verification for Ada. It uses the user-provided contracts and attempts

### 2021-05-21 SecDev21/paper.tex
fixing ADA and SPARK section

@@
 Given the reliance of the other software-level protocols on the TCP protocol, formalizing the its specification and verifying its implementation is critical.
-Achieving these two goals in practice is challening.
+Achieving these two goals in practice is challenging.
 This is mainly because TCP's specification, \cite{rfc793}, is written in English rather than any formal language, leaving many parts underspecified.
@@
 
-Ada is a general-purpose procedural programming language. The design of the Ada
-language puts great emphasis on the safety and correctness of the program. This
-objective is realized by using a readable syntax that uses keywords instead of
-symbols where reasonable. The type system is strong and strict and many
-potential violations of type constraints can be detected statically by the
-compiler. If not, a run-time check is inserted into the program, to guarantee
-the detection of incorrect situations. Ada 2012 introduced contract based
-programming to Ada. In particular, it is possible to attach pre- and
-postconditions to subprograms\footnote{In Ada, a distinction is made between
-  functions that return a value, and procedures, which do
-  not. \emph{Subprogram} is the term that designates both.}.  These conditions
-can be checked during the execution of the program, just like assertions.
+Ada is a general-purpose, procedural programming language that puts great emphasis on the safety and correctness of the program.
+Some of the safety characteristics of the language are its strong typing system and its extensive compile-time and runtime checks.
+Common C programming vulnerabilities like using unsafe pointers or improper null termination of strings can not exist in an Ada program, while issues like buffer overflow or integer overflow can be dynamically captured by Ada’s runtime checks and dealt with via exception handlers.
+Furthermore, Ada introduced contracts, such as post- and pre-conditions, as part of the language’s standard syntax.
+Such contracts are vital for explicitly expressing software and verification requirements in the source code to allow for static analyzers or runtime checks to verify that the stated requirements are met.
 
-SPARK~\cite{mccormick_chapin_2015} is the name of a platform that provides
-formal verification for Ada. It uses the user-provided contracts and attempts
-to prove that the runtime checks cannot fail and that postconditions are
-established by the corresponding subprograms.  As formal verification for the
-whole Ada language would be intractable, SPARK is also the name of the subset
-of the Ada language that is supported by the SPARK tool called
-GNATprove\footnote{\url{http://docs.adacore.com/spark2014-docs/html/ug/}}.
-This subset contains almost all features of Ada, though sometimes in a
-restricted form.  In particular, expressions should be free from side effects,
-and aliasing is forbidden (no two variables should share the same memory
-location or overlap in memory), including when using pointers thanks to the use
-of an ownership policy~\cite{dross2020recursive}.  This restriction greatly
-simplifies the memory model used in the SPARK tool: any program variables can
-be reasoned about independently from other variables.
+By subsetting Ada, the SPARK programming language \cite{mccormick_chapin_2015} emerged to provide the largest possible subset of Ada, which is suitable for functional specification and static verification.
+For instance, in SPARK, expressions should be free from side effects, and aliasing is forbidden (no two variables should share the exact memory location or overlap in memory), including when using pointers thanks to the use of an ownership policy~\cite{dross2020recursive}.
 
-GNATprove performs two kinds of analysis modularly on individual subprograms:
-information flow to detect reads of uninitialized data and violations of flow
-contracts; deductive verification based on the Why3 platform to generate
-verification conditions for SMT solvers via a weakest-precondition calculus to
-detect possible runtime errors and violations of functional contracts.
+\emph{GNATprove} \cite{GNATProve:users_manual} is the platform that provides formal verification for SPARK through two kinds of analysis.
+The first enables flow analysis to check the initialization of variables, look at data dependencies between inputs and outputs of subprograms, and detects unused assignments and unmodified variables.
+The second uses deductive verification to generate verification conditions for SMT solvers via a weakest-precondition calculus to detect possible runtime errors and violations of functional contracts.
 

### 2021-05-21 SecDev21/paper.tex
Change title and remove reference to code rep for anonymity
fixing C to SPARK section

@@
 By using as a starting point such a well-tested and widely used TCP/IP implementation, the IoT industry can have an immediate benefit by adopting this work's security-enhanced version of the CycloneTCP library.
-\todo[inline]{pending contributions list and fixing the next paragraph.}
+\todo[inline]{KG:pending contributions list and fixing the next paragraph. Will be fixed when the rest of the paper is finalized.}
 We present the TCP protocol in section~\ref{sec:TCP} and then briefly the
@@
 
-To reduce the size of the work, only few functions of interest have been
-translated in SPARK. Most of the C code has been kept.
+The verification of the entire functional specification of the \emph{CycloneTCP} library is beyond the scope of this work.
+Instead, the aim is to demonstrate how the use of SPARK can provide significant security hardening to widely used C-written libraries.
+Thus, this work focuses on the hardening of the TCP library areas that its original authors designated as the most vulnerable or crucial to conform to their functional specifications, namely, the Application Programming Interface (API).
+This mainly falls under two categories: a)  hardening the user's API to enforce its correct usage by the users, discussed in \Cref{sec:API}, and b) proving the conformance to the protocol's functional specification for the translated code, discussed in \Cref{sec:verif}.
+
+\todo[inline]{KG: fixed up to here}
+
 Hopefully, SPARK integrates a mechanism to interface C code. No verification are

### 2021-05-21 SecDev21/paper.tex

@@
 
-\title{Verification of a TCP/IP stack}
+\title{Layered Formal Verification of an Industrial TCP/IP Stack}
 
@@
 By using as a starting point such a well-tested and widely used TCP/IP implementation, the IoT industry can have an immediate benefit by adopting this work's security-enhanced version of the CycloneTCP library.
+
 \todo[inline]{KG:pending contributions list and fixing the next paragraph. Will be fixed when the rest of the paper is finalized.}
+
 We present the TCP protocol in section~\ref{sec:TCP} and then briefly the
@@
 
-The reader can refer to the code at \url{https://github.com/AdaCore/Http_Cyclone}.
+% The reader can refer to the code at \url{https://github.com/AdaCore/Http_Cyclone}.
 

### 2021-05-21 SecDev21/paper.tex
cleanup intro

@@
 Any vulnerability in software libraries that are widely used by embedded systems can potentially open the door for compromising the security of billions of deployed IoT devices.
-In 2019, critical security vulnerabilities were discovered in a TCP implementation that has been used potentially by billions of IoT devices for more than 13 years \cite{zero_days_vuln:ARMIS_white_paper}, with some of them even allowing remote code execution.
+In 2019, critical security vulnerabilities were discovered in a TCP implementation that has been used by billions of IoT devices for more than 13 years~\cite{zero_days_vuln:ARMIS_white_paper}, some of them allowing remote code execution.
 
@@
 
-Existing research deals with formally verifying several protocols TCP/IP stack levels.
-For example, the work in \emph{miTLS} \cite{bhargavan2013implementing} formally verifies an SSL/TLS protocol implementation, and the work in \cite{Reiher_2020} uses a technology called RecordFlux to safely parse data segments.
-Given the reliance of the other software-level protocols on the TCP protocol, formalizing the its specification and verifying its implementation is critical.
-Achieving these two goals in practice is challenging.
-This is mainly because TCP's specification, \cite{rfc793}, is written in English rather than any formal language, leaving many parts underspecified.
-Moreover, the TCP protocol is inherently concurrent, making it complex to specify and verify.
-Previous work focused on formalizing the protocol to either prove key properties \cite{smith1996formal} or fully verify the TCP protocol's specifications \cite{ridge2008rigorous} using the HOL proof assistant tool.
-While for the later work, the authors claim that it is possible from their formalization to extract an actual implementation in Haskell, this has not been achieved yet.
-At this point of writing, we are not aware of any formally verified TCP protocol implementation.
+Existing research deals with formally verifying several protocols of TCP/IP stack levels.
+Early work formalized the TCP protocol to prove some key properties of the protocol itself~\cite{smith1996formal}.
+More recently, researchers have formally specified TCP, UDP and the Sockets API in HOL~\cite{tcp2005sigcomm,ridge2008rigorous} and used the resulting specification for testing existing libraries.
+Others have formally specified and implemented the SSL/TLS protocol in F*\cite{bhargavan2013implementing}, and integrated the resulting implementation in existing libraries.
+To the best of our knowledge, no prior work attempted to formally verify an implementation of TCP, although the authors of~\cite{ridge2008rigorous} suggest that it would possible to extract an actual implementation in Haskell from their specification.
 
-In this paper, we take a practical approach, where the security of an existing TCP/IP stack implementation written in ANSI C programming language is incrementally enhanced by replacing parts of its code with formally verified code in SPARK \cite{mccormick_chapin_2015}.
-SPARK 2014 is a programming language designed as a subset of the Ada language to produce highly reliable software through formal verification.
-For this work, the CycloneTCP, a professional-grade embedded TCP/IP library developed by the Oryx Embedded company \cite{CycloneTCP}.
+In this paper, we take a practical approach, where the security of an existing TCP/IP stack implementation written in ANSI C programming language is incrementally enhanced by replacing parts of its code with formally verified code in SPARK~\cite{mccormick_chapin_2015}, a subset of the Ada programming language supported by formal verification tools.
+We chose CycloneTCP, a professional-grade embedded TCP/IP library developed by the Oryx Embedded company~\cite{CycloneTCP}.
 \Cref{Fig:TcpStack} shows the protocols supported by the library.
@@
 The quality of the CycloneTCP is acknowledged by the AMNESIA:33 report \cite{AMNESIA33}, which classifies it as one of the most resilient TCP/IP stack.
-By using as a starting point such a well-tested and widely used TCP/IP implementation, the IoT industry can have an immediate benefit by adopting this work's security-enhanced version of the CycloneTCP library.
 

### 2021-05-21 SecDev21/paper.tex
interfacing data objects text fixes

--- 2021-05-21 SecDev21/paper.tex
@@
 
+Ada and SPARK offer an easy-to-use mechanism for interfacing with C code \cite{}.
+The mechanism allows C-written function to be called by Ada/SPARK code and vice versa and for representing shared data objects.
+The memory representations of a shared data object must be the same in both the C and the SPARK code.
+For instance, if a C function is called with a 32-bits integer argument, the SPARK interface of this function should also be called with a 32-bits integer argument.
+It is the responsibility of the user to ensure this.
+
 \todo[inline]{KG: fixed up to here}
 
-Hopefully, SPARK integrates a mechanism to interface C code. No verification are
+No verification are
 performed on the C code, but pre- and postconditions can be attached to
@@
 programmer to ensure the correctness of the contracts.
-
-The types in SPARK can be more expressive than the types in C which adds more
-safety to the SPARK code.
-But the C and the SPARK code can share objects and it is essential that the
-memory representations of an object are the same in the C and in the SPARK side.
-For instance, if a C function is supposed to be called with a 32-bits integer,
-it is the responsibility of the programmer to ensure that the function is
-actually called with a 32-bits integer in the SPARK code.
 

### 2021-05-21 SecDev21/paper.tex
fixing text for  contracts on C interfaced functions

@@
 
+Although formal verification can only be applied to the SPARK translated code,  pre- and post-conditions can be attached to C functions interfaced in SPARK to allow reasoning about the side effects of calling these functions by the SPARK code.
+Since the functional specification of these contracts can not be proved, it is the role of the programmer to ensure their correctnes.
+
 \todo[inline]{KG: fixed up to here}
-
-No verification are
-performed on the C code, but pre- and postconditions can be attached to
-subprograms and they can be called in SPARK code.
-The functional specification are not proved, thus it is the role of the
-programmer to ensure the correctness of the contracts.
 

### 2021-05-21 SecDev21/paper.tex
fixing typos
more Enlgish fixes

@@
 Although formal verification can only be applied to the SPARK translated code,  pre- and post-conditions can be attached to C functions interfaced in SPARK to allow reasoning about the side effects of calling these functions by the SPARK code.
-Since the functional specification of these contracts can not be proved, it is the role of the programmer to ensure their correctnes.
-
-\todo[inline]{KG: fixed up to here}
+Since the functional specification of these contracts can not be proved, it is the programmer's role to ensure their correctness.
+Interfacing C and SPARK code can be error-prone if it is not done carefully.
+\Cref{sec:verif} discusses how to mitigate this issue effectively.
 
@@
 problems to ensure the quality of our work.
+
+\todo[inline]{KG: fixed up to here}
 

### 2021-05-21 SecDev21/paper.tex
removing old text
further cleanup

@@
 TCP is a connection-oriented protocol; a connection between the sender and the receiver must be established before transmitting any data.
-Furthermore, TCP is a reliable protocol since it guarantees the delivery of all the messages and that they are delivered in the order, they were sent.
-It also provides error-checking mechanisms to discard and recover any corrupted data.
+Furthermore, TCP is a reliable protocol since it guarantees the delivery of all the messages and that they are delivered in the order they were sent.
+It also provides error-checking mechanisms to discard and recover from corrupted data.
 
@@
 Generally, every TCP communication session will go through the following three phases (if no error occurs): 1)~opening the connection, 2)~sending and receiving the data, and 3)~closing the connection.
-The behaviour of the three-phase communication can be described by the state machine given in \cref{Fig:statemachine}.
+The behavior of the three-phase communication can be described by the state machine given in \cref{Fig:statemachine}.
 An edge represents a state transition.
-A transition is either triggered by an explicit action, stated in italics format on the label's edge, for example, \textit{Close}, or is triggered by the arrival of a specific flag/s.
-An edge-label in the format of $\frac{\ \ x \ \ }{\ \ y \ \ }$ represents the associated flags transmitted with each transition, where $x$ is either send or receive and $y$ refers to the actual flag/s transmitted in response to the action that caused the state transition.
-An explicit action is either a user-triggered action (through the user's API) or an automatically performed action triggered by a timeout event.
+A transition is either triggered by an explicit action, stated in italics format on the label's edge, for example, \textit{Close}, or is triggered by the arrival of specific flags.
+An edge-label in the format of $\frac{\ \ x \ \ }{\ \ y \ \ }$ represents the associated flags transmitted with each transition, where $x$ is either send or receive and $y$ refers to the actual flags transmitted in response to the action that caused the state transition.
+An explicit action is either a user-triggered action (through the user's API) or an action automatically triggered by a timeout event.
 The flags are embedded in the header section of a transmitted segment and can be one of the following:
@@
 \Cref{sec:verif} discusses how to mitigate this issue effectively.
-
-Interfacing C and SPARK code is a great opportunity to plug our work in an
-existing library, but it is also be a source of errors if it is not done
-carefully. In section~\ref{sec:verif} it is explain how we have dealt with these
-problems to ensure the quality of our work.
 

### 2021-05-21 SecDev21/paper.tex
describing things better

--- 2021-05-21 SecDev21/paper.tex
@@
 
-Ada and SPARK offer an easy-to-use mechanism for interfacing with C code \cite{}.
-The mechanism allows C-written function to be called by Ada/SPARK code and vice versa and for representing shared data objects.
+Ada and SPARK offer an easy-to-use mechanism for interfacing with C code.
+The mechanism allows  C-written functions to be called by Ada/SPARK code and vice versa, via function bindings \cite{Ada_C_Binding}, and for the representation of shared data objects by the two languages.
 The memory representations of a shared data object must be the same in both the C and the SPARK code.
-For instance, if a C function is called with a 32-bits integer argument, the SPARK interface of this function should also be called with a 32-bits integer argument.
+For instance, if a C function is called with a 32-bits integer argument, the SPARK binding of this function should also be called with a 32-bits integer argument.
 It is the responsibility of the user to ensure this.
 
-Although formal verification can only be applied to the SPARK translated code,  pre- and post-conditions can be attached to C functions interfaced in SPARK to allow reasoning about the side effects of calling these functions by the SPARK code.
+Although formal verification can only be applied to the SPARK translated code, pre- and post-conditions can be attached to C functions bindings in SPARK to allow the formal verification tools to reason about side effects of calling these functions that can affect the SPARK code.
 Since the functional specification of these contracts can not be proved, it is the programmer's role to ensure their correctness.

### 2021-05-24 SecDev21/paper.tex
more fixes
rephrasing

--- 2021-05-21 SecDev21/paper.tex
@@
 Any vulnerability in software libraries that are widely used by embedded systems can potentially open the door for compromising the security of billions of deployed IoT devices.
-In 2019, critical security vulnerabilities were discovered in a TCP implementation that has been used by billions of IoT devices for more than 13 years~\cite{zero_days_vuln:ARMIS_white_paper}, some of them allowing remote code execution.
+Researchers have shown that such vulnerabilities exist: in 2019, critical security vulnerabilities were discovered in a TCP implementation that has been used by billions of IoT devices for more than 13 years~\cite{zero_days_vuln:ARMIS_white_paper}, some of them allowing remote code execution.
 
@@
 
-Existing research deals with formally verifying several protocols of TCP/IP stack levels.
+Existing research has dealt with formally verifying several protocols of TCP/IP stack levels.
 Early work formalized the TCP protocol to prove some key properties of the protocol itself~\cite{smith1996formal}.
@@
 In this paper, we take a practical approach, where the security of an existing TCP/IP stack implementation written in ANSI C programming language is incrementally enhanced by replacing parts of its code with formally verified code in SPARK~\cite{mccormick_chapin_2015}, a subset of the Ada programming language supported by formal verification tools.
-We chose CycloneTCP, a professional-grade embedded TCP/IP library developed by the Oryx Embedded company~\cite{CycloneTCP}.
+We chose CycloneTCP, a professional-grade open source embedded TCP/IP library developed by the Oryx Embedded company~\cite{CycloneTCP}.
 \Cref{Fig:TcpStack} shows the protocols supported by the library.
-This work focuses on the TCP and socket components of the stack.
-The library's TCP implementation is meant to conform with the RFC 793 protocol specifications \cite{rfc793}.
+This work focuses on the TCP and Socket components of the stack.
+The library's TCP implementation is meant to conform with the RFC 793 protocol specifications~\cite{rfc793}.
 The quality of the CycloneTCP is acknowledged by the AMNESIA:33 report \cite{AMNESIA33}, which classifies it as one of the most resilient TCP/IP stack.
@@
 An edge represents a state transition.
-A transition is either triggered by an explicit action, stated in italics format on the label's edge, for example, \textit{Close}, or is triggered by the arrival of specific flags.
-An edge-label in the format of $\frac{\ \ x \ \ }{\ \ y \ \ }$ represents the associated flags transmitted with each transition, where $x$ is either send or receive and $y$ refers to the actual flags transmitted in response to the action that caused the state transition.
+An edge-label in the format of $\frac{\ \ x \ \ }{\ \ y \ \ }$ represents the actions associated to the state transition, where $x$ is triggering the transition, and $y$ is produced by the transition.
+The trigger $x$ is either an explicit action like \textit{Close}, or the arrival of specific flags, while $y$ always represents the flags transmitted in response to the action that caused the state transition, when present.
 An explicit action is either a user-triggered action (through the user's API) or an action automatically triggered by a timeout event.
@@
 \ssynsent and \ssynrcv represent the initialization of the connection that ends when \sestab is reached.
-\sfwone, \sfwtwo, \sclosing, and \stimewait are a group of states that represents waiting for a connection termination request from remote TCP.
+\sfwone, \sfwtwo, \sclosing, and \stimewait are a group of states that represent waiting for a connection termination request from the remote TCP.
 \sclosew and \slastack represent waiting for a connection termination request from the local user.
@@
 Different tasks can interact and update the socket data structure, used to retain the status of a TCP connection, to handle the various events that can occur within a TCP session.
-The TCP norm, \cite{rfc793} , under Section "3.9.
-Event Processing" page 52, describes a possible implementation of how to handle these events based on three tasks: one for the \emph{user calls}, one for the \emph{arriving segments} and one for the \emph{timers}.
-The role of each of the three tasks can be summarized as follow: User calls refer to functions, namely \texttt{open}, \texttt{close}, \texttt{abort}, \texttt{send} and \texttt{receive}, that can be called by the user to control the connection, send, or receive data.
+RFC 793, under Section ``3.9.
+Event Processing'' page 52, describes a possible implementation of how to handle these events based on three tasks: one for the \emph{user calls}, one for the \emph{arriving segments} and one for the \emph{timers}.
+The role of each of the three tasks can be summarized as follows.
+User calls refer to functions, namely \texttt{open}, \texttt{close}, \texttt{abort}, \texttt{send} and \texttt{receive}, that can be called by the user to control the connection, send, or receive data.
 These functions can trigger transitions between the connection's states since they are intended to control the connection.
@@
 Thus, corresponding transitions to the timeout events can also be triggered by this task.
-A complete description of the protocol's specification can be found at the \emph{RFC 793} document \cite{rfc793}.
+A complete description of the protocol's specification can be found in the RFC 793 document~\cite{rfc793}.
 
-The CycloneTCP implementation adopts the three-task-based model and the socket data structure defined by the \cite{rfc793}.
-Thus, it has tasks dedicated to the user's code, to the timers, and to the processing of the incoming segments.
+The CycloneTCP implementation adopts the three-task-based model and the socket data structure defined by the RFC 793.
+Thus, it has tasks dedicated to the user's code, to the timers, and to the processing of incoming segments.
 Moreover, the user can manipulate a socket through the TCP's library Application Programming Interface (API) to perform various TCP operations, such as data transmission.
@@
 Some of the safety characteristics of the language are its strong typing system and its extensive compile-time and runtime checks.
-Common C programming vulnerabilities like using unsafe pointers or improper null termination of strings can not exist in an Ada program, while issues like buffer overflow or integer overflow can be dynamically captured by Ada’s runtime checks and dealt with via exception handlers.
-Furthermore, Ada introduced contracts, such as post- and pre-conditions, as part of the language’s standard syntax.
+Common C programming vulnerabilities like using unsafe pointers or improper null termination of strings cannot exist in an Ada program, while issues like buffer overflow or integer overflow can be dynamically captured by Ada’s runtime checks and dealt with via exception handlers.
+Furthermore, Ada introduced contracts, such as preconditions and postconditions, as part of the language’s standard syntax.
 Such contracts are vital for explicitly expressing software and verification requirements in the source code to allow for static analyzers or runtime checks to verify that the stated requirements are met.
@@
 
-\emph{GNATprove} \cite{GNATProve:users_manual} is the platform that provides formal verification for SPARK through two kinds of analysis.
+\emph{GNATprove} \cite{GNATProve:users_manual} is the tool that provides formal verification for SPARK through two kinds of analyses.
 The first enables flow analysis to check the initialization of variables, look at data dependencies between inputs and outputs of subprograms, and detects unused assignments and unmodified variables.
@@
 The verification of the entire functional specification of the \emph{CycloneTCP} library is beyond the scope of this work.
-Instead, the aim is to demonstrate how the use of SPARK can provide significant security hardening to widely used C-written libraries.
+Instead, the aim is to provide significant security hardening to mostly-C libraries by selective use of SPARK.
 Thus, this work focuses on the hardening of the TCP library areas that its original authors designated as the most vulnerable or crucial to conform to their functional specifications, namely, the Application Programming Interface (API).
-This mainly falls under two categories: a)  hardening the user's API to enforce its correct usage by the users, discussed in \Cref{sec:API}, and b) proving the conformance to the protocol's functional specification for the translated code, discussed in \Cref{sec:verif}.
+This mainly falls under two categories: a) proving the conformance to the protocol's functional specification for the translated code, discussed in \Cref{sec:verif}, and b) hardening the user's API to enforce its correct usage by users, discussed in \Cref{sec:API}.
 
@@
 
-Although formal verification can only be applied to the SPARK translated code,  pre- and post-conditions can be attached to C functions interfaced in SPARK to allow reasoning about the side effects of calling these functions by the SPARK code.
-Since the functional specification of these contracts can not be proved, it is the programmer's role to ensure their correctness.
+Although formal verification can only be applied to the SPARK translated code,  preconditions and postconditions can be attached to C functions interfaced in SPARK to allow reasoning about the side effects of calling these functions by the SPARK code.
+Since the functional specification of these contracts cannot be proved, it is the programmer's role to ensure their correctness.
 Interfacing C and SPARK code can be error-prone if it is not done carefully.
@@
 
-In CycloneTCP library, pointers are often manipulated at top-level and transmitted
-to C functions. They are also used as fields in structures, for instance in
-the socket structure to store other complex data-structures and recursive pointer
-based data-structures. SPARK use a pointer ownership model to prevent bugs
-involving pointers such as memory leak or double free. When a memory area is
-allocated, a pointer points to this memory area and thus it is owned.
+In the CycloneTCP library, pointers are used both as the type of variables, to pass objects when calling other functions, and the type of structure fields, for instance in
+the socket structure to store other complex data-structures and recursive pointer-based data-structures. SPARK use a pointer ownership model to prevent bugs
+involving pointers such as memory leak or double free, which imposes some restrictions on how pointers are manipulated.
+When a memory area is allocated, a pointer points to this memory area and thus it is owned.
 The ownership of the memory can be moved but the memory must be deallocated
@@
 
-The socket structure is shared between different functions, and contains a lot
-of fields. Most of the fields are used in other parts of library and are not
-relevant for our verification. We only need to concentrate on a subset of fields
-to make the specifications easier to write, and to understand.
-This can be achieved in SPARK by using \emph{models} and \emph{ghost code}
-which are used by GNATprove for analysis but not compiled in the final executable.
+The socket structure is shared between different functions, and contains lots
+of fields. Most of the fields are used in other parts of the library and are not
+relevant for our verification. Thus we use \emph{ghost code} in SPARK
+to focus the verification effort on the relevant subset of the fields,
+which also makes the specifications easier to write and to understand.
+This is a solution to the classical problem of specifying the \emph{frame condition} for a function, that is, the set of possibly modified objects.
+Ghost code is used by GNATprove for analysis but not compiled in the final executable.
 
-In the contract, we use the construction \spark{Model(Sock)} to refer the
-relevant fields of the object \spark{Sock} where
-\spark{Model} is a ghost function that extract the fields of interest from
-\spark{Sock}, like in the construction frequently used in postconditions
+More specifically, we define a \emph{model} of the actual socket as a ghost function for use in contracts: \spark{Model(Sock)}
+extracts the fields of interest from socket \spark{Sock}.
+A frequent usage in postconditions is
 \begin{lstlisting}[language=Ada, basicstyle=\small\ttfamily]
@@
 that states that relevant fields of \spark{Sock} selected by \spark{Model} are
-not changed by the function call, except the fields \spark{State} that is equal
+not changed by the function call, except for the field \spark{State} that is equal
 to \spark{TCP_STATE_CLOSED} after the call.
@@
 
-\subsection{Extracting a specification from the TCP norm}
+\subsection{Extracting a specification from RFC 793}
 
 Extracting a specification
-is mandatory to provide functional contracts to the user functions. That can
-be achieved by reading the norm. However, the norm is underspecified in many
-cases and only one possible implementation is depicted. Thus, the specifications
-we have extracted are conformant with the norm but they slightly differ to what
-can be found in the RFC.
-Among the specification we have extracted:
+is mandatory for provide functional contracts to the user functions. That can
+be achieved by reading the text of RFC 793. However, the text is underspecified in many
+cases and only one possible implementation is described. We made the following choices in our formalization of the TCP specification:
 \begin{itemize}
-\item The transitions between the states must respect the order given by the
+\item The transitions between states must respect the order given by the
 state machine described in figure~\ref{Fig:statemachine}.
@@
 these two states.
-\item The user functions contain functional specifications in the norm, that
-have to be done  when a function is called depending on the state of the
-socket~\cite[p. 52]{rfc793}. In particular it is specified in what state the
-function can be called without returning an error.
+\item RFC 793 describes conditions on the state of the socket for calling user functions,
+which need to hold when the function is called~\cite[p. 52]{rfc793}. In particular it is specified in what state the
+function can be called without returning an error. These became preconditions in our formalization.
 \end{itemize}
@@
 
-The user functions of the TCP protocol have been rewritten in SPARK. Besides
-proving the absence of run-time error in these functions, it has been possible
-to check that the functional specifications of the TCP protocol are not violated
-by the code.
+The user functions defining the API of the TCP protocol were fully rewritten in SPARK.
+SPARK bindings were also provided to the majority of the rest of the library’s C functions to allow their invocation from the SPARK code,
+taking into account their effect on the TCP state machine in their contracts.
+GNATprove was used to prove that the implementation of user functions is free of run-time errors and
+respects the functional specifications of the TCP protocol that we formalized in contracts.
 
@@
 
-When a transition has to be made in the code, it is made through the helper
-function \spark{TCP_Change_State}. The precondition of this function describes
-all the allowed transitions. Therefore, when we try to do a forbidden transition,
-GNATprove will display a message. This is the most direct way to make a
-transition in the state machine.
-As the TCP user functions are all protected with a mutex to avoid interferences
-from other threads that could change the TCP state during their execution, in
-particular when a new segment is received, it is quite easy to know the
-possible set of state that can have the socket during the all execution of the
-functions.
+To model transitions of the TCP state machine inside the code, we defined the helper
+function \spark{TCP_Change_State}, whose contract describes
+all the transitions allowed.
+As calls to user functions are protected with a mutex to avoid interferences
+from other threads, the only changes to the TCP state machine inside user functions are
+related to their actions. We model these changes by calling \spark{TCP_Change_State}
+at places that correspond to a transition of the TCP machine.
 
-There is only one function that makes the verification harder because it
-releases the mutex: the function \spark{TCP_Wait_For_Events}.
-It takes as argument an event to wait and returns once the event
-is triggered by the reception of a segment.
-The role of this
-function is to release the mutex and allow the reception task to process the
-incoming segments. This function
-can be used, for example, in the function \spark{open} when the connection
-is initialized to wait until the connection has been established or not to
+% Yannick: I don't understand the following, needs to be rephrased if we want to keep it
+%
+% particular when a new segment is received, it is quite easy to know the
+% possible set of state that can have the socket during the all execution of the
+% functions.
+
+The function \spark{TCP_Wait_For_Events} requires special handling because it may
+release the mutex. This function takes in argument an event on which it should wait,
+and returns once the event is triggered by the reception of a segment.
+While waiting, it releases the mutex to allow the reception task to process incoming segments.
+For example, this function is called from \spark{open} after the connection
+has been initialized, to wait until the connection has been established or not, in order to
 correctly set the error that indicates to the user the state of the connection.
-If we consider that everything can happen during a call to \spark{TCP_Wait_For_Events},
-the verification becomes too imprecise and it becomes impossible to prove the
-functional specifications extracted from the norm. Instead we have decided to
-find a better contract to this function.
+
+% Yannick: we talk about error above, but there is no error if the connection is established???
+% This needs rephrasing
+
+A correct but imprecise contract for \spark{TCP_Wait_For_Events} would allow any change in the TCP
+state machine while waiting. But it would make it impossible to prove that its callers
+satisfy their functional specification. We need a more precise contract for this function.
 
@@
 
+% Yannick: below "a new segement...if it contains a new segment" is not phrased correctly
+
 The function \spark{TCP_Wait_For_Events} first checks if the event is not already
 true. If not, the mutex is released until a segment that triggered the event is
-received or until the timeout is reach. Each time a new segment is received, it
+received or until the timeout is reached. Each time a new segment is received, it
 is processed and if it contains a segment that causes a change of state, the TCP
@@
 true due to the change of state, it is triggered.
+
+% Yannick: the above final sentence is unclear: what is triggered?
 
@@
 transitions in the TCP state machine, but some other properties are formulated
-in the TCP norm, and could be verified in the implementation to make it even
+in RFC 793, and could be verified in the implementation to make it even
 more robust.

### 2021-05-24 SecDev21/paper.tex
more rephrasing

@@
 
+% Yannick: the caption below used the term "user-task related function" which is used nowhere else
+
 \begin{algorithm}[t]
@@
 
+% Yannick: the whole explanation is unclear. We should rephrase to explain that:
+% - a suitable contract is added to function \spark{TCP_Wait_For_Events}
+% - to make sure this contract is correct, we implement a ghost procedure
+%   \spark{TCP_Wait_For_Events_Proof} as shown in Algorithm 1, that models
+%   the possible chain of events that occur while waiting, with the same contract
+%   as \spark{TCP_Wait_For_Events}. This procedure is proved automatically
+%   without even a loop invariant thanks to loop unrolling in GNATprove.
+
 An event is considered to be true or not depending on the state of the socket.
@@
 
-Rewriting the part that process the incoming segments was out of the scope of
+Rewriting the part that processes the incoming segments was out of the scope of
 this work. Instead, we have opted for the use of symbolic execution on the
@@
 execution to improve deductive verification, and in~\cite{kassios2012comparing}
-the authors explain that symbolic execution can be good at infering contracts.
+the authors explain that symbolic execution can be good at inferring contracts.
 For our work, KLEE has been used. It is built on top of LLVM as it takes LLVM
@@
 the subtleties of CycloneTCP. The next step is to write a driver to run KLEE and
-know if the contracts that we have infered are correct or not.
+know if the contracts that we have inferred are correct or not.
 A simplified example is shown in figure~\ref{code:kleedriver}. The first lines
@@
 segment is called. It is followed at line~\ref{code:kleedriver:assert} by a KLEE
-builtin that check if an assert is valid for all the symbolic paths
+builtin that checks if an assert is valid for all the symbolic paths
 explored before.
@@
 
-It appears that lot of programmers do not correctly use the socket API that is
-provided by the library. Providing verified user functions for TCP is not enough
-to guarantee the safety of software if the users do not correctly use them.
-An important part of the work was to ensure that code that uses the socket API's
-functions are correctly called. By ``correctly'' we mean: 1)~the functions
-are called with respect to a defined order and 2)~the return code is checked
-before continuing the processing.
+A common problem faced by users of the CycloneTCP library is the need to
+properly chain calls of the Socket API, in a way that respects the TCP
+automaton. This entails verifying that 1)~the functions are called with respect
+to a defined order and 2)~the return code is checked before continuing the
+processing.
 
@@
 figure~\ref{Fig:statemachine}.
-Ada's pre- and postconditions are a powerful tool to express such inter-function
+Ada's preconditions and postconditions are a powerful tool to express such inter-function
 dependencies, while SPARK technology can be used to guarantee that the
-assertions always hold. Thus, pre- and postconditions were introduced to model
+assertions always hold. Thus, preconditions and postconditions were introduced to model
 a partial order on the calls to the API's functions.
@@
 section~\ref{sec:pointers}, we have found that a memory leak can happen when
-the connection have been closed and the buffer are cleaning.
+the connection has been closed and the buffer are cleaning.
 
@@
 
-Another very interesting and very subtle bug have been found thanks to SPARK is
-reproduced in the snippet bellow.
+A subtle bug found thanks to SPARK is
+reproduced in the snippet below.
 
@@
 In this scenario, after the call to \spark{Tcp_Send}, when the error cases have
-been filter, the state of the socket can either be \sestab or \sclosew. The call
-at line~\ref{bugProg:tcpWaitForEvents} and now the state of the socket can be
-still be \sestab or \sclosew, or if a \rst flag has been received when the mutex
-was released, the state can be \sclosed too.
+been filtered, the state of the socket can either be \sestab or \sclosew. After the call
+at line~\ref{bugProg:tcpWaitForEvents}, the state of the socket can
+still be \sestab or \sclosew, or it can be \sclosed if a \rst flag has been received when the mutex
+was released.
 At line~\ref{bugProg:tcpWaitForEventsPost}, a \fin segment is sent, which does
@@
 
-But these transition are not allowed by the norm, and thus not allowed by the
+But these transitions are not allowed by RFC 793, and thus not allowed by the
 precondition of \spark{TCP_Change_State}. The problem is detected by SPARK and a fix
-has been proposed.
+has been applied to the SPARK and C implementations of CycloneTCP.
+
+% The above is uncler, the transitions from CLOSED and CLOSED-WAIT are allowed, it's only the transition from ESTABLISHED which is not allowed.
 
@@
 
+% above: need time to prove automatically with GNATprove with the spec of your laptop (num of cores, amount of memory)
+
 \section{Conclusion and future work}
@@
 results. It has helped to find bugs in the existing implementation. But more
-than finding bugs, the interest of this work lie in the face that it is a big
+than finding bugs, the interest of this work lie in the face that it is a
 step toward a secure implementation. It has been proved by the tool GNATprove
@@
 
+% examples of such properties?
+
+% ambition for industrial use?
+
 \bibliographystyle{IEEEtran}

### 2021-05-26 SecDev21/paper.tex
some more fixes

@@
 
-\todo[inline]{KG: fixed up to here}
-
 \subsection{Dealing with pointers}
@@
 
-In the CycloneTCP library, pointers are used both as the type of variables, to pass objects when calling other functions, and the type of structure fields, for instance in
-the socket structure to store other complex data-structures and recursive pointer-based data-structures. SPARK use a pointer ownership model to prevent bugs
-involving pointers such as memory leak or double free, which imposes some restrictions on how pointers are manipulated.
-When a memory area is allocated, a pointer points to this memory area and thus it is owned.
-The ownership of the memory can be moved but the memory must be deallocated
-before the end of the execution of the program.
+In the CycloneTCP library, pointers are used both as the type of variables to pass objects when calling other functions and as the type of structure fields; for instance, in the socket structure to store other complex data structures and recursive pointer-based data structures.
+SPARK uses a pointer ownership model to prevent bugs involving pointers such as memory leak or double free, which imposes some restrictions on how pointers are being manipulated.
+When a memory area is allocated, a pointer points to this memory area, and thus, it is considered "owned.'' The ownership of the memory can be moved, but the memory must be deallocated before the end of the execution of the program.
 
@@
 
-The socket structure is shared between different functions, and contains lots
-of fields. Most of the fields are used in other parts of the library and are not
-relevant for our verification. Thus we use \emph{ghost code} in SPARK
-to focus the verification effort on the relevant subset of the fields,
-which also makes the specifications easier to write and to understand.
+The socket structure is shared between different functions and contains a large number of fields.
+Most of the fields are used in library parts that are not considered by this work, and thus, are not relevant for our verification.
+Thus we use \emph{ghost code} in SPARK to focus the verification effort on the relevant subset of the fields, which also makes the specifications easier to write and understand.
 This is a solution to the classical problem of specifying the \emph{frame condition} for a function, that is, the set of possibly modified objects.
-Ghost code is used by GNATprove for analysis but not compiled in the final executable.
+\emph{Ghost code} is used by GNATprove for analysis but is not being compiled in the final executable.
+More specifically, we define a \emph{model} of the actual socket as a ghost function that can be used in contracts: \spark{Model(Sock)} extracts the fields of interest from socket \spark{Sock}.
+A frequent usage in postconditions is
 
-More specifically, we define a \emph{model} of the actual socket as a ghost function for use in contracts: \spark{Model(Sock)}
-extracts the fields of interest from socket \spark{Sock}.
-A frequent usage in postconditions is
 \begin{lstlisting}[language=Ada, basicstyle=\small\ttfamily]
@@
 \end{lstlisting}
-that states that relevant fields of \spark{Sock} selected by \spark{Model} are
-not changed by the function call, except for the field \spark{State} that is equal
-to \spark{TCP_STATE_CLOSED} after the call.
+
+This states that the relevant fields of \spark{Sock} selected by \spark{Model} are not changed by the function call, except for the field \spark{State} that is equal to \spark{TCP_STATE_CLOSED} after the call.
+
+\todo[inline]{KG: fixed up to here}
 

### 2021-05-26 SecDev21/paper.tex
more text fixes

@@
 
-\todo[inline]{KG: fixed up to here}
-
 \section{Conformance to the TCP protocol}
@@
 
-Extracting a specification
-is mandatory for provide functional contracts to the user functions. That can
-be achieved by reading the text of RFC 793. However, the text is underspecified in many
-cases and only one possible implementation is described. We made the following choices in our formalization of the TCP specification:
+Extracting a specification is mandatory for providing functional contracts to the user's functions.
+This can be achieved by interpreting the text of RFC 793.
+However, the text is underspecified, and thus, only one of the possible implementations is used by this work.
+The following choices were made in our formalization of the TCP specification:
+
 \begin{itemize}
 \item The transitions between states must respect the order given by the
-state machine described in figure~\ref{Fig:statemachine}.
-In particular, a transition from a state to
-another is valid only if there exists a transition in the state machine between
-these two states.
-\item RFC 793 describes conditions on the state of the socket for calling user functions,
-which need to hold when the function is called~\cite[p. 52]{rfc793}. In particular it is specified in what state the
-function can be called without returning an error. These became preconditions in our formalization.
+state machine described in figure~\ref{Fig:statemachine}. In particular, a transition from a state to another is valid only if the state machine allows for a transition between these two states.
+\item RFC 793 describes the conditions for the state of the socket which need to hold when a user's function is called~\cite[p.
+52]{rfc793}.
+In particular, these conditions specify in what state the function can be called without returning an error.
+These conditions were turned into preconditions in our formalization.
 \end{itemize}
+
+\todo[inline]{KG: fixed up to here}
 

### 2021-05-26 SecDev21/paper.tex
rework the TCP user function section

@@
 The library's TCP implementation is meant to conform with the RFC 793 protocol specifications~\cite{rfc793}.
-The quality of the CycloneTCP is acknowledged by the AMNESIA:33 report \cite{AMNESIA33}, which classifies it as one of the most resilient TCP/IP stack.
+The quality of the CycloneTCP is acknowledged by the AMNESIA:33 report~\cite{AMNESIA33}, which classifies it as one of the most resilient TCP/IP stack.
 
@@
 
-The user functions defining the API of the TCP protocol were fully rewritten in SPARK.
-SPARK bindings were also provided to the majority of the rest of the library’s C functions to allow their invocation from the SPARK code,
-taking into account their effect on the TCP state machine in their contracts.
-GNATprove was used to prove that the implementation of user functions is free of run-time errors and
-respects the functional specifications of the TCP protocol that we formalized in contracts.
+The original \emph{CycloneTCP} library provides no guarantees that the implementation respects the TCP functional specification.
+Thus, there is no assurance that the implementation will only allow valid transitions between the different states that a TCP session can exhibit.
+We aim to use the SPARK technology to verify that the implementation respects the TCP automaton, given in \cref{Fig:statemachine}.
+Although, this work is focused on hardening the API of the user task, mainly the high-level user functions, to verify the state transitions the rest of the TCP protocol has to be taken into account.
+The reason for this is two-fold.
+First, although user-functions can trigger state transitions, the actual transitions are done by library functions that do not belong to the user-task set of functions.
+Second, other parts of the library can trigger state transitions during and between the user-task function calls, which can affect the intermediate and final states that a user related function can exhibit.
+
+The TCP user API related functions were fully translated to SPARK.
+SPARK bindings were also provided to the majority of the rest of the library's C functions to allow their invocation from the SPARK code.
+
+RFC 793 document provides all the information needed for the allowed transitions between the different TCP possible states.
+For any transitions that are directly triggered through the call of the user-tasks functions, SPARK contracts, that represent transitions allowed by the functional specification, are embedded within the SPARK translated code.
+The \texttt{TCP\_Change\_State} is a helper function that is called every time a state transition needs to be made, to update the socket's state.
+An incorrect transition allowed by the code will be detected by the prover.
 
@@
 
-To model transitions of the TCP state machine inside the code, we defined the helper
-function \spark{TCP_Change_State}, whose contract describes
-all the transitions allowed.
-As calls to user functions are protected with a mutex to avoid interferences
-from other threads, the only changes to the TCP state machine inside user functions are
-related to their actions. We model these changes by calling \spark{TCP_Change_State}
-at places that correspond to a transition of the TCP machine.
-
 % Yannick: I don't understand the following, needs to be rephrased if we want to keep it
@@
 
-The function \spark{TCP_Wait_For_Events} requires special handling because it may
-release the mutex. This function takes in argument an event on which it should wait,
-and returns once the event is triggered by the reception of a segment.
-While waiting, it releases the mutex to allow the reception task to process incoming segments.
-For example, this function is called from \spark{open} after the connection
-has been initialized, to wait until the connection has been established or not, in order to
-correctly set the error that indicates to the user the state of the connection.
+The concurrent implementation of the TCP protocol allows for multiple interactions between the tasks, synchronous (with a call to \spark{TCP_Wait_For_Events}) or asynchronous (between two user function calls), and thus, numerous possible state changes at any given moment.
+This makes the functional-specification verification significantly challenging, as SPARK does not have a native mode to deal precisely with interactions related to concurrency.
+To address this, we introduced SPARK functions with contracts that model how the different possible interactions can modify the state of a socket.
+Finding a correct and precise contract for \spark{TCP_Wait_For_Events} is challenging but needed for the verification of the TCP user functions.
+The solution will be elaborated in \cref{subsubsec:dealing_concurrency}.
 
-% Yannick: we talk about error above, but there is no error if the connection is established???
-% This needs rephrasing
-
-A correct but imprecise contract for \spark{TCP_Wait_For_Events} would allow any change in the TCP
-state machine while waiting. But it would make it impossible to prove that its callers
-satisfy their functional specification. We need a more precise contract for this function.
+\todo[inline]{GC: modified up to here}
 
 \subsection{Dealing with synchronous changes of states}
+\label{subsubsec:dealing_concurrency}
 
@@
 
+The specifications are fully proved with GNATprove in 9m54s on a 12 processors with 16GB RAM laptop.
+
 % above: need time to prove automatically with GNATprove with the spec of your laptop (num of cores, amount of memory)

### 2021-05-27 SecDev21/paper.tex
Synchronous exchange part

@@
 Finding a correct and precise contract for \spark{TCP_Wait_For_Events} is challenging but needed for the verification of the TCP user functions.
-The solution will be elaborated in \cref{subsubsec:dealing_concurrency}.
+The solution will be elaborated in the next section.
 
-\todo[inline]{GC: modified up to here}
-
-\subsection{Dealing with synchronous changes of states}
-\label{subsubsec:dealing_concurrency}
+\subsection{Dealing with synchronous exchanges}
 
@@
 
-The function \spark{TCP_Wait_For_Events} first checks if the event is not already
-true. If not, the mutex is released until a segment that triggered the event is
-received or until the timeout is reached. Each time a new segment is received, it
-is processed and if it contains a segment that causes a change of state, the TCP
-state is updated. Then the events are updated, and if the waited event becomes
-true due to the change of state, it is triggered.
+Sometimes the resulting state of a user function can be affected by the rest of the TCP tasks is when the user's functions release the mutex to allow for other events to take place.
+In this case, the function \spark{TCP_Wait_For_Events} is called from the user function to check if the event requested is completed.
+For example, the \spark{TCP_Connect} user function calls \spark{TCP_Wait_For_Events} to wait for the completion of the socket connection event, namely the \spark{SOCKET\_EVENT\_CONNECTED}.
+The call to the \spark{TCP_Wait_For_Events} releases the mutex, and when the \emph{arriving} task receives the appropriate segments, in this case, a segment with the \syn flag, the connection will be established.
+This will move the socket to the \sestab state.
+In the meantime, the \spark{TCP\_Update\_Events} function is monitoring for any state changes and when it notices that the connection is established it will trigger the \spark{SOCKET\_EVENT\_CONNECTED} event, using the OS event mechanism.
+This will allow the \spark{TCP\_Connect} user function to resume execution.
 
@@
     \caption{Function to compute the possible state after the completion of a
-particular event that is requested by a user-task related function.}
+particular event that is requested by a user function.}
 \label{algo:waitForEvents}
@@
 
-An event is considered to be true or not depending on the state of the socket.
-This computation is done by the function \spark{TCP_Update_Events}. We have put
-a contract on it, which is proved to be correct by SPARK. Then we can write a
-function that emulates what is done when \spark{TCP_Wait_For_Events} is called.
-We have written a procedure \spark{TCP_Wait_For_Events_Proof}, reproduced in
-algorithm~\ref{algo:waitForEvents} to compute all the possible TCP states that a
-socket can have after a wait. This algorithm simulates the reception of
-one message, update the events and return if the event waited is reached.
-Otherwise it repeats the procedures. The for loop only requires three steps
-because all the step in the state machine can be reached in three segment
-receptions without user interaction.
+Essentially, the function \spark{TCP_Wait_For_Events} releases the mutex to allow for required events to happen.
+The function \spark{TCP_Update_Events} detects that the event expected outcome is completed by monitoring the changes to the TCP states.
+Then it updates the specified event to be true in the socket structure, and it raises the desired event to allow the resuming of the user function that requested the event in the first place.
+Thus, the \spark{TCP_Update_Events} function is called when a segment is being received, and a state change took place.
+This makes the \spark{TCP_Update_Events} function a perfect candidate to introduce contracts in SPARK and model the possible states after the completion of each event.
 
-On \spark{TCP_Wait_For_Events_Proof}, a contract has been put to describe all
-the possible states that a socket can have after the wait. It is only possible
-to prove this contract with SPARK if the function named
-\spark{TCP_Process_One_Segment} in algorithm~\ref{algo:waitForEvents}
-has a postcondition.
+To model any possible state-changes upon the reception of a segment in SPARK, the function \spark{TCP_Process_One_Segment} is introduced.
+The behavior of \spark{TCP_Wait_For_Events} is modeled by the ghost function \spark{TCP_Wait_For_Events_Proof} shown in \cref{algo:waitForEvents}.
+It computes the set of possible final states after the completion of an event by reusing the function \spark{TCP_Process_One_Segment}.
+Both functions have the same behavior, and thus, the contract of the function \spark{TCP_Wait_For_Events_Proof} is also attached to \spark{TCP_Wait_For_Events}.
+This ghost function is proved automatically without even a loop invariant thanks to loop unrolling in GNATprove.
+
+\todo[inline]{GC: modified up to here}
 
 \subsection{Symbolic execution to extract contracts}
+
+\todo[inline]{Is this part good enough or does it need to be rewritten?}
 

### 2021-05-28 SecDev21/paper.tex
slightly change the part on symbolic execution

@@
 Generally, every TCP communication session will go through the following three phases (if no error occurs): 1)~opening the connection, 2)~sending and receiving the data, and 3)~closing the connection.
-The behavior of the three-phase communication can be described by the state machine given in \cref{Fig:statemachine}.
+The behavior of the three-phase communication can be described by the state machine given in \Cref{Fig:statemachine}.
 An edge represents a state transition.
@@
 
-The state machine in \cref{Fig:statemachine} does not represent the complete protocol specification; for example, it does not reflect error conditions or any actions which are not connected with the state changes.
+The state machine in \Cref{Fig:statemachine} does not represent the complete protocol specification; for example, it does not reflect error conditions or any actions which are not connected with the state changes.
 Instead, it gives an overview of all the possible states a TCP connection could reach over its lifetime.
@@
 Thus, there is no assurance that the implementation will only allow valid transitions between the different states that a TCP session can exhibit.
-We aim to use the SPARK technology to verify that the implementation respects the TCP automaton, given in \cref{Fig:statemachine}.
+We aim to use the SPARK technology to verify that the implementation respects the TCP automaton, given in \Cref{Fig:statemachine}.
 Although, this work is focused on hardening the API of the user task, mainly the high-level user functions, to verify the state transitions the rest of the TCP protocol has to be taken into account.
@@
 To model any possible state-changes upon the reception of a segment in SPARK, the function \spark{TCP_Process_One_Segment} is introduced.
-The behavior of \spark{TCP_Wait_For_Events} is modeled by the ghost function \spark{TCP_Wait_For_Events_Proof} shown in \cref{algo:waitForEvents}.
+The behavior of \spark{TCP_Wait_For_Events} is modeled by the ghost function \spark{TCP_Wait_For_Events_Proof} shown in \Cref{algo:waitForEvents}.
 It computes the set of possible final states after the completion of an event by reusing the function \spark{TCP_Process_One_Segment}.
@@
 
-Rewriting the part that processes the incoming segments was out of the scope of
-this work. Instead, we have opted for the use of symbolic execution on the
+The soundness of our verificatin requires all the contracts to be correct, in particular this is true for the contract of \spark{TCP_Process_One_Segment}.
+Get this contract can be done by rewriting in SPARK the part that processes the
+incoming segments, but this is out of the scope of this work.
+Instead, we have opted for the use of symbolic execution on the
 original CycloneTCP C code.
@@
 the authors explain that symbolic execution can be good at inferring contracts.
-For our work, KLEE has been used. It is built on top of LLVM as it takes LLVM
+For our work, KLEE\footnote{\url{https://klee.github.io}} has been used.
+It is built on top of LLVM as it takes LLVM
 bytecode in input and works well with C code thanks to a very simple interface.
@@
 know if the contracts that we have inferred are correct or not.
-A simplified example is shown in figure~\ref{code:kleedriver}. The first lines
-create a symbolic incoming segment to process as well as a symbolic socket.
-At line~\ref{code:kleedriver:tcpProcessSegment} the function that processes the
-segment is called. It is followed at line~\ref{code:kleedriver:assert} by a KLEE
-builtin that checks if an assert is valid for all the symbolic paths
-explored before.
+A simplified example is shown in \Cref{code:kleedriver}. The first lines
+create a symbolic incoming segment to process and a symbolic socket.
+At \cref{code:kleedriver:tcpProcessSegment} the function that processes the
+segment is executed. It is followed at \cref{code:kleedriver:assert} by a KLEE
+builtin that checks if the code leads to a valid state identified in the
+protocol’s specification for all the paths explored.
 
@@
 challenge has been solved and the function
-\spark{TCP_Wait_For_Events} have a correct and an accurate contract, which
+\spark{TCP_Wait_For_Events} has a correct and an accurate contract, which
 makes possible the verification of the functional correctness of the TCP user

### 2021-05-31 SecDev21/paper.tex
Correct sentences

@@
 
-\todo[inline]{Is this part good enough or does it need to be rewritten?}
-
-The soundness of our verificatin requires all the contracts to be correct, in particular this is true for the contract of \spark{TCP_Process_One_Segment}.
-Get this contract can be done by rewriting in SPARK the part that processes the
+The soundness of our verification requires all the contracts to be correct, in particular this is true for the contract of \spark{TCP_Process_One_Segment} that is used to prove the function \spark{TCP_Wait_For_Events}.
+Getting this contract can be done by rewriting in SPARK the part that processes the
 incoming segments, but this is out of the scope of this work.
@@
 order also is being conveyed by the protocol's state machine given in
-figure~\ref{Fig:statemachine}.
+\Cref{Fig:statemachine}.
 Ada's preconditions and postconditions are a powerful tool to express such inter-function
@@
 Simplified contracts extracted from our library are reproduced in
-figure~\ref{fig:functionorder}. If no error occurs, the function
+\Cref{fig:functionorder}. If no error occurs, the function
 \spark{Socket_Connect} sets \spark{Sock.S_Remote_Ip_Addr} to

### 2021-06-01 SecDev21/paper.tex
more reviewing

@@
 
-\todo[inline]{KG: fixed up to here}
-
 \subsection{Rewriting the TCP user functions}
@@
 We aim to use the SPARK technology to verify that the implementation respects the TCP automaton, given in \Cref{Fig:statemachine}.
-Although, this work is focused on hardening the API of the user task, mainly the high-level user functions, to verify the state transitions the rest of the TCP protocol has to be taken into account.
+Although this work is focused on hardening the API of the user task, mainly the high-level user functions, to verify the state transitions, the rest of the TCP protocol has to be taken into account.
 The reason for this is two-fold.
 First, although user-functions can trigger state transitions, the actual transitions are done by library functions that do not belong to the user-task set of functions.
-Second, other parts of the library can trigger state transitions during and between the user-task function calls, which can affect the intermediate and final states that a user related function can exhibit.
+Second, other parts of the library can trigger state transitions during and between the user-task function calls, which can affect the intermediate and final states that a user-related function can exhibit.
 
-The TCP user API related functions were fully translated to SPARK.
-SPARK bindings were also provided to the majority of the rest of the library's C functions to allow their invocation from the SPARK code.
+The TCP user API-related functions were fully translated to SPARK. SPARK bindings were also provided to most of the rest of the library's C functions to allow their invocation from the SPARK code.
 
 RFC 793 document provides all the information needed for the allowed transitions between the different TCP possible states.
-For any transitions that are directly triggered through the call of the user-tasks functions, SPARK contracts, that represent transitions allowed by the functional specification, are embedded within the SPARK translated code.
-The \texttt{TCP\_Change\_State} is a helper function that is called every time a state transition needs to be made, to update the socket's state.
+For any transitions directly triggered by the call of the user-task functions, SPARK contracts, representing transitions allowed by the functional specification, are embedded within the SPARK translated code.
+The \texttt{TCP\_Change\_State} is a helper function that is called every time a state transition needs to be made to update the socket's state.
 An incorrect transition allowed by the code will be detected by the prover.
@@
 % functions.
+
+\todo[inline]{KG: fixed up to here}
 

### 2021-06-01 SecDev21/paper.tex
fixing the synchronous exchanges section

@@
 
-\todo[inline]{KG: fixed up to here}
-
 The concurrent implementation of the TCP protocol allows for multiple interactions between the tasks, synchronous (with a call to \spark{TCP_Wait_For_Events}) or asynchronous (between two user function calls), and thus, numerous possible state changes at any given moment.
-This makes the functional-specification verification significantly challenging, as SPARK does not have a native mode to deal precisely with interactions related to concurrency.
+This makes the functional specification verification significantly challenging, as SPARK does not have a native mode to deal precisely with interactions related to concurrency.
 To address this, we introduced SPARK functions with contracts that model how the different possible interactions can modify the state of a socket.
 Finding a correct and precise contract for \spark{TCP_Wait_For_Events} is challenging but needed for the verification of the TCP user functions.
-The solution will be elaborated in the next section.
+The solution is elaborated in the following section.
 
@@
 
-Sometimes the resulting state of a user function can be affected by the rest of the TCP tasks is when the user's functions release the mutex to allow for other events to take place.
+In the synchronous scenario of interaction between the TCP tasks, the resulting state of a user function can be affected by the rest of the TCP tasks when the user's functions release the mutex to allow for other events to occur.
 In this case, the function \spark{TCP_Wait_For_Events} is called from the user function to check if the event requested is completed.
@@
 This will move the socket to the \sestab state.
-In the meantime, the \spark{TCP\_Update\_Events} function is monitoring for any state changes and when it notices that the connection is established it will trigger the \spark{SOCKET\_EVENT\_CONNECTED} event, using the OS event mechanism.
+In the meantime, the \spark{TCP\_Update\_Events} function is monitoring for any state changes, and when it notices that the connection is established, it will trigger the \spark{SOCKET\_EVENT\_CONNECTED} event, using the OS event mechanism.
 This will allow the \spark{TCP\_Connect} user function to resume execution.
@@
 %   without even a loop invariant thanks to loop unrolling in GNATprove.
+
+\todo[inline]{KG: fixed up to here}
 

### 2021-06-01 SecDev21/paper.tex
adding a todo for Guillaume

@@
 
-\todo[inline]{KG: fixed up to here}
+\todo[inline]{KG to GC : when you are talking about functions you should say where they are located.
+There will be readers that will want to see the coresponding code in the repository.
+Look for examples in the HICLASS report: e.g.
+This is implemented by the Tcp\_Wait\_For\_Events\_Proof function, located in the tcp\_misc\_binding.adb}
 
 Essentially, the function \spark{TCP_Wait_For_Events} releases the mutex to allow for required events to happen.
-The function \spark{TCP_Update_Events} detects that the event expected outcome is completed by monitoring the changes to the TCP states.
+The function \spark{TCP_Update_Events} detects that the event's expected outcome is completed by monitoring the changes to the TCP states.
 Then it updates the specified event to be true in the socket structure, and it raises the desired event to allow the resuming of the user function that requested the event in the first place.
-Thus, the \spark{TCP_Update_Events} function is called when a segment is being received, and a state change took place.
+Thus, the \spark{TCP_Update_Events} function is called when a segment is being received and a state change took place.
 This makes the \spark{TCP_Update_Events} function a perfect candidate to introduce contracts in SPARK and model the possible states after the completion of each event.
 
-To model any possible state-changes upon the reception of a segment in SPARK, the function \spark{TCP_Process_One_Segment} is introduced.
+To model any possible state changes upon the reception of a segment in SPARK, the function \spark{TCP_Process_One_Segment} is introduced.
 The behavior of \spark{TCP_Wait_For_Events} is modeled by the ghost function \spark{TCP_Wait_For_Events_Proof} shown in \Cref{algo:waitForEvents}.
@@
 Both functions have the same behavior, and thus, the contract of the function \spark{TCP_Wait_For_Events_Proof} is also attached to \spark{TCP_Wait_For_Events}.
-This ghost function is proved automatically without even a loop invariant thanks to loop unrolling in GNATprove.
+This ghost function is proved automatically without even a loop invariant, thanks to loop unrolling in GNATprove.
+
+\todo[inline]{KG: fixed up to here}
 

### 2021-06-01 SecDev21/paper.tex
fixing the SE section

@@
 \todo[inline]{KG to GC : when you are talking about functions you should say where they are located.
-There will be readers that will want to see the coresponding code in the repository.
+There will be readers that will want to see the corresponding code in the repository.
 Look for examples in the HICLASS report: e.g.
@@
 
-\todo[inline]{KG: fixed up to here}
-
 \todo[inline]{GC: modified up to here}
@@
 
-The soundness of our verification requires all the contracts to be correct, in particular this is true for the contract of \spark{TCP_Process_One_Segment} that is used to prove the function \spark{TCP_Wait_For_Events}.
-Getting this contract can be done by rewriting in SPARK the part that processes the
-incoming segments, but this is out of the scope of this work.
-Instead, we have opted for the use of symbolic execution on the
-original CycloneTCP C code.
-Symbolic execution is a mean to execute programs with symbolic values rather
-than concrete values. Symbolic execution has already been successfully
-used with other verification methods.
-Indeed, in~\cite{vanoverberghe2008using} the authors present how to use symbolic
-execution to improve deductive verification, and in~\cite{kassios2012comparing}
-the authors explain that symbolic execution can be good at inferring contracts.
-For our work, KLEE\footnote{\url{https://klee.github.io}} has been used.
-It is built on top of LLVM as it takes LLVM
-bytecode in input and works well with C code thanks to a very simple interface.
+The soundness of our verification requires all the contracts to be complete and correct; all the possible state transitions of a TCP connection must be modeled and proved.
+In particular, this is true for the contract of \spark{TCP_Process_One_Segment} that is used to prove the function \spark{TCP_Wait_For_Events}.
+In our case, because not all the TCP library is translated to SPARK, any transitions triggered from the C part of the library are not accounted, since they can not be proved.
+To overcome this issue, symbolic execution was used on the original CycloneTCP C code to exhaustively verify any contracts related to transitions triggered from the C parts of the library.
+Once these contracts are verified, the relevant C code is proved to respect the TCP functional specification protocol.
+
+Symbolic execution is a means of executing programs with symbolic values rather than concrete values \cite{symbolic_exec_survey:2018}.
+Exhaustive symbolic execution can be considered sound and complete (it prevents false negatives and false positives), and thus, it is suitable for formal verification.
+In \cite{vanoverberghe2008using}, the authors present how symbolic execution can be used to improve deductive verification, and in \cite{kassios2012comparing} the authors demonstrated the technique's value in inferring contracts.
+For our work,  the KLEE\footnote{\url{https://klee.github.io}} symbolic execution engine has been used.
+KLEE is built upon the LLVM optimizer's bytecode, and it offers a very simple yet powerful C interface.
 
@@
 \end{figure}
+
+\todo[inline]{KG: fixed up to here}
 

### 2021-06-01 SecDev21/paper.tex
Rephrasing the se section

@@
 In the synchronous scenario of interaction between the TCP tasks, the resulting state of a user function can be affected by the rest of the TCP tasks when the user's functions release the mutex to allow for other events to occur.
-In this case, the function \spark{TCP_Wait_For_Events} is called from the user function to check if the event requested is completed.
+In this case, the function \spark{TCP_Wait_For_Events} (located in the file \texttt{tcp\_misc\_binding.ads}) is called from the user function to check if the event requested is completed.
 For example, the \spark{TCP_Connect} user function calls \spark{TCP_Wait_For_Events} to wait for the completion of the socket connection event, namely the \spark{SOCKET\_EVENT\_CONNECTED}.
@@
 
-\todo[inline]{KG to GC : when you are talking about functions you should say where they are located.
+% \todo[inline]{KG to GC : when you are talking about functions you should say where they are located.
 There will be readers that will want to see the corresponding code in the repository.
@@
 
-To model any possible state changes upon the reception of a segment in SPARK, the function \spark{TCP_Process_One_Segment} is introduced.
-The behavior of \spark{TCP_Wait_For_Events} is modeled by the ghost function \spark{TCP_Wait_For_Events_Proof} shown in \Cref{algo:waitForEvents}.
+To model any possible state changes upon the reception of a segment in SPARK, the function \spark{TCP_Process_One_Segment} located in the file \texttt{tcp\_fsm\_binding.ads} is introduced.
+The behavior of \spark{TCP_Wait_For_Events} is modeled by the ghost function \spark{TCP_Wait_For_Events_Proof} (that can be found in the file \texttt{tcp\_misc\_binding.ads}) shown in \Cref{algo:waitForEvents}.
 It computes the set of possible final states after the completion of an event by reusing the function \spark{TCP_Process_One_Segment}.
@@
 This ghost function is proved automatically without even a loop invariant, thanks to loop unrolling in GNATprove.
-
-\todo[inline]{GC: modified up to here}
 
@@
 
-The first step in order to extract contracts was to read the state machine to
-know all the possible transitions that can be done by the reception of a flag,
-and adapt it to
-the subtleties of CycloneTCP. The next step is to write a driver to run KLEE and
-know if the contracts that we have inferred are correct or not.
-A simplified example is shown in \Cref{code:kleedriver}. The first lines
-create a symbolic incoming segment to process and a symbolic socket.
-At \cref{code:kleedriver:tcpProcessSegment} the function that processes the
-segment is executed. It is followed at \cref{code:kleedriver:assert} by a KLEE
-builtin that checks if the code leads to a valid state identified in the
-protocol’s specification for all the paths explored.
+We want to prove that the C code implementation of the respects rules extracted from RFC 793.
+To achieve this, symbolic execution is used to verify that all the possible paths when executing the \spark{tcpProcessSegment} function will result in a state that respects the RFC 793 rules.
+This process requires three steps, that are show in \Cref{code:kleedriver}.
+A symbolic incoming segment is created in the first lines to allow us to account for all possible incoming segments.
+Then at \cref{code:kleedriver:tcpProcessSegment} the function \texttt{tcpProcessSegment} is executed symbolically with KLEE.
+Finally, it is followed at \cref{code:kleedriver:assert} by a call to \texttt{klee\_assert} that checks if the code leads to a valid state identified in the protocol’s specification.
 
-Once the contract is validated by KLEE, it can be copied out on the SPARK
-function \spark{TCP_Process_One_Segment}. With this step done, the concurrency
-challenge has been solved and the function
-\spark{TCP_Wait_For_Events} has a correct and an accurate contract, which
-makes possible the verification of the functional correctness of the TCP user
+Once the \texttt{klee\_assert} is validated by KLEE, which means that the code respects the RFC 793 specification, it can be copied out as a valid SPARK contract for the function \spark{TCP_Process_One_Segment}.
+With this step done, the concurrency challenge has been solved and the function \spark{TCP_Wait_For_Events} has a correct and an accurate contract, which makes possible the verification of the functional correctness of the TCP user
 functions thanks to SPARK.
@@
 
-However weeknesses exist in our implementation, in particular because all the
+However weaknesses exist in our implementation, in particular because all the
 underlying layers are still written in C. The principal function of those layers

### 2021-06-02 SecDev21/paper.tex
finishing the KLEE part

@@
 Symbolic execution is a means of executing programs with symbolic values rather than concrete values \cite{symbolic_exec_survey:2018}.
-Exhaustive symbolic execution can be considered sound and complete (it prevents false negatives and false positives), and thus, it is suitable for formal verification.
+Exhaustive symbolic execution can be considered sound and complete (it prevents false negatives and false positives), and thus, it is complementary to formal verification.
 In \cite{vanoverberghe2008using}, the authors present how symbolic execution can be used to improve deductive verification, and in \cite{kassios2012comparing} the authors demonstrated the technique's value in inferring contracts.
-For our work,  the KLEE\footnote{\url{https://klee.github.io}} symbolic execution engine has been used.
+For our work, the KLEE\footnote{\url{https://klee.github.io}} symbolic execution engine is used.
 KLEE is built upon the LLVM optimizer's bytecode, and it offers a very simple yet powerful C interface.
@@
 
+Symbolic execution is used to verify that all the possible paths when executing the \spark{tcpProcessSegment} function will result in a state that respects the RFC 793 extracted specifications.
+This process requires three steps shown in \Cref{code:kleedriver}.
+Firstly, a symbolic incoming segment is created, allowing to account for all possible incoming segments (see lines 1-4).
+Then at \cref{code:kleedriver:tcpProcessSegment}, function \texttt{tcpProcessSegment} is executed symbolically by KLEE.
+Finally, the assertion, \texttt{klee\_assert}, at \cref{code:kleedriver:assert} checks if the code leads to a valid state, one that is within the protocol's specification.
+If there is no raised assertion, then the contract represented by the assertion is deemed proved, and the relevant C code respects the protocol's specification regarding that contract.
+The assertion is then converted to a SPARK contract and added to the function \spark{TCP_Process_One_Segment}.
+This enables the SPARK prover to reason about the possible side-effects caused by the C-code on the state of a TCP connection and thus, enables the functional correctness verification of the TCP user functions.
+
 \todo[inline]{KG: fixed up to here}
-
-We want to prove that the C code implementation of the respects rules extracted from RFC 793.
-To achieve this, symbolic execution is used to verify that all the possible paths when executing the \spark{tcpProcessSegment} function will result in a state that respects the RFC 793 rules.
-This process requires three steps, that are show in \Cref{code:kleedriver}.
-A symbolic incoming segment is created in the first lines to allow us to account for all possible incoming segments.
-Then at \cref{code:kleedriver:tcpProcessSegment} the function \texttt{tcpProcessSegment} is executed symbolically with KLEE.
-Finally, it is followed at \cref{code:kleedriver:assert} by a call to \texttt{klee\_assert} that checks if the code leads to a valid state identified in the protocol’s specification.
-
-Once the \texttt{klee\_assert} is validated by KLEE, which means that the code respects the RFC 793 specification, it can be copied out as a valid SPARK contract for the function \spark{TCP_Process_One_Segment}.
-With this step done, the concurrency challenge has been solved and the function \spark{TCP_Wait_For_Events} has a correct and an accurate contract, which makes possible the verification of the functional correctness of the TCP user
-functions thanks to SPARK.
-
 \section{Hardening the user's API}

### 2021-06-02 SecDev21/paper.tex
reviewing user's api section

@@
 
-\todo[inline]{KG: fixed up to here}
 \section{Hardening the user's API}
@@
 
-A common problem faced by users of the CycloneTCP library is the need to
-properly chain calls of the Socket API, in a way that respects the TCP
-automaton. This entails verifying that 1)~the functions are called with respect
-to a defined order and 2)~the return code is checked before continuing the
-processing.
+One of the most significant problems pointed to by the primary author of the \emph{CycloneTCP} library is the incorrect usage of the library's API.
+More specifically, the library's users tend to call the API functions in the wrong order and forget to check the return code of a function call.
+Thus, they are subsequently allowing their  TCP implementation to behave outside of the functional specification of the protocol.
+The following sections demonstrate how SPARK can be used to enforce the correct usage of the library.
+
+\todo[inline]{KG: fixed up to here}
 

### 2021-06-02 SecDev21/paper.tex
moving the info part from results to intro

@@
 The quality of the CycloneTCP is acknowledged by the AMNESIA:33 report~\cite{AMNESIA33}, which classifies it as one of the most resilient TCP/IP stack.
+
+In total, for this work, 50 subprograms have been translated from C to SPARK.
+This accounts for 2266 lines of code, of which 1165 are SPARK logic code (e.g\.~ contracts) used to represent and prove the protocol's specification.
+The overall proving time with GNATprove is 9m and 54s on a 12-processor Intel(R) Core(TM) i7-9750H CPU with 2.60GHz speed and 16GB of RAM.
 
@@
 
-\todo[inline]{KG: fixed up to here}
-
 \subsection{Enforcing the correct order of API functions}
 
-The TCP protocol implies a specific order such that the user can call the API's
-functions without breaking the functional specification of the protocol. This
-order also is being conveyed by the protocol's state machine given in
-\Cref{Fig:statemachine}.
-Ada's preconditions and postconditions are a powerful tool to express such inter-function
-dependencies, while SPARK technology can be used to guarantee that the
-assertions always hold. Thus, preconditions and postconditions were introduced to model
-a partial order on the calls to the API's functions.
+The TCP protocol implies a specific order such that the user can call the API's functions without breaking the functional specification of the protocol.
+This order also is being conveyed by the protocol's state machine given in \Cref{Fig:statemachine}.
+Ada's preconditions and postconditions are a powerful tool to express such inter-function dependencies, while SPARK technology can be used to guarantee that the assertions always hold.
+Thus, preconditions and postconditions were introduced to model a partial order on the calls to the API's functions.
 
@@
 
-Simplified contracts extracted from our library are reproduced in
-\Cref{fig:functionorder}. If no error occurs, the function
-\spark{Socket_Connect} sets \spark{Sock.S_Remote_Ip_Addr} to
-\spark{Remote_Ip_Addr} which is supposed to be
-initialized when the function is called. \spark{Socket_Send} requires to be
-called with a socket not null such that its field \spark{Remote_Ip_Addr} is
-initialized. The only way to ensure that the precondition hold is to
-call \spark{Socket_Connect} before \spark{Socket_Send}.
-The tool GNATprove can statically find if a call to \spark{Socket_Send}
-is not preceded by a call to \spark{Socket_Connect}.
+Simplified contracts extracted from our SPARK implementation are reproduced in \Cref{fig:functionorder}, to demonstrate how SPARK can be used to enforce the correct function calls' ordering.
+If no error occurs, the function \spark{Socket_Connect} sets \spark{Sock.S_Remote_Ip_Addr} to \spark{Remote_Ip_Addr} which is supposed to be initialized when the function is called.
+\spark{Socket_Send} requires to be called with a socket not null such that its field \spark{Remote_Ip_Addr} is initialized.
+The only way to ensure that the precondition hold is to call \spark{Socket_Connect} before \spark{Socket_Send}.
+The tool GNATprove can statically detect if a call to \spark{Socket_Send} is not preceded by a call to \spark{Socket_Connect}.
 
@@
 
-A common mistake is to forget to check the return codes after a call to a socket
-API's function. For instance, the following code would be reject by SPARK:
+The following code demonstrates how SPARK will enforce to the check of the return codes of a function and allow execution to proceed only if no error code is returned:
+
 \begin{lstlisting}[language=Ada,basicstyle=\small\ttfamily]
@@
 \end{lstlisting}
-Indeed, if \spark{Socket_Connect} fails and \spark{Error} is not equal to
-\spark{NO_ERROR}, \spark{Sock.S_Remote_Ip_Addr} might be non initialized, and
-then the function \spark{Socket_Send} cannot be called because of its
-precondition.
+
+In the above scenario the post-condition of  \spark{Socket_Connect} will set the value of \spark{Sock.S_Remote_Ip_Addr} if the function succesfully executes, and thus, \spark{Error} is not equal to \spark{NO_ERROR}.
+In the case an error occurs, \spark{Sock.S_Remote_Ip_Addr} won't be initialized, and thus, the function \spark{Socket_Send} cannot be called because its precondition wont be satisfied.
+
+\todo[inline]{KG: fixed up to here}
 
@@
 
-\subsection{Verified parts}
-
-For achieving this verification, 50 subprograms have been translated from C to
-SPARK. This accounts for \np{2266} lines of code of which \np{1165} are logical
-code to write and prove the specifications.
-
-The specifications are fully proved with GNATprove in 9m54s on a 12 processors with 16GB RAM laptop.
-
-% above: need time to prove automatically with GNATprove with the spec of your laptop (num of cores, amount of memory)
-
 \section{Conclusion and future work}

### 2021-06-02 SecDev21/paper.tex
text improvments

@@
 We aim to use the SPARK technology to verify that the implementation respects the TCP automaton, given in \Cref{Fig:statemachine}.
-Although this work is focused on hardening the API of the user task, mainly the high-level user functions, to verify the state transitions, the rest of the TCP protocol has to be taken into account.
+The work is focused on hardening the API of the user task, mainly the high-level user functions, to verify the state transitions.
+To achieve this, the rest of the TCP protocol, not related to the user task, also has to be taken into account.
 The reason for this is two-fold.
@@
 Second, other parts of the library can trigger state transitions during and between the user-task function calls, which can affect the intermediate and final states that a user-related function can exhibit.
-
-The TCP user API-related functions were fully translated to SPARK. SPARK bindings were also provided to most of the rest of the library's C functions to allow their invocation from the SPARK code.
+The TCP user API-related functions were fully translated to SPARK.
+SPARK bindings were also provided to most of the rest of the library's C functions to allow their invocation from the SPARK code.
 
@@
 
-In the above scenario the post-condition of  \spark{Socket_Connect} will set the value of \spark{Sock.S_Remote_Ip_Addr} if the function succesfully executes, and thus, \spark{Error} is not equal to \spark{NO_ERROR}.
+In the above scenario the post-condition of \spark{Socket_Connect} will set the value of \spark{Sock.S_Remote_Ip_Addr} if the function successfully executes, and thus, \spark{Error} is not equal to \spark{NO_ERROR}.
 In the case an error occurs, \spark{Sock.S_Remote_Ip_Addr} won't be initialized, and thus, the function \spark{Socket_Send} cannot be called because its precondition wont be satisfied.
@@
 
-\section{Results}
+\section{Bugs Captured}
 \label{sec:results}
-
-\subsection{Bugs found}
 

### 2021-06-02 SecDev21/paper.tex
adding KLEE citation to save space

@@
 The original \emph{CycloneTCP} library provides no guarantees that the implementation respects the TCP functional specification.
-Thus, there is no assurance that the implementation will only allow valid transitions between the different states that a TCP session can exhibit.
+Thus, there is no assurance that the implementation will allow only valid transitions between the different states that a TCP session can exhibit.
 We aim to use the SPARK technology to verify that the implementation respects the TCP automaton, given in \Cref{Fig:statemachine}.
@@
 In \cite{vanoverberghe2008using}, the authors present how symbolic execution can be used to improve deductive verification, and in \cite{kassios2012comparing} the authors demonstrated the technique's value in inferring contracts.
-For our work, the KLEE\footnote{\url{https://klee.github.io}} symbolic execution engine is used.
+For our work, the KLEE v2.1 symbolic execution engine is used \cite{KLEE}.
 KLEE is built upon the LLVM optimizer's bytecode, and it offers a very simple yet powerful C interface.

### 2021-06-02 SecDev21/paper.tex
rewording

--- 2021-06-02 SecDev21/paper.tex
@@
 
-\title{Layered Formal Verification of an Industrial TCP/IP Stack}
+\title{Layered Formal Verification of a TCP/IP Stack}
 
@@
 
-Existing research has dealt with formally verifying several protocols of TCP/IP stack levels.
-Early work formalized the TCP protocol to prove some key properties of the protocol itself~\cite{smith1996formal}.
-More recently, researchers have formally specified TCP, UDP and the Sockets API in HOL~\cite{tcp2005sigcomm,ridge2008rigorous} and used the resulting specification for testing existing libraries.
-Others have formally specified and implemented the SSL/TLS protocol in F*\cite{bhargavan2013implementing}, and integrated the resulting implementation in existing libraries.
+Existing research has dealt with formally verifying several protocols of the TCP/IP stack.
+Early work formalized and proved some key properties of the TCP protocol~\cite{smith1996formal}.
+More recently, researchers have formally specified TCP, UDP and the Sockets API in HOL and tested existing libraries against the formal specification~\cite{tcp2005sigcomm,ridge2008rigorous}.
+Others have formally specified and implemented the SSL/TLS protocol in F* and integrated the resulting implementation in existing libraries~\cite{bhargavan2013implementing}.
 To the best of our knowledge, no prior work attempted to formally verify an implementation of TCP, although the authors of~\cite{ridge2008rigorous} suggest that it would possible to extract an actual implementation in Haskell from their specification.
 
-In this paper, we take a practical approach, where the security of an existing TCP/IP stack implementation written in ANSI C programming language is incrementally enhanced by replacing parts of its code with formally verified code in SPARK~\cite{mccormick_chapin_2015}, a subset of the Ada programming language supported by formal verification tools.
+In this paper, we take a practical approach, where the security of an existing TCP/IP stack implementation written in the C programming language is incrementally enhanced by replacing parts of its code with formally verified code in SPARK~\cite{mccormick_chapin_2015}, a subset of the Ada programming language supported by formal verification tools.
 We chose CycloneTCP, a professional-grade open source embedded TCP/IP library developed by the Oryx Embedded company~\cite{CycloneTCP}.
@@
 This work focuses on the TCP and Socket components of the stack.
-The library's TCP implementation is meant to conform with the RFC 793 protocol specifications~\cite{rfc793}.
+The library's TCP implementation is meant to conform to the RFC 793 protocol specifications~\cite{rfc793}.
 The quality of the CycloneTCP is acknowledged by the AMNESIA:33 report~\cite{AMNESIA33}, which classifies it as one of the most resilient TCP/IP stack.
@@
 
-We present the TCP protocol in section~\ref{sec:TCP} and then briefly the
-chosen stack, the Oryx Embedded one, in section~\ref{sec:stack}
-before presenting the specification techniques used in section~\ref{sec:spec}
-and then in sections~\ref{sec:verif} and~\ref{sec:API} we explain how we
-formally implemented in SPARK the TCP user functions and how we hardened the
-user's API. Finally we end in section~\ref{sec:results} by showing the results
-of our work.
+We present the TCP protocol in section~\ref{sec:TCP} and the specification
+techniques used in section~\ref{sec:spec}. We explain how we formally
+implemented in SPARK the TCP user functions in section~\ref{sec:verif} and how
+we hardened the user's API in section~\ref{sec:API}. Finally we conclude with a
+summary of the results of our work in section~\ref{sec:results}.
 
@@
 An edge represents a state transition.
-An edge-label in the format of $\frac{\ \ x \ \ }{\ \ y \ \ }$ represents the actions associated to the state transition, where $x$ is triggering the transition, and $y$ is produced by the transition.
-The trigger $x$ is either an explicit action like \textit{Close}, or the arrival of specific flags, while $y$ always represents the flags transmitted in response to the action that caused the state transition, when present.
+An edge-label in the format of $\frac{\ \ x \ \ }{\ \ y \ \ }$ represents the actions associated with the state transition, where $x$ is triggering the transition, and $y$ is produced by the transition.
+The trigger $x$ is either an explicit action like \textit{Close}, or the arrival of specific flags, while $y$ always represents the flags transmitted in response when present.
 An explicit action is either a user-triggered action (through the user's API) or an action automatically triggered by a timeout event.
@@
 
-The state machine in \Cref{Fig:statemachine} does not represent the complete protocol specification; for example, it does not reflect error conditions or any actions which are not connected with the state changes.
+The state machine in \Cref{Fig:statemachine} does not represent the complete protocol specification; for example, it does not reflect error conditions or any actions which are not connected to the state changes.
 Instead, it gives an overview of all the possible states a TCP connection could reach over its lifetime.
-Each state has a meaning.
+Each state has a clear meaning.
 \slisten represents waiting for a connection request from any remote TCP and port.
@@
 Finally, timers control the timeouts, for example, the retransmission timeout to retransmit a message or the time-wait timeout to close the connection after a specified amount of time elapsed.
-Thus, corresponding transitions to the timeout events can also be triggered by this task.
+Thus, transitions corresponding to the timeout events can also be triggered by this task.
 A complete description of the protocol's specification can be found in the RFC 793 document~\cite{rfc793}.
@@
 \emph{GNATprove} \cite{GNATProve:users_manual} is the tool that provides formal verification for SPARK through two kinds of analyses.
-The first enables flow analysis to check the initialization of variables, look at data dependencies between inputs and outputs of subprograms, and detects unused assignments and unmodified variables.
+The first enables flow analysis to check the initialization of variables, look at data dependencies between inputs and outputs of subprograms, and detect useless code.
 The second uses deductive verification to generate verification conditions for SMT solvers via a weakest-precondition calculus to detect possible runtime errors and violations of functional contracts.
@@
 
-The verification of the entire functional specification of the \emph{CycloneTCP} library is beyond the scope of this work.
+The verification of the entire functional specification of the CycloneTCP library is beyond the scope of this work.
 Instead, the aim is to provide significant security hardening to mostly-C libraries by selective use of SPARK.
@@
 
-Ada and SPARK offer an easy-to-use mechanism for interfacing with C code \cite{}.
+Ada and SPARK offer an easy-to-use mechanism for interfacing with C code.
 The mechanism allows C-written function to be called by Ada/SPARK code and vice versa and for representing shared data objects.
@@
 SPARK uses a pointer ownership model to prevent bugs involving pointers such as memory leak or double free, which imposes some restrictions on how pointers are being manipulated.
-When a memory area is allocated, a pointer points to this memory area, and thus, it is considered "owned.'' The ownership of the memory can be moved, but the memory must be deallocated before the end of the execution of the program.
+When a memory area is allocated, a pointer points to this memory area, and thus, it is considered ``owned.'' The ownership of the memory can be moved, but the memory must be deallocated before the end of the execution of the program.
 
@@
 Thus we use \emph{ghost code} in SPARK to focus the verification effort on the relevant subset of the fields, which also makes the specifications easier to write and understand.
-This is a solution to the classical problem of specifying the \emph{frame condition} for a function, that is, the set of possibly modified objects.
+This is a solution to the classical problem of specifying the \emph{frame condition} for a function, that is, the set of objects possibly modified by the function.
 \emph{Ghost code} is used by GNATprove for analysis but is not being compiled in the final executable.
@@
 
+\noindent
 This states that the relevant fields of \spark{Sock} selected by \spark{Model} are not changed by the function call, except for the field \spark{State} that is equal to \spark{TCP_STATE_CLOSED} after the call.
@@
 \item The transitions between states must respect the order given by the
-state machine described in figure~\ref{Fig:statemachine}. In particular, a transition from a state to another is valid only if the state machine allows for a transition between these two states.
+state machine described in \Cref{Fig:statemachine}. In particular, a transition from a state to another is valid only if the state machine allows for a transition between these two states.
 \item RFC 793 describes the conditions for the state of the socket which need to hold when a user's function is called~\cite[p.
-52]{rfc793}.
-In particular, these conditions specify in what state the function can be called without returning an error.
+52]{rfc793}, to avoid errors.
 These conditions were turned into preconditions in our formalization.
@@
 
-The original \emph{CycloneTCP} library provides no guarantees that the implementation respects the TCP functional specification.
+The original CycloneTCP library provides no guarantees that the implementation respects the TCP functional specification.
 Thus, there is no assurance that the implementation will only allow valid transitions between the different states that a TCP session can exhibit.
 We aim to use the SPARK technology to verify that the implementation respects the TCP automaton, given in \Cref{Fig:statemachine}.
-The work is focused on hardening the API of the user task, mainly the high-level user functions, to verify the state transitions.
-To achieve this, the rest of the TCP protocol, not related to the user task, also has to be taken into account.
+Although this work is focused on hardening the API of the user task, mainly the high-level user functions, to verify the state transitions, the rest of the TCP protocol has to be taken into account.
 The reason for this is two-fold.
-First, although user-functions can trigger state transitions, the actual transitions are done by library functions that do not belong to the user-task set of functions.
-Second, other parts of the library can trigger state transitions during and between the user-task function calls, which can affect the intermediate and final states that a user-related function can exhibit.
-The TCP user API-related functions were fully translated to SPARK.
-SPARK bindings were also provided to most of the rest of the library's C functions to allow their invocation from the SPARK code.
+First, although TCP user functions can trigger state transitions, the actual transitions are done by other library functions called from TCP user functions.
+Second, other parts of the library can trigger state transitions during and between TCP user function calls, which can affect the intermediate and final states that a TCP user function can exhibit.
+The TCP user functions were fully translated to SPARK. SPARK bindings were also provided for most of the rest of the library's C functions to allow their invocation from the SPARK code.
 
 RFC 793 document provides all the information needed for the allowed transitions between the different TCP possible states.
-For any transitions directly triggered by the call of the user-task functions, SPARK contracts, representing transitions allowed by the functional specification, are embedded within the SPARK translated code.
-The \texttt{TCP\_Change\_State} is a helper function that is called every time a state transition needs to be made to update the socket's state.
-An incorrect transition allowed by the code will be detected by the prover.
+For any transitions directly triggered by calling the TCP user functions, SPARK contracts, representing transitions allowed by the functional specification, are embedded within the SPARK translated code.
+The helper function \texttt{TCP\_Change\_State} is called explicitly every time a state transition needs to be made to update the socket's state.
+Any incorrect transition allowed by the code will be detected by GNATprove.
 
@@
 
-The concurrent implementation of the TCP protocol allows for multiple interactions between the tasks, synchronous (with a call to \spark{TCP_Wait_For_Events}) or asynchronous (between two user function calls), and thus, numerous possible state changes at any given moment.
+The concurrent implementation of the TCP protocol allows for multiple interactions between the tasks, both synchronous (with a call to \spark{TCP_Wait_For_Events}) or asynchronous (between two user function calls), and thus, numerous possible state changes at any given moment.
 This makes the functional specification verification significantly challenging, as SPARK does not have a native mode to deal precisely with interactions related to concurrency.
 To address this, we introduced SPARK functions with contracts that model how the different possible interactions can modify the state of a socket.
-Finding a correct and precise contract for \spark{TCP_Wait_For_Events} is challenging but needed for the verification of the TCP user functions.
+Finding a correct and precise contract for \spark{TCP_Wait_For_Events} was challenging but needed for the verification of the TCP user functions.
 The solution is elaborated in the following section.
@@
 
-% Yannick: below "a new segement...if it contains a new segment" is not phrased correctly
-
-In the synchronous scenario of interaction between the TCP tasks, the resulting state of a user function can be affected by the rest of the TCP tasks when the user's functions release the mutex to allow for other events to occur.
-In this case, the function \spark{TCP_Wait_For_Events} (located in the file \texttt{tcp\_misc\_binding.ads}) is called from the user function to check if the event requested is completed.
-For example, the \spark{TCP_Connect} user function calls \spark{TCP_Wait_For_Events} to wait for the completion of the socket connection event, namely the \spark{SOCKET\_EVENT\_CONNECTED}.
-The call to the \spark{TCP_Wait_For_Events} releases the mutex, and when the \emph{arriving} task receives the appropriate segments, in this case, a segment with the \syn flag, the connection will be established.
+In the synchronous scenario of interaction between the TCP tasks, the resulting state of a user function can be affected by the rest of the TCP tasks when the TCP user function releases the mutex to allow for other events to occur.
+In this case, the function \spark{TCP_Wait_For_Events} is called from the TCP user function to check if the event requested is completed.
+For example, the \spark{TCP_Connect} user function calls \spark{TCP_Wait_For_Events} to wait for the completion of the socket connection event, namely \spark{SOCKET\_EVENT\_CONNECTED}.
+The call to \spark{TCP_Wait_For_Events} releases the mutex, and when the relevant task receives the appropriate segment, in this case, a segment with the \syn flag, the connection will be established.
 This will move the socket to the \sestab state.
@@
     \caption{Function to compute the possible state after the completion of a
-particular event that is requested by a user function.}
+particular event that is requested by a TCP user function.}
 \label{algo:waitForEvents}
@@
 The function \spark{TCP_Update_Events} detects that the event's expected outcome is completed by monitoring the changes to the TCP states.
-Then it updates the specified event to be true in the socket structure, and it raises the desired event to allow the resuming of the user function that requested the event in the first place.
+Then it updates the specified event to be true in the socket structure, and it raises the desired event to allow resuming the TCP user function that requested the event in the first place.
 Thus, the \spark{TCP_Update_Events} function is called when a segment is being received and a state change took place.
@@
 
-To model any possible state changes upon the reception of a segment in SPARK, the function \spark{TCP_Process_One_Segment} located in the file \texttt{tcp\_fsm\_binding.ads} is introduced.
-The behavior of \spark{TCP_Wait_For_Events} is modeled by the ghost function \spark{TCP_Wait_For_Events_Proof} (that can be found in the file \texttt{tcp\_misc\_binding.ads}) shown in \Cref{algo:waitForEvents}.
-It computes the set of possible final states after the completion of an event by reusing the function \spark{TCP_Process_One_Segment}.
+To model any possible state changes upon the reception of a segment in SPARK, the function \spark{TCP_Process_One_Segment} is introduced.
+The behavior of \spark{TCP_Wait_For_Events} is modeled by the ghost function \spark{TCP_Wait_For_Events_Proof} shown in \Cref{algo:waitForEvents}.
+It computes the set of possible final states after the completion of an event by calling repeatedly the function \spark{TCP_Process_One_Segment}.
 Both functions have the same behavior, and thus, the contract of the function \spark{TCP_Wait_For_Events_Proof} is also attached to \spark{TCP_Wait_For_Events}.
-This ghost function is proved automatically without even a loop invariant, thanks to loop unrolling in GNATprove.
+This ghost function is proved automatically without even requiring a loop invariant, thanks to loop unrolling in GNATprove.
 

### 2021-06-02 SecDev21/paper.tex
cutting text

--- 2021-06-02 SecDev21/paper.tex
@@
 More specifically, we define a \emph{model} of the actual socket as a ghost function that can be used in contracts: \spark{Model(Sock)} extracts the fields of interest from socket \spark{Sock}.
-A frequent usage in postconditions is
+A common usage is within a function postcondition:
 
@@
 
-This states that the relevant fields of \spark{Sock} selected by \spark{Model} are not changed by the function call, except for the field \spark{State} that is equal to \spark{TCP_STATE_CLOSED} after the call.
+This states that the fields of \spark{Sock} used by \spark{Model} are not changed when calling the function, except for the field \spark{State} that is equal to \spark{TCP_STATE_CLOSED} after the call.
 
@@
 The work is focused on hardening the API of the user task, mainly the high-level user functions, to verify the state transitions.
-To achieve this, the rest of the TCP protocol, not related to the user task, also has to be taken into account.
+To achieve this, the rest of the TCP protocol, not related to the user task, has to be taken into account.
 The reason for this is two-fold.

### 2021-06-02 SecDev21/paper.tex
rewording

--- 2021-06-02 SecDev21/paper.tex
@@
 The soundness of our verification requires all the contracts to be complete and correct; all the possible state transitions of a TCP connection must be modeled and proved.
-In particular, this is true for the contract of \spark{TCP_Process_One_Segment} that is used to prove the function \spark{TCP_Wait_For_Events}.
-In our case, because not all the TCP library is translated to SPARK, any transitions triggered from the C part of the library are not accounted, since they can not be proved.
+In particular, this is true for the contract of \spark{TCP_Process_One_Segment} that is used to prove the function \spark{TCP_Wait_For_Events}, as it models calls to parts of the library in C that are not proved.
 To overcome this issue, symbolic execution was used on the original CycloneTCP C code to exhaustively verify any contracts related to transitions triggered from the C parts of the library.
-Once these contracts are verified, the relevant C code is proved to respect the TCP functional specification protocol.
 
@@
     numberstyle=\tiny, frame=bottomline, escapeinside={(*@}{@*)}]
-// creation of a fake incoming segment
+// Create a fake incoming segment
 TCPheader *segment = malloc(sizeof(TCPheader));
@@
 klee_assume(segment->flag <= 31);
-// Creation of the socket
+// Create the socket
 Socket *sock = malloc(sizeof(Socket)), oldSock;
@@
 
-// The function that process the segments is called
+// Call the function to process the segment
 tcpProcessSegment(sock, segment); (*@\label{code:kleedriver:tcpProcessSegment}@*)
 
-// Contract to check
+// Check the expected postcondition
 klee_assert( (*@\label{code:kleedriver:assert}@*)
@@
 One of the most significant problems pointed to by the primary author of the \emph{CycloneTCP} library is the incorrect usage of the library's API.
-More specifically, the library's users tend to call the API functions in the wrong order and forget to check the return code of a function call.
-Thus, they are subsequently allowing their  TCP implementation to behave outside of the functional specification of the protocol.
-The following sections demonstrate how SPARK can be used to enforce the correct usage of the library.
+More specifically, the users of the library sometimes call the TCP user functions in the wrong order and forget to check the return code after some calls.
+This causes their TCP implementation to behave outside of the functional specification of the protocol.
+The following sections demonstrate how SPARK can be used to enforce a correct usage of the library.
 
-\subsection{Enforcing the correct order of API functions}
+\subsection{Enforcing the correct order of TCP user functions}
 
-The TCP protocol implies a specific order such that the user can call the API's functions without breaking the functional specification of the protocol.
-This order also is being conveyed by the protocol's state machine given in \Cref{Fig:statemachine}.
+The TCP protocol implies a specific order such that the user can call the TCP user functions without breaking the functional specification of the protocol.
+This order is also being conveyed by the protocol's state machine given in \Cref{Fig:statemachine}.
 Ada's preconditions and postconditions are a powerful tool to express such inter-function dependencies, while SPARK technology can be used to guarantee that the assertions always hold.
-Thus, preconditions and postconditions were introduced to model a partial order on the calls to the API's functions.
+Thus, preconditions and postconditions were introduced to model a partial order on the calls to the TCP user functions.
 
@@
 \end{lstlisting}
-\caption{An example of how function calls can be ordered by pre- and
+\caption{An example of how function calls can be ordered by preconditions and
 postconditions}
@@
 If no error occurs, the function \spark{Socket_Connect} sets \spark{Sock.S_Remote_Ip_Addr} to \spark{Remote_Ip_Addr} which is supposed to be initialized when the function is called.
-\spark{Socket_Send} requires to be called with a socket not null such that its field \spark{Remote_Ip_Addr} is initialized.
+\spark{Socket_Send} requires to be called with a not null socket such that its field \spark{Remote_Ip_Addr} is initialized.
 The only way to ensure that the precondition hold is to call \spark{Socket_Connect} before \spark{Socket_Send}.
-The tool GNATprove can statically detect if a call to \spark{Socket_Send} is not preceded by a call to \spark{Socket_Connect}.
+By analyzing user's code, GNATprove will statically detect if a call to \spark{Socket_Send} is not preceded by a call to \spark{Socket_Connect}.
 
@@
 
-The following code demonstrates how SPARK will enforce to the check of the return codes of a function and allow execution to proceed only if no error code is returned:
+The following code demonstrates how SPARK will enforce that return codes of TCP user functions are checked, and allow execution to proceed only if no error code is returned:
 
@@
 
-In the above scenario the post-condition of \spark{Socket_Connect} will set the value of \spark{Sock.S_Remote_Ip_Addr} if the function successfully executes, and thus, \spark{Error} is not equal to \spark{NO_ERROR}.
-In the case an error occurs, \spark{Sock.S_Remote_Ip_Addr} won't be initialized, and thus, the function \spark{Socket_Send} cannot be called because its precondition wont be satisfied.
+In the above scenario, the postcondition of \spark{Socket_Connect} specifies that \spark{Sock.S_Remote_Ip_Addr} is initialized only if the function successfully executes, and thus, \spark{Error} is equal to \spark{NO_ERROR}.
+In the case where an error occurs, nothing is specified about the value of \spark{Sock.S_Remote_Ip_Addr}, and thus, the function \spark{Socket_Send} cannot be called because its precondition won't be satisfied.
 
@@
 
-The use of SPARK for the verification of the TCP protocol has given good
-results. It has helped to find bugs in the existing implementation. But more
-than finding bugs, the interest of this work lie in the face that it is a
-step toward a secure implementation. It has been proved by the tool GNATprove
-that our implementation is free of run-time errors and that all the transitions
-in the TCP state machine are done with respect to the norm.
+We selectively applied formal verification to a layer of a TCP/IP library, by
+translating TCP user functions into SPARK, and modelling asynchronous events
+with ghost code. This allowed us to detect two bugs in the original C
+implementation, related to memory management and concurrency. This is a step
+towards a more secure implementation. The translated code has been proven to be
+free of run-time errors and to respect the transitions of the TCP state machine
+as specified in RFC 793.
 
-However weaknesses exist in our implementation, in particular because all the
-underlying layers are still written in C. The principal function of those layers
-is to format packet before they are sent, or parse incoming packets, check their
-integrity and transmit the resulting payload to the corresponding upper level
+These results depend upon other layers in the C programming language, which
+were not formally verified to the same level. Instead, we used symbolic
+execution with KLEE to formally verify the contracts used in SPARK to model the
+corresponding behaviors. The principal function of those layers
+is to format packets before they are sent, or parse incoming packets, check their
+integrity and transmit the resulting payload to the corresponding upper-level
 layer. This processing part can be a source of errors and bugs as explain
 in~\cite{Reiher2019RecordFluxFM}.
-Using the RecordFlux DSL to parse the packets of the different protocols is the
-next step to make the stack safer. Using RecordFlux to parse TCP message
-would be a big step to finish the translation of the TCP protocol in SPARK.
-The postconditions extracted by KLEE could be proved by GNATprove for a
-safer result.
+As an alternative to manual translation into SPARK, we envision 
+using the RecordFlux DSL to specify the parser and printer of the packets of the
+different protocols, and to generate provable SPARK code from this specification.
+This would complete our current verification of the CycloneTCP stack towards
+the lower layers.
 
-Our verification of the TCP protocol only focuses on the validity of the
-transitions in the TCP state machine, but some other properties are formulated
-in RFC 793, and could be verified in the implementation to make it even
-more robust.
+%Finally, our verification of the TCP protocol only focuses on the validity of the
+%transitions in the TCP state machine, but other properties are formulated
+%in RFC 793, and could be verified in the implementation to make it even
+%more robust.
 

### 2021-06-02 SecDev21/paper.tex

--- 2021-06-02 SecDev21/paper.tex
@@
 \noindent
-This states that the relevant fields of \spark{Sock} selected by \spark{Model} are not changed by the function call, except for the field \spark{State} that is equal to \spark{TCP_STATE_CLOSED} after the call.
+This states that the fields of \spark{Sock} used by \spark{Model} are not changed by the function call, except for the field \spark{State} that is equal to \spark{TCP_STATE_CLOSED} after the call.
 
@@
 We aim to use the SPARK technology to verify that the implementation respects the TCP automaton, given in \Cref{Fig:statemachine}.
-Although this work is focused on hardening the API of the user task, mainly the high-level user functions, to verify the state transitions, the rest of the TCP protocol has to be taken into account.
+The work is focused on hardening the API of the user task, mainly the high-level user functions, to verify the state transitions.
+To achieve this, the rest of the TCP protocol, not related to the user task, has to be taken into account.
 The reason for this is two-fold.
@@
 Second, other parts of the library can trigger state transitions during and between TCP user function calls, which can affect the intermediate and final states that a TCP user function can exhibit.
-The TCP user functions were fully translated to SPARK. SPARK bindings were also provided for most of the rest of the library's C functions to allow their invocation from the SPARK code.
+The TCP user functions were fully translated to SPARK.
+SPARK bindings were also provided for most of the rest of the library's C functions to allow their invocation from the SPARK code.
 

### 2021-06-02 SecDev21/paper.tex
fixing e.g.
reviewing the conclusion

@@
 In total, for this work, 50 subprograms have been translated from C to SPARK.
-This accounts for 2266 lines of code, of which 1165 are SPARK logic code (e.g\.~ contracts) used to represent and prove the protocol's specification.
+This accounts for 2266 lines of code, of which 1165 are SPARK logic code (e.g.\ contracts) used to represent and prove the protocol's specification.
 The overall proving time with GNATprove is 9m and 54s on a 12-processor Intel(R) Core(TM) i7-9750H CPU with 2.60GHz speed and 16GB of RAM.
@@
 
-We selectively applied formal verification to a layer of a TCP/IP library, by
-translating TCP user functions into SPARK, and modelling asynchronous events
-with ghost code. This allowed us to detect two bugs in the original C
-implementation, related to memory management and concurrency. This is a step
-towards a more secure implementation. The translated code has been proven to be
-free of run-time errors and to respect the transitions of the TCP state machine
-as specified in RFC 793.
+We selectively applied formal verification to a TCP/IP library layer by translating TCP user functions into SPARK and modeling asynchronous events with ghost code.
+This enabled the detection of two bugs in the original C implementation related to memory management and concurrency.
+The work is a step towards a more secure implementation.
+The translated code has been proven to be free of run-time errors and to respect the transitions of the TCP state machine, specified in RFC 793 \cite{rfc793}.
 
-These results depend upon other layers in the C programming language, which
-were not formally verified to the same level. Instead, we used symbolic
-execution with KLEE to formally verify the contracts used in SPARK to model the
-corresponding behaviors. The principal function of those layers
-is to format packets before they are sent, or parse incoming packets, check their
-integrity and transmit the resulting payload to the corresponding upper-level
-layer. This processing part can be a source of errors and bugs as explain
-in~\cite{Reiher2019RecordFluxFM}.
-As an alternative to manual translation into SPARK, we envision 
-using the RecordFlux DSL to specify the parser and printer of the packets of the
-different protocols, and to generate provable SPARK code from this specification.
-This would complete our current verification of the CycloneTCP stack towards
-the lower layers.
+These results depend upon other layers in the C programming language, which were not formally verified to the same level.
+Instead, we used symbolic execution with KLEE to formally verify the contracts used in SPARK to model the corresponding behaviors.
+The principal function of those layers is to format packets before they are sent, or parse incoming packets, check their integrity, and transmit the resulting payload to the corresponding upper-level layer.
+This processing part can be a source of errors and bugs as explain in~\cite{Reiher2019RecordFluxFM}.
+As an alternative to manual translation into SPARK, we envision using the RecordFlux DSL \cite{Reiher2019RecordFluxFM} to specify the parser and printer of the different protocols' packets and generate provable SPARK code from this specification.
+This would complete our current verification of the CycloneTCP stack towards the lower layers.
 

### 2021-06-03 SecDev21/paper.tex
removing a todo
fixing the memory leak text

@@
 
-\todo[inline]{KG:pending contributions list and fixing the next paragraph. Will be fixed when the rest of the paper is finalized.}
+%\todo[inline]{KG:pending contributions list and fixing the next paragraph. Will be fixed when the rest of the paper is finalized.}
 
@@
 
-\todo[inline]{KG: fixed up to here}
-
 \section{Bugs Captured}
@@
 
-Thanks to the pointer ownership policy adopted by SPARK and detailed in
-section~\ref{sec:pointers}, we have found that a memory leak can happen when
-the connection has been closed and the buffer are cleaning.
+The pointer ownership policy of SPARK, detailed in \Cref{sec:pointers}, enabled the detection of a memory leak when a socket closes.
+This was present due to partially freeing the memory allocated for the buffer that stores the incoming segments.
+This bug was due to using the wrong library function call to free the memory.
+SPARK ownership policy detected this because, after the call to this function, the pointer to the buffer was declared not own anymore, but the memory pointed by that pointer was partially deallocated.
 
+\todo[inline]{KG: fixed up to here}
 \subsubsection{Violation of the TCP protocol}

### 2021-06-03 SecDev21/OryxStack.tex
fixing the second bug text

@@
 
-    \caption{Overview of the TCP/IP CycloneTCP stack in the OSI model. Green parts represent
-             what have been translated in SPARK and in orange the underlying untrusted part written in C.}
+    \caption{Overview of the TCP/IP CycloneTCP stack (OSI model). Green parts have been translated into SPARK. They depend on orange parts written in C.}
     \label{Fig:TcpStack}

### 2021-06-03 SecDev21/paper.tex
fixing the second bug text

@@
 
-\todo[inline]{KG: fixed up to here}
 \subsubsection{Violation of the TCP protocol}
 
-A subtle bug found thanks to SPARK is
-reproduced in the snippet below.
+A subtle bug that violates the protocol's extracted specifications captured by SPARK is reproduced below:
 
@@
 
-In this scenario, after the call to \spark{Tcp_Send}, when the error cases have
-been filtered, the state of the socket can either be \sestab or \sclosew. After the call
-at line~\ref{bugProg:tcpWaitForEvents}, the state of the socket can
-still be \sestab or \sclosew, or it can be \sclosed if a \rst flag has been received when the mutex
-was released.
-At line~\ref{bugProg:tcpWaitForEventsPost}, a \fin segment is sent, which does
-not modify the state of the socket. Finally, at line~\ref{bugProg:tcpChangeState}
-we try to jump to the state \sfwone and thus the transitions $\text{\sclosew}
-\rightarrow \text{\sfwone}$ and $\text{\sclosed} \rightarrow \text{\sfwone}$ are
-possible.
+In the original implementation, after the call to \spark{Tcp_Send}, if no error occurs, the state of the socket can either be \sestab or \sclosew.
+After the call at line~\ref{bugProg:tcpWaitForEvents}, the state of the socket can still be \sestab or \sclosew, or \sclosed if a \rst flag has been received when the mutex was released.
+At line~\ref{bugProg:tcpWaitForEventsPost}, a \fin segment is sent, which does not modify the state of the socket.
+Finally, at line~\ref{bugProg:tcpChangeState} the code tries to move the connection to the \sfwone state, and thus, the transitions $\text{\sclosew} \rightarrow \text{\sfwone}$ and $\text{\sclosed} \rightarrow \text{\sfwone}$ are possible.
 
-But these transitions are not allowed by RFC 793, and thus not allowed by the
-precondition of \spark{TCP_Change_State}. The problem is detected by SPARK and a fix
-has been applied to the SPARK and C implementations of CycloneTCP.
+These transitions are not allowed by RFC 793, and thus, not allowed by the precondition of \spark{TCP_Change_State}, which models all the allowed transitions by the protocol's specification.
+The problem was both corrected in the C original implementation and the SPARK translated code.
 

### 2021-06-03 SecDev21/paper.tex
Rewording, add abstract + contributions

@@
 
-\title{Layered Formal Verification of a TCP/IP Stack}
+\title{Layered Formal Verification of a TCP Stack}
 
@@
 \begin{abstract}
-\todo{Write the abstract.}
+The Transmission Control Protocol (TCP) at the heart of TCP/IP protocol stacks
+is a critical part of our current digital infrastructure. In this article, we
+show how an existing professional-grade open source embedded TCP/IP library can
+benefit from a formally verified TCP reimplementation. Our approach is to apply
+formal verification to the TCP layer only, relying on validated models of the
+lower layers on which it depends.
 \end{abstract}
@@
 In total, for this work, 50 subprograms have been translated from C to SPARK.
-This accounts for 2266 lines of code, of which 1165 are SPARK logic code (e.g.\ contracts) used to represent and prove the protocol's specification.
-The overall proving time with GNATprove is 9m and 54s on a 12-processor Intel(R) Core(TM) i7-9750H CPU with 2.60GHz speed and 16GB of RAM.
+This accounts for 2266 lines of code, of which 1165 are SPARK logic code (e.g.\ contracts) used to represent and prove TCP specification.
+The overall proving time with GNATprove is 10 minutes on a 12-processor Intel(R) Core(TM) i7-9750H CPU with 2.60GHz speed and 16GB of RAM.
 
-%\todo[inline]{KG:pending contributions list and fixing the next paragraph. Will be fixed when the rest of the paper is finalized.}
+The contributions of this article are the following.
+First, we show that formal verification can be applied selectively to a layer of an existing TCP/IP library, by translating the TCP layer to SPARK and modelling the lower layers on which it depends.
+The benefits are three-fold: we found two bugs in the TCP layer, we proved the correction of that layer, and we defined formal contracts for the upper layers to call the TCP layer.
+Second, we showed how the models of the lower layers can be validated by symbolic execution with KLEE on the C code of those layers.
 
@@
 Thus, transitions corresponding to the timeout events can also be triggered by this task.
-A complete description of the protocol's specification can be found in the RFC 793 document~\cite{rfc793}.
+A complete description of TCP specification can be found in the RFC 793 document~\cite{rfc793}.
 
@@
 Then at \cref{code:kleedriver:tcpProcessSegment}, function \texttt{tcpProcessSegment} is executed symbolically by KLEE.
-Finally, the assertion, \texttt{klee\_assert}, at \cref{code:kleedriver:assert} checks if the code leads to a valid state, one that is within the protocol's specification.
-If there is no raised assertion, then the contract represented by the assertion is deemed proved, and the relevant C code respects the protocol's specification regarding that contract.
+Finally, the assertion, \texttt{klee\_assert}, at \cref{code:kleedriver:assert} checks if the code leads to a valid state, one that is within TCP specification.
+If there is no raised assertion, then the contract represented by the assertion is deemed proved, and the relevant C code respects TCP specification regarding that contract.
 The assertion is then converted to a SPARK contract and added to the function \spark{TCP_Process_One_Segment}.
@@
 
+As a result of this work, two bugs were detected and reported to the authors of
+CycloneTCP, and they fixed them in the C implementation.
+
 \subsubsection{Memory leak}
 
-The pointer ownership policy of SPARK, detailed in \Cref{sec:pointers}, enabled the detection of a memory leak when a socket closes.
-This was present due to partially freeing the memory allocated for the buffer that stores the incoming segments.
-This bug was due to using the wrong library function call to free the memory.
-SPARK ownership policy detected this because, after the call to this function, the pointer to the buffer was declared not own anymore, but the memory pointed by that pointer was partially deallocated.
+A memory leak was detected thanks to the pointer ownership policy of SPARK
+detailed in \Cref{sec:pointers}. When closing a socket, the associated memory
+allocated for the buffer that stores incoming segments was only partially
+freed, due to the use of the wrong library function to free the memory.
 
@@
 
-A subtle bug that violates the protocol's extracted specifications captured by SPARK is reproduced below:
+A subtle bug violating TCP specification is reproduced below:
 
@@
 Finally, at line~\ref{bugProg:tcpChangeState} the code tries to move the connection to the \sfwone state, and thus, the transitions $\text{\sclosew} \rightarrow \text{\sfwone}$ and $\text{\sclosed} \rightarrow \text{\sfwone}$ are possible.
-
-These transitions are not allowed by RFC 793, and thus, not allowed by the precondition of \spark{TCP_Change_State}, which models all the allowed transitions by the protocol's specification.
-The problem was both corrected in the C original implementation and the SPARK translated code.
-
-% The above is uncler, the transitions from CLOSED and CLOSED-WAIT are allowed, it's only the transition from ESTABLISHED which is not allowed.
+These transitions are not allowed by RFC 793, and thus, not allowed by the precondition of \spark{TCP_Change_State}, which models all the transitions allowed by TCP specification.
+Thus GNATprove reported an unproved precondition check here.
 

### 2021-06-03 SecDev21/paper.tex
Minor additions

@@
 This accounts for 2266 lines of code, of which 1165 are SPARK logic code (e.g.\ contracts) used to represent and prove TCP specification.
-The overall proving time with GNATprove is 10 minutes on a 12-processor Intel(R) Core(TM) i7-9750H CPU with 2.60GHz speed and 16GB of RAM.
+The overall proving time with GNATprove from the Community 2021 release is 10 minutes on a 12-processor Intel(R) Core(TM) i7-9750H CPU with 2.60GHz speed and 16GB of RAM.
+The tools and code are freely available online.
 

### 2021-06-03 SecDev21/paper.tex
Uniformize the figures style

@@
 \begin{figure}
-\begin{lstlisting}[language=Ada,basicstyle=\footnotesize\ttfamily]
+\begin{lstlisting}[language=Ada,basicstyle=\footnotesize\ttfamily, frame=bottomline]
 procedure Socket_Connect
@@
 \begin{lstlisting}[language=Ada, basicstyle=\footnotesize\ttfamily,
-                    numbers=left, numberstyle=\tiny, escapechar=\%]
+                    numbers=left, numberstyle=\tiny, escapechar=\%,
+                    numbersep=0pt]
 case Sock.State is

### 2021-06-04 SecDev21/paper.tex
Clement's feedback

@@
 
-The original CycloneTCP library provides no guarantees that the implementation respects the TCP functional specification.
+While being extensively tested against industry standard regression test suites, the original CycloneTCP library provides no formal guarantees that the implementation respects the TCP functional specification.
 Thus, there is no assurance that the implementation will only allow valid transitions between the different states that a TCP session can exhibit.

### 2021-06-11 SecDev21/paper.tex
Final English review by Ben

@@
 \and
-\IEEEauthorblockN{Kyriakos Georgious}
+\IEEEauthorblockN{Kyriakos Georgiou}
 \IEEEauthorblockA{\textit{AdaCore \& University of Bristol}}
@@
 By 2030 it is estimated that one trillion Internet of Things (IoT) devices will be deployed and be part of every aspect of our daily life~\cite{IoT_route:ARM_white_paper}.
-At the same time, it is made evident that security for such systems is a major challenge~\cite{Sec_IoT_Challenges:paper,Cyber_Systems_security:paper}.
+At the same time, it is evident that security for such systems is a major challenge~\cite{Sec_IoT_Challenges:paper,Cyber_Systems_security:paper}.
 Any vulnerability in software libraries that are widely used by embedded systems can potentially open the door for compromising the security of billions of deployed IoT devices.
@@
 Layering is an essential design requirement for both models to achieve modularity, flexibility, and abstraction.
-Applications that only need to use the lower levels' functionality can drop all the unnecessary upper levels.
+Applications that only need to use the lower levels' functionality can ignore all the unnecessary upper levels.
 \Cref{Fig:TcpStack} shows the various layers of the OSI model.
@@
 While the \emph{application} level has a wide range of protocols that can be used depending on the application specification, the \emph{transport} layer is heavily dependent on the two dominant protocols: Transmission Control Protocol (TCP) and User Datagram Protocol (UDP).
-Today, almost every networking application, including IoT devices, is using either the TCP or the UDP protocols to traffic data between the Network and the Application layers.
+Today, almost every networking application, including IoT devices, is using either the TCP or the UDP protocols to traffic data between the Network and the Session layers.
 %Thus, the reliability, assurance, and security of these protocols are of paramount importance for critical applications.
@@
 The library's TCP implementation is meant to conform to the RFC 793 protocol specifications~\cite{rfc793}.
-The quality of the CycloneTCP is acknowledged by the AMNESIA:33 report~\cite{AMNESIA33}, which classifies it as one of the most resilient TCP/IP stack.
+The quality of the CycloneTCP is acknowledged by the AMNESIA:33 report~\cite{AMNESIA33}, which classifies it as one of the most resilient TCP/IP stacks.
 
-In total, for this work, 50 subprograms have been translated from C to SPARK.
+In total, for this work, 50 functions have been translated from C to SPARK.
 This accounts for 2266 lines of code, of which 1165 are SPARK logic code (e.g.\ contracts) used to represent and prove TCP specification.
@@
 First, we show that formal verification can be applied selectively to a layer of an existing TCP/IP library, by translating the TCP layer to SPARK and modelling the lower layers on which it depends.
-The benefits are three-fold: we found two bugs in the TCP layer, we proved the correction of that layer, and we defined formal contracts for the upper layers to call the TCP layer.
+The benefits are three-fold: we found and corrected two bugs in the TCP layer, we proved the correction of that layer, and we defined formal contracts for the upper layers to call the TCP layer.
 Second, we showed how the models of the lower layers can be validated by symbolic execution with KLEE on the C code of those layers.
@@
 The role of each of the three tasks can be summarized as follows.
-User calls refer to functions, namely \texttt{open}, \texttt{close}, \texttt{abort}, \texttt{send} and \texttt{receive}, that can be called by the user to control the connection, send, or receive data.
+\textit{User calls} refer to functions, namely \texttt{open}, \texttt{close}, \texttt{abort}, \texttt{send} and \texttt{receive}, that can be called by the user to control the connection, send, or receive data.
 These functions can trigger transitions between the connection's states since they are intended to control the connection.
-In the arriving segments task, the received segments are being processed,
+In the \textit{arriving segments} task, the received segments are being processed,
 and the corresponding messages are sent back.
 Transitions between states can be triggered on the reception of a segment.
-Finally, timers control the timeouts, for example, the retransmission timeout to retransmit a message or the time-wait timeout to close the connection after a specified amount of time elapsed.
+Finally, \textit{timers} control the timeouts, for example, the retransmission timeout to retransmit a message or the time-wait timeout to close the connection after a specified amount of time elapsed.
 Thus, transitions corresponding to the timeout events can also be triggered by this task.
@@
 
-Ada is a general-purpose, procedural programming language that puts great emphasis on the safety and correctness of the program.
+Ada is a general-purpose, procedural and Object-Oriented programming language that puts great emphasis on the safety and correctness of the program.
 Some of the safety characteristics of the language are its strong typing system and its extensive compile-time and runtime checks.
 Common C programming vulnerabilities like using unsafe pointers or improper null termination of strings cannot exist in an Ada program, while issues like buffer overflow or integer overflow can be dynamically captured by Ada’s runtime checks and dealt with via exception handlers.
-Furthermore, Ada introduced contracts, such as preconditions and postconditions, as part of the language’s standard syntax.
+Furthermore, Ada provides contracts, such as preconditions and postconditions, as part of the language’s standard syntax.
 Such contracts are vital for explicitly expressing software and verification requirements in the source code to allow for static analyzers or runtime checks to verify that the stated requirements are met.
 
-By subsetting Ada, the SPARK programming language \cite{mccormick_chapin_2015} emerged to provide the largest possible subset of Ada, which is suitable for functional specification and static verification.
-For instance, in SPARK, expressions should be free from side effects, and aliasing is forbidden (no two variables should share the exact memory location or overlap in memory), including when using pointers thanks to the use of an ownership policy~\cite{dross2020recursive}.
+By subsetting Ada, the SPARK programming language \cite{mccormick_chapin_2015} provides the largest possible subset of Ada that is suitable for functional specification and static verification.
+For instance, in SPARK, expressions are free from side effects, and aliasing is forbidden (no two variables can share the exact memory location or overlap in memory), including when using pointers thanks to the use of an ownership policy~\cite{dross2020recursive}.
 
 \emph{GNATprove} \cite{GNATProve:users_manual} is the tool that provides formal verification for SPARK through two kinds of analyses.
-The first enables flow analysis to check the initialization of variables, look at data dependencies between inputs and outputs of subprograms, and detect useless code.
+The first enables flow analysis to check the initialization of variables, verify data dependencies between inputs and outputs of subprograms, and detect useless code.
 The second uses deductive verification to generate verification conditions for SMT solvers via a weakest-precondition calculus to detect possible runtime errors and violations of functional contracts.
@@
 Instead, the aim is to provide significant security hardening to mostly-C libraries by selective use of SPARK.
-Thus, this work focuses on the hardening of the TCP library areas that its original authors designated as the most vulnerable or crucial to conform to their functional specifications, namely, the Application Programming Interface (API).
+Thus, this work focuses on the hardening of the TCP library areas that its original authors designated as the most vulnerable or crucial to conform to their functional specifications, namely, the API.
 This mainly falls under two categories: a) proving the conformance to the protocol's functional specification for the translated code, discussed in \Cref{sec:verif}, and b) hardening the user's API to enforce its correct usage by users, discussed in \Cref{sec:API}.
@@
 Ada and SPARK offer an easy-to-use mechanism for interfacing with C code.
-The mechanism allows C-written function to be called by Ada/SPARK code and vice versa and for representing shared data objects.
+The mechanism allows C-written functions to be called by Ada/SPARK code and vice versa and for representing shared data objects.
 The memory representations of a shared data object must be the same in both the C and the SPARK code.
-For instance, if a C function is called with a 32-bits integer argument, the SPARK interface of this function should also be called with a 32-bits integer argument.
-It is the responsibility of the user to ensure this.
+For instance, if a C function is called with a 32-bit integer argument, the SPARK interface of this function should also be called with a 32-bits integer argument.
+It is the responsibility of the user to ensure this property.
 
@@
 
-In the CycloneTCP library, pointers are used both as the type of variables to pass objects when calling other functions and as the type of structure fields; for instance, in the socket structure to store other complex data structures and recursive pointer-based data structures.
+In the CycloneTCP library, pointers are used both as the type of variables passed to functions and as the type of structure fields; for instance, in the socket structure to store other complex data structures and recursive pointer-based data structures.
 SPARK uses a pointer ownership model to prevent bugs involving pointers such as memory leak or double free, which imposes some restrictions on how pointers are being manipulated.
-When a memory area is allocated, a pointer points to this memory area, and thus, it is considered ``owned.'' The ownership of the memory can be moved, but the memory must be deallocated before the end of the execution of the program.
+When a memory area is allocated, a pointer points to this memory area, and thus, it is considered ``owned.'' The ownership of the memory can be transferred, but the memory must be deallocated before the end of the execution of the program.
 
@@
 Most of the fields are used in library parts that are not considered by this work, and thus, are not relevant for our verification.
-Thus we use \emph{ghost code} in SPARK to focus the verification effort on the relevant subset of the fields, which also makes the specifications easier to write and understand.
+Thus we use \emph{ghost code} in SPARK (that is, special code only meant for verification and not included in the final executable) to focus the verification effort on the relevant subset of the fields, which also makes the specifications easier to write and understand.
 This is a solution to the classical problem of specifying the \emph{frame condition} for a function, that is, the set of objects possibly modified by the function.
-\emph{Ghost code} is used by GNATprove for analysis but is not being compiled in the final executable.
+\emph{Ghost code} is used by GNATprove for analysis but is not compiled into the final executable.
 More specifically, we define a \emph{model} of the actual socket as a ghost function that can be used in contracts: \spark{Model(Sock)} extracts the fields of interest from socket \spark{Sock}.
@@
 \item The transitions between states must respect the order given by the
-state machine described in \Cref{Fig:statemachine}. In particular, a transition from a state to another is valid only if the state machine allows for a transition between these two states.
+state machine described in \Cref{Fig:statemachine}. In particular, a transition from one state to another is valid only if the state machine allows a transition between these two states.
 \item RFC 793 describes the conditions for the state of the socket which need to hold when a user's function is called~\cite[p.
@@
 
-RFC 793 document provides all the information needed for the allowed transitions between the different TCP possible states.
+RFC 793 provides all the information needed for the allowed transitions between the different TCP possible states.
 For any transitions directly triggered by calling the TCP user functions, SPARK contracts, representing transitions allowed by the functional specification, are embedded within the SPARK translated code.
@@
 
-The concurrent implementation of the TCP protocol allows for multiple interactions between the tasks, both synchronous (with a call to \spark{TCP_Wait_For_Events}) or asynchronous (between two user function calls), and thus, numerous possible state changes at any given moment.
+The concurrent implementation of the TCP protocol allows for multiple interactions between the tasks, both synchronous (with a call to \spark{TCP_Wait_For_Events}) and asynchronous (between two user function calls), and thus, numerous possible state changes at any given moment.
 This makes the functional specification verification significantly challenging, as SPARK does not have a native mode to deal precisely with interactions related to concurrency.
@@
 Then at \cref{code:kleedriver:tcpProcessSegment}, function \texttt{tcpProcessSegment} is executed symbolically by KLEE.
-Finally, the assertion, \texttt{klee\_assert}, at \cref{code:kleedriver:assert} checks if the code leads to a valid state, one that is within TCP specification.
-If there is no raised assertion, then the contract represented by the assertion is deemed proved, and the relevant C code respects TCP specification regarding that contract.
+Finally, the assertion, \texttt{klee\_assert}, at \cref{code:kleedriver:assert} checks if the code leads to a valid state, one that is within the TCP specification.
+If there is no raised assertion, then the contract represented by the assertion is deemed proved, and the relevant C code respects the TCP specification regarding that contract.
 The assertion is then converted to a SPARK contract and added to the function \spark{TCP_Process_One_Segment}.
@@
 One of the most significant problems pointed to by the primary author of the \emph{CycloneTCP} library is the incorrect usage of the library's API.
-More specifically, the users of the library sometimes call the TCP user functions in the wrong order and forget to check the return code after some calls.
+More specifically, users of the library might call the TCP user functions in the wrong order and forget to check the return code after some calls.
 This causes their TCP implementation to behave outside of the functional specification of the protocol.
@@
 The TCP protocol implies a specific order such that the user can call the TCP user functions without breaking the functional specification of the protocol.
-This order is also being conveyed by the protocol's state machine given in \Cref{Fig:statemachine}.
+This order is also conveyed by the protocol's state machine given in \Cref{Fig:statemachine}.
 Ada's preconditions and postconditions are a powerful tool to express such inter-function dependencies, while SPARK technology can be used to guarantee that the assertions always hold.
@@
 If no error occurs, the function \spark{Socket_Connect} sets \spark{Sock.S_Remote_Ip_Addr} to \spark{Remote_Ip_Addr} which is supposed to be initialized when the function is called.
-\spark{Socket_Send} requires to be called with a not null socket such that its field \spark{Remote_Ip_Addr} is initialized.
+\spark{Socket_Send} requires being called with a not null socket such that its field \spark{Remote_Ip_Addr} is initialized.
 The only way to ensure that the precondition hold is to call \spark{Socket_Connect} before \spark{Socket_Send}.
-By analyzing user's code, GNATprove will statically detect if a call to \spark{Socket_Send} is not preceded by a call to \spark{Socket_Connect}.
+By analyzing the user's code, GNATprove will statically detect if a call to \spark{Socket_Send} is not preceded by a call to \spark{Socket_Connect}.
 
@@
 The principal function of those layers is to format packets before they are sent, or parse incoming packets, check their integrity, and transmit the resulting payload to the corresponding upper-level layer.
-This processing part can be a source of errors and bugs as explain in~\cite{Reiher2019RecordFluxFM}.
+This processing part can be a source of errors and bugs as explained in~\cite{Reiher2019RecordFluxFM}.
 As an alternative to manual translation into SPARK, we envision using the RecordFlux DSL \cite{Reiher2019RecordFluxFM} to specify the parser and printer of the different protocols' packets and generate provable SPARK code from this specification.

### 2021-06-11 SecDev21/paper.tex
change back

@@
 While the \emph{application} level has a wide range of protocols that can be used depending on the application specification, the \emph{transport} layer is heavily dependent on the two dominant protocols: Transmission Control Protocol (TCP) and User Datagram Protocol (UDP).
-Today, almost every networking application, including IoT devices, is using either the TCP or the UDP protocols to traffic data between the Network and the Session layers.
+Today, almost every networking application, including IoT devices, is using either the TCP or the UDP protocols to traffic data between the Network and the Application layers.
 %Thus, the reliability, assurance, and security of these protocols are of paramount importance for critical applications.

### 2021-08-26 SecDev21/paper.tex
fix biblio
Corrections for Fig. 3 and Algorithm 1

@@
 It computes the set of possible final states after the completion of an event by calling repeatedly the function \spark{TCP_Process_One_Segment}.
+It is enough to perform only three iterations in the loop of this algorithm because the maximum path between two states that does not require a user interaction has a size of three.
 Both functions have the same behavior, and thus, the contract of the function \spark{TCP_Wait_For_Events_Proof} is also attached to \spark{TCP_Wait_For_Events}.
@@
     (oldSock.state == TCP_STATE_ESTABLISHED) ?
-        sock.state == TCP_STATE_ESTABLISHED ||
-        sock.state == TCP_STATE_CLOSE_WAIT ||
-        sock.state == TCP_STATE_CLOSED :
+        sock->state == TCP_STATE_ESTABLISHED ||
+        sock->state == TCP_STATE_CLOSE_WAIT ||
+        sock->state == TCP_STATE_CLOSED :
     (oldSock.state == ...) ? ... )
@@
 
-These results depend upon other layers in the C programming language, which were not formally verified to the same level.
+These results depend upon other layers written in the C programming language, which were not formally verified to the same level.
 Instead, we used symbolic execution with KLEE to formally verify the contracts used in SPARK to model the corresponding behaviors.

### 2021-08-26 SecDev21/paper.tex
more corrections

@@
 
-In total, for this work, 50 functions have been translated from C to SPARK.
-This accounts for 2266 lines of code, of which 1165 are SPARK logic code (e.g.\ contracts) used to represent and prove TCP specification.
-The overall proving time with GNATprove from the Community 2021 release is 10 minutes on a 12-processor Intel(R) Core(TM) i7-9750H CPU with 2.60GHz speed and 16GB of RAM.
-The tools and code are freely available online.
-
 The contributions of this article are the following.
@@
 Second, we showed how the models of the lower layers can be validated by symbolic execution with KLEE on the C code of those layers.
+The tools and code\footnote{\url{https://github.com/Adacore/Http_Cyclone}} are freely available online.
 
@@
 TCP is a connection-oriented protocol; a connection between the sender and the receiver must be established before transmitting any data.
-Furthermore, TCP is a reliable protocol since it guarantees the delivery of all the messages and that they are delivered in the order they were sent.
+Furthermore, TCP is a reliable protocol since it guarantees the delivery of all the messages in absence of a network outage and that they are delivered in the order they were sent.
 It also provides error-checking mechanisms to discard and recover from corrupted data.
@@
 Thus, it has tasks dedicated to the user's code, to the timers, and to the processing of incoming segments.
-Moreover, the user can manipulate a socket through the TCP's library Application Programming Interface (API) to perform various TCP operations, such as data transmission.
+These  three tasks communicate together, either synchronously: a task requests for an action to be done by another task; or asynchronously: some actions are performed on a socket by a task between two uses of the socket.
+All the uses of a socket are protected by a global mutex, common to all the sockets.
+It ensures that only one task can modify a socket at the same time.
 
@@
 The memory representations of a shared data object must be the same in both the C and the SPARK code.
-For instance, if a C function is called with a 32-bit integer argument, the SPARK interface of this function should also be called with a 32-bits integer argument.
+For instance, if a C function is called with a 32-bits integer argument, the SPARK interface of this function should also be called with a 32-bits integer argument.
+It is also true for the structures.
 It is the responsibility of the user to ensure this property.
+Nevertheless, the compiler GCC offers an option \texttt{-fdump-ada-spec} that helps the user in writing correct Ada specifications by generating them directly from the C code.
+The layout of the structures are then the same by construction in C and SPARK.
 
@@
 This states that the fields of \spark{Sock} used by \spark{Model} are not changed by the function call, except for the field \spark{State} that is equal to \spark{TCP_STATE_CLOSED} after the call.
+Nothing is stated on the field that are not selected by \spark{Model(Sock)}.
+In particular they can have changed during the execution of the function or not.
 
@@
 This makes the functional specification verification significantly challenging, as SPARK does not have a native mode to deal precisely with interactions related to concurrency.
-To address this, we introduced SPARK functions with contracts that model how the different possible interactions can modify the state of a socket.
+To address this, we avoid using concurrency in SPARK and instead we introduced sequential SPARK functions with contracts that model how the different possible interactions can modify the state of a socket.
 Finding a correct and precise contract for \spark{TCP_Wait_For_Events} was challenging but needed for the verification of the TCP user functions.
@@
 If there is no raised assertion, then the contract represented by the assertion is deemed proved, and the relevant C code respects the TCP specification regarding that contract.
-The assertion is then converted to a SPARK contract and added to the function \spark{TCP_Process_One_Segment}.
+The assertion is then manually converted to a SPARK contract and added to the function \spark{TCP_Process_One_Segment}.
 This enables the SPARK prover to reason about the possible side-effects caused by the C-code on the state of a TCP connection and thus, enables the functional correctness verification of the TCP user functions.
+
+Using KLEE on the function that process incoming segments helped us to make faster the verification of the user function.
+Applied to the user function, this method would not have been enough.
+KLEE can only unroll loops when the loops present in the user functions can finish after a high number of iterations which may leads to a combinatorial explosion in the number of paths explored by KLEE.
+KLEE has other limitations that were useful in SPARK for our verification like the detection of memory leaks.
 
@@
 Thus, preconditions and postconditions were introduced to model a partial order on the calls to the TCP user functions.
+The pre- and postconditions in Ada can be enabled with a pragm in order to be checked at the runtime to assure, even if they are not previously verified with SPARK, that they are still true when the program is running.
 
@@
 
+\section{Results}
+
+In total, for this work, 50 functions have been translated from C to SPARK.
+This accounts for 2266 lines of code, of which 1165 are SPARK logic code (e.g.\ contracts) used to represent and prove TCP specification.
+All the socket functions and the user functions of TCP have been translated and proved in SPARK, then in particular \spark{TCP_Init}, \spark{TCP_Connect}, \spark{TCP_Listen}, \spark{TCP_Send}, \spark{TCP_Receive}, \spark{TCP_Shutdown}, \spark{TCP_Accept} and \spark{TCP_Abort}.
+The corresponding socket function that makes the interface between Socket function have also been fully translated to SPARK.
+At the end, 25\% of the functions that implement the TCP protocol have been rewritten in SPARK.
+
+The overall proving time with GNATprove from the Community 2021 release is 10 minutes on a 12-processor Intel(R) Core(TM) i7-9750H CPU with 2.60GHz speed and 16GB of RAM.
+
+% I'll try to add comparison between the time taken by the C and the Ada code.
+
 \section{Conclusion and future work}

### 2021-08-30 SecDev21/paper.tex
fix typos

@@
 Thus, it has tasks dedicated to the user's code, to the timers, and to the processing of incoming segments.
-These  three tasks communicate together, either synchronously: a task requests for an action to be done by another task; or asynchronously: some actions are performed on a socket by a task between two uses of the socket.
+These  three tasks communicate together, either synchronously: a task requests for an action to be done by another task; or asynchronously: some actions are performed on a socket by a task between two uses of the socket by another task.
 All the uses of a socket are protected by a global mutex, common to all the sockets.
@@
 For instance, if a C function is called with a 32-bits integer argument, the SPARK interface of this function should also be called with a 32-bits integer argument.
-It is also true for the structures.
+It is also true for record types.
 It is the responsibility of the user to ensure this property.
 Nevertheless, the compiler GCC offers an option \texttt{-fdump-ada-spec} that helps the user in writing correct Ada specifications by generating them directly from the C code.
-The layout of the structures are then the same by construction in C and SPARK.
+The layout of the record types are then the same by construction in C and SPARK.
 
@@
 This states that the fields of \spark{Sock} used by \spark{Model} are not changed by the function call, except for the field \spark{State} that is equal to \spark{TCP_STATE_CLOSED} after the call.
-Nothing is stated on the field that are not selected by \spark{Model(Sock)}.
+Nothing is stated on the fields that are not selected by \spark{Model(Sock)}.
 In particular they can have changed during the execution of the function or not.
@@
 
-The concurrent implementation of the TCP protocol allows for multiple interactions between the tasks, both synchronous (with a call to \spark{TCP_Wait_For_Events}) and asynchronous (between two user function calls), and thus, numerous possible state changes at any given moment.
+The concurrent implementation of the TCP protocol allows for multiple interactions between the tasks, both synchronous (with a call to \spark{TCP_Wait_For_Events}) and asynchronous (between two user function calls), and thus, numerous possible state changes are possible at any given moment.
 This makes the functional specification verification significantly challenging, as SPARK does not have a native mode to deal precisely with interactions related to concurrency.
@@
 The assertion is then manually converted to a SPARK contract and added to the function \spark{TCP_Process_One_Segment}.
-This enables the SPARK prover to reason about the possible side-effects caused by the C-code on the state of a TCP connection and thus, enables the functional correctness verification of the TCP user functions.
+This enables the SPARK prover to reason about the possible side-effects caused by the C code on the state of a TCP connection and thus, enables the functional correctness verification of the TCP user functions.
 
@@
 Thus, preconditions and postconditions were introduced to model a partial order on the calls to the TCP user functions.
-The pre- and postconditions in Ada can be enabled with a pragm in order to be checked at the runtime to assure, even if they are not previously verified with SPARK, that they are still true when the program is running.
+The preconditions in Ada can be enabled at runtime to protect against a violation when called from code that was not proved, for example from C code.
 
@@
 In total, for this work, 50 functions have been translated from C to SPARK.
-This accounts for 2266 lines of code, of which 1165 are SPARK logic code (e.g.\ contracts) used to represent and prove TCP specification.
-All the socket functions and the user functions of TCP have been translated and proved in SPARK, then in particular \spark{TCP_Init}, \spark{TCP_Connect}, \spark{TCP_Listen}, \spark{TCP_Send}, \spark{TCP_Receive}, \spark{TCP_Shutdown}, \spark{TCP_Accept} and \spark{TCP_Abort}.
+This accounts for 2266 lines of code, of which 1165 are SPARK logic code (e.g.\ contracts) used to represent and prove the TCP specification.
+All the socket functions and the user functions of TCP have been translated and proved in SPARK, in particular \spark{TCP_Init}, \spark{TCP_Connect}, \spark{TCP_Listen}, \spark{TCP_Send}, \spark{TCP_Receive}, \spark{TCP_Shutdown}, \spark{TCP_Accept} and \spark{TCP_Abort}.
 The corresponding socket function that makes the interface between Socket function have also been fully translated to SPARK.

### 2021-08-30 SecDev21/paper.tex
clarifications

@@
 
-Using KLEE on the function that process incoming segments helped us to make faster the verification of the user function.
-Applied to the user function, this method would not have been enough.
-KLEE can only unroll loops when the loops present in the user functions can finish after a high number of iterations which may leads to a combinatorial explosion in the number of paths explored by KLEE.
-KLEE has other limitations that were useful in SPARK for our verification like the detection of memory leaks.
+Our choice of combining deductive verification with SPARK and symbolic execution with KLEE was driven by practical considerations which are typical of industrial applications.
+Using KLEE on user functions would be difficult due to the presence of loops with a possible high number of iterations, which symbolic execution will need to unroll, leading to combinatorial explosion.
+Using SPARK on the function that processes incoming segments would be desirable, and may be done in the future, but was not doable in the timeframe of this project.
+Note that KLEE has other limitations compared to SPARK in terms of the verification scope.
+For example it is lacking a detection of memory leaks, which turned out to be useful in SPARK, see section~\ref{sec:memory-leak}.
 
@@
 
+\label{sec:memory-leak}
 \subsubsection{Memory leak}
@@
 All the socket functions and the user functions of TCP have been translated and proved in SPARK, in particular \spark{TCP_Init}, \spark{TCP_Connect}, \spark{TCP_Listen}, \spark{TCP_Send}, \spark{TCP_Receive}, \spark{TCP_Shutdown}, \spark{TCP_Accept} and \spark{TCP_Abort}.
-The corresponding socket function that makes the interface between Socket function have also been fully translated to SPARK.
+The corresponding socket functions from the Socket interface have also been fully translated to SPARK.
 At the end, 25\% of the functions that implement the TCP protocol have been rewritten in SPARK.

### 2021-09-08 SecDev21/paper.tex
adding acknowledgment

@@
 
+\section*{Acknowledgments}
+This research is part of the ``High-Integrity Complex Large Software and Electronic Systems'' (HICLASS) project that is supported by the Aerospace Technology Institute (ATI) Programme, a joint Government and industry investment to maintain and grow the UK's competitive position in civil aerospace design and manufacture, under grant agreement No.
+113213.
+We also thank the reviewers for their valuable feedback.
+
 \bibliographystyle{IEEEtran}

### 2021-09-08 SecDev21/paper.tex
Comparison between C & SPARK code

@@
 \definecolor{urlcolor}{RGB}{148,39,38}
+
+\usepackage{multirow}
+\usepackage{array}
 
@@
 \begin{lstlisting}[language=C, basicstyle=\footnotesize\ttfamily, numbers=left,
-    numberstyle=\tiny, frame=bottomline, escapeinside={(*@}{@*)}]
+    numberstyle=\tiny, frame=bottomline, escapeinside={(*@}{@*)}, numbersep=2pt]
 // Create a fake incoming segment
@@
 
+\begin{table}[!t]
+    \caption{Comparison of the number of ASM instructions between the C and the SPARK implementation}
+    \label{table:instr}
+    \renewcommand{\arraystretch}{1.3}
+    \centering
+    \begin{tabular}{lcc}
+    \hline
+    \multirow{2}{*}{Function}&\multicolumn{2}{c}{Number of ASM instructions} \\
+    &C&SPARK\\ \hline
+    \spark{TCP_Listen}   & 18                  & 11                   \\
+    \spark{TCP_Accept}   & 153                 & 167                  \\
+    \spark{TCP_Connect}  & 103                 & 114                  \\
+    \spark{TCP_Send}     & 94                  & 124                  \\
+    \spark{TCP_Receive}  & 113                 & 153                  \\
+    \spark{TCP_Shutdown} & 107                 & 122                  \\
+    \spark{TCP_Abort}    & 42                  & 50                   \\ \hline
+    \end{tabular}
+\end{table}
+
 In total, for this work, 50 functions have been translated from C to SPARK.
@@
 
-% I'll try to add comparison between the time taken by the C and the Ada code.
+Comparing the performance of each C and SPARK functions can be random and might not reflect the real time taken by the functions due to the network latency.
+Instead, we have reproduced in \Cref*{table:instr} the number of ARM assembly instructions of the TCP functions written in C and in SPARK.
+We observe that the C functions contain less instructions and then the performance of a C code using the TCP functions will be better than the performance of the same code written in SPARK.
 

### 2021-09-08 SecDev21/paper.tex
taille du code

@@
     \centering
-    \begin{tabular}{lcc}
+    \begin{tabular}{lccc}
     \hline
-    \multirow{2}{*}{Function}&\multicolumn{2}{c}{Number of ASM instructions} \\
-    &C&SPARK\\ \hline
-    \spark{TCP_Listen}   & 18                  & 11                   \\
-    \spark{TCP_Accept}   & 153                 & 167                  \\
-    \spark{TCP_Connect}  & 103                 & 114                  \\
-    \spark{TCP_Send}     & 94                  & 124                  \\
-    \spark{TCP_Receive}  & 113                 & 153                  \\
-    \spark{TCP_Shutdown} & 107                 & 122                  \\
-    \spark{TCP_Abort}    & 42                  & 50                   \\ \hline
+    \multirow{2}{*}{Function}&\multicolumn{2}{c}{Number of ASM instructions}&\multirow{2}{*}{$\Delta$ num. of instr.} \\
+    &C&SPARK & \\ \hline
+    \spark{TCP_Listen}   & 18 & 11                   & $-$39\%\\
+    \spark{TCP_Accept}   & 153 & 167                  & +9\%\\
+    \spark{TCP_Connect}  & 103 & 114                  & +10\%\\
+    \spark{TCP_Send}     & 94 & 124                  & +31\%\\
+    \spark{TCP_Receive}  & 113 & 153                  & +35\%\\
+    \spark{TCP_Shutdown} & 107 & 122                  & +14\%\\
+    \spark{TCP_Abort}    & 42 & 50                   & +19\%\\ \hline
     \end{tabular}
@@
 Instead, we have reproduced in \Cref*{table:instr} the number of ARM assembly instructions of the TCP functions written in C and in SPARK.
+Both C and SPARK code have been compiled with gcc by enabling the option \texttt{-Os} that performs many optimizations similar to \texttt{-O2} that do not increase the size of the generated code.
 We observe that the C functions contain less instructions and then the performance of a C code using the TCP functions will be better than the performance of the same code written in SPARK.

### 2021-09-09 SecDev21/paper.tex
final changes

@@
 Early work formalized and proved some key properties of the TCP protocol~\cite{smith1996formal}.
-More recently, researchers have formally specified TCP, UDP and the Sockets API in HOL and tested existing libraries against the formal specification~\cite{tcp2005sigcomm,ridge2008rigorous}.
-Others have formally specified and implemented the SSL/TLS protocol in F* and integrated the resulting implementation in existing libraries~\cite{bhargavan2013implementing}.
+In later work, researchers have formally specified TCP, UDP and the Sockets API in HOL and tested existing libraries against the formal specification~\cite{tcp2005sigcomm,ridge2008rigorous}.
+More recently, others have formally specified and implemented the SSL/TLS protocol in F* and integrated the resulting implementation in existing libraries~\cite{bhargavan2013implementing}.
 To the best of our knowledge, no prior work attempted to formally verify an implementation of TCP, although the authors of~\cite{ridge2008rigorous} suggest that it would possible to extract an actual implementation in Haskell from their specification.
@@
 The library's TCP implementation is meant to conform to the RFC 793 protocol specifications~\cite{rfc793}.
-The quality of the CycloneTCP is acknowledged by the AMNESIA:33 report~\cite{AMNESIA33}, which classifies it as one of the most resilient TCP/IP stacks.
+The quality of CycloneTCP is acknowledged by the AMNESIA:33 report~\cite{AMNESIA33}, which classifies it as one of the most resilient TCP/IP stacks.
 
@@
 Second, we showed how the models of the lower layers can be validated by symbolic execution with KLEE on the C code of those layers.
-The tools and code\footnote{\url{https://github.com/Adacore/Http_Cyclone}} are freely available online.
+The tools\footnote{\url{https://www.adacore.com/download}} and code\footnote{\url{https://github.com/Adacore/Http_Cyclone}} are freely available online.
 
@@
 It computes the set of possible final states after the completion of an event by calling repeatedly the function \spark{TCP_Process_One_Segment}.
-It is enough to perform only three iterations in the loop of this algorithm because the maximum path between two states that does not require a user interaction has a size of three.
+It is enough to perform only three iterations in the loop of this algorithm because the maximum path between two states from the TCP automaton, given in \Cref{Fig:statemachine}, that does not require a user interaction has a size of three.
 Both functions have the same behavior, and thus, the contract of the function \spark{TCP_Wait_For_Events_Proof} is also attached to \spark{TCP_Wait_For_Events}.
@@
 Symbolic execution is a means of executing programs with symbolic values rather than concrete values \cite{symbolic_exec_survey:2018}.
-Exhaustive symbolic execution can be considered sound and complete (it prevents false negatives and false positives), and thus, it is complementary to formal verification.
+Exhaustive symbolic execution can be considered sound and complete (it prevents false negatives and false positives), and thus, it is on par with deductive verification.
 In \cite{vanoverberghe2008using}, the authors present how symbolic execution can be used to improve deductive verification, and in \cite{kassios2012comparing} the authors demonstrated the technique's value in inferring contracts.
@@
    Post =>
-    if Sock.S_Type = SOCKET_TYPE_STREAM then
+    (if Sock.S_Type = SOCKET_TYPE_STREAM then
       (if Error = NO_ERROR then
@@
     else
-      Sock.S_Remote_Ip_Addr = Remote_Ip_Addr
+      Sock.S_Remote_Ip_Addr = Remote_Ip_Addr)
 
@@
 \spark{Socket_Send} requires being called with a not null socket such that its field \spark{Remote_Ip_Addr} is initialized.
-The only way to ensure that the precondition hold is to call \spark{Socket_Connect} before \spark{Socket_Send}.
+The only way to ensure that the precondition holds is to call \spark{Socket_Connect} before \spark{Socket_Send}.
 By analyzing the user's code, GNATprove will statically detect if a call to \spark{Socket_Send} is not preceded by a call to \spark{Socket_Connect}.
@@
 
-As a result of this work, two bugs were detected and reported to the authors of
-CycloneTCP, and they fixed them in the C implementation.
+As a result of this work, two bugs were detected and reported to the fourth author of this article, also main developer of CycloneTCP, and were fixed in the C implementation.
 
@@
 
-Comparing the performance of each C and SPARK functions can be random and might not reflect the real time taken by the functions due to the network latency.
+Comparing the performance of C and SPARK functions can be very volatile and might not reflect the real time taken by the functions due to network latency.
 Instead, we have reproduced in \Cref*{table:instr} the number of ARM assembly instructions of the TCP functions written in C and in SPARK.
-Both C and SPARK code have been compiled with gcc by enabling the option \texttt{-Os} that performs many optimizations similar to \texttt{-O2} that do not increase the size of the generated code.
-We observe that the C functions contain less instructions and then the performance of a C code using the TCP functions will be better than the performance of the same code written in SPARK.
+Both C and SPARK code have been compiled with GCC by enabling the option \texttt{-Os} that performs many optimizations similar to \texttt{-O2} that do not increase the size of the generated code.
+We observe that the C functions contain fewer instructions than their SPARK counterparts, hence it is expected that their runtime performance will be better.
+However, we don't expect this to be of practical significance, given the modest amount of time that an application will spend running these functions.
 
